### 2020-05-21 battery-preprocessing_v5.Rmd
this includes lower number of trails needed for inclusion

---
title: "Online Laterality Battery Pre-Processing - short version"
author: "Adam Parker + Dorothy Bishop"
date: "updated April 22nd 2020"
output:
  pdf_document: default
  html_document: default
---
NB. 2 changes Apr 22nd by DB: 
a) all data now in a Data folder, so when reading need to specify that; 
b) *demographics* data frame renamed to *allsum*.

This R Markdown file prcoesses the various components of the Online Laterality Battery which is implemented online via Gorilla.sc.

This version updated on 1st Feb 2020 to make script more reproducible and compact by using functions for repeated operations.
Adapted in April 2020 for processing dataset on 300 people

The script pulls in csv files and processes them so that pre-registered analyses can be conducted (pre-registered on the OSF: ). First, questionnaires are analyses followed by chimeric, dichotic, visual half field and tapping tasks. This is done for time 1 and time 2. A "full" laterality index output is then appended to a file with the demographics. This is then used in subsequent R markdown scripts.

This script analyses the following: 
1) Questionnaires
  - demographics
  - Miles test
  - Porta test
  - Edinburgh Handedness Inventory
  - lexTALE
2) Chimeric face task
3) Dichotic listening
4) Rhyme detection task
5) Key Pressing Task (finger dexterity)
And then produces summary of demog and results for the final sample

# Questionnaires

This R Markdown processes the following questionnaires from the Online battery:

- demographics
- Miles test
- Porta test
- Edinburgh Handedness Inventory
- lexTALE

Once processed, a file is made.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#NB for original data setwd("~/Dropbox/Online LDT/300 Ps")
# libraries 
library(readr)
library(yarrr)
library(dplyr)
library(tidyr)
library(car)
library(ggplot2)
library(knitr)
require(readr)  # for read_csv()
require(dplyr)  # for mutate()
require(tidyr)  # for unnest()
require(purrr)  # for map(), reduce()
library(GGally)
require(stringr) # string matching
library(ggpirate)
library(corrr) #easier correlations!
```

First, the demographics is processed.This gives us a participant ID, age, gender, footedness, and handedness.

```{r demograhics, warning=FALSE}
demo_dat <- read_csv("./data/Demographics.csv", na = c("empty", "NA")) # read in data
# relabel 
demo_dat$subject <- demo_dat$Participant.Public.ID
demo_dat$subject <- as.factor(demo_dat$subject)
# count and print subjects
nsub <-length(levels(demo_dat$subject))
print(paste0("subjects: ", nsub))

# create data frame
allsum <- data.frame(matrix(ncol = 7, nrow = nsub))
  colnames(allsum) <- c("subject", "age", "gender", "footedness", "handedness", "bilingual", "strongest.language")
myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  subname <- levels(demo_dat$subject)[i] # find subject ID
  myrows <- which(demo_dat$subject==subname) # select rows for this subject
  tmp <- data.frame(demo_dat[myrows,])

    myrow <- myrow+1

    allsum$subject[myrow] <- subname
    allsum$age[myrow] <- as.numeric(tmp[tmp$Question.Key == "Age",]$Response)
    allsum$gender[myrow] <- tmp[tmp$Question.Key == "Gender",]$Response
    allsum$footedness[myrow] <- tmp[tmp$Question.Key == "footedness",]$Response
    allsum$handedness[myrow] <- tmp[tmp$Question.Key == "categorical_hand",]$Response
    allsum$bilingual[myrow] <- tmp[tmp$Question.Key == "bilingual",]$Response
    allsum$strongest.language[myrow] <- tolower(tmp[tmp$Question.Key == "bilingual-text",]$Response)

    allsum$bilingual <- as.character(allsum$bilingual)
    allsum$bilingual <- gsub('. My strongest language is:', '', allsum$bilingual)
}

#create a numerical code for handedness - can be used in plots to distinguish handedness groups (eg if pch set to handpch, get circles and triangles)
allsum$handpch <-1
allsum$handpch[allsum$handedness=='Left'] <-2
```

Now we import and append the Porta and Miles Test to the allsum data frame.

```{r sightednesss, warning=FALSE}
miles_dat <- read_csv("./data/Miles.csv", na = c("empty", "NA")) # read in data
porta_dat <- read_csv("./data/Porta.csv", na = c("empty", "NA")) # read in data
# add test ID
miles_dat$test_ID <- "miles"
porta_dat$test_ID <- "porta"
# merge the two
sight_dat <- rbind(miles_dat, porta_dat)
# relabel 
sight_dat$subject <- sight_dat$Participant.Public.ID
sight_dat$subject <- as.factor(sight_dat$subject)

# count and print subjects
nsub <-length(levels(sight_dat$subject))
print(paste0("subjects: ", nsub))

# create data frame
sightedness <- data.frame(matrix(ncol = 3, nrow = nsub))
  colnames(sightedness) <- c("subject", "miles", "porta")
myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  subname <- levels(sight_dat$subject)[i] # find subject ID
  myrows <- which(sight_dat$subject==subname) # select rows for this subject
  tmp <- data.frame(sight_dat[myrows,])

    myrow <- myrow+1

    sightedness$subject[myrow] <- subname
    sightedness$miles[myrow] <- tmp[tmp$Question.Key == "response-2",]$Response
    sightedness$porta[myrow] <- tmp[tmp$Question.Key == "response-2",]$Response
  }

# now merge sightedness and allsum
allsum <- merge(allsum, sightedness, by= "subject")
```

Now we add the EHI. Here we score and plot the Edinburgh handedness inventory (Oldfield, 1971) in its short form.

```{r EHI, warning=FALSE}
EHI <- read_csv("./data/EHI.csv", na = "empty") # read in data

# select only responses
EHI <- EHI[EHI$Question.Key == "response-2-quantised" | EHI$Question.Key == "response-3-quantised" | 
           EHI$Question.Key == "response-4-quantised" | EHI$Question.Key == "response-5-quantised" |
           EHI$Question.Key == "response-6-quantised" | EHI$Question.Key == "response-7-quantised" | 
           EHI$Question.Key == "response-8-quantised" | EHI$Question.Key == "response-9-quantised" |
           EHI$Question.Key == "response-10-quantised" | EHI$Question.Key == "response-11-quantised" | 
           EHI$Question.Key == "response-12-quantised" | EHI$Question.Key == "response-13-quantised" |
           EHI$Question.Key == "response-14-quantised" | EHI$Question.Key == "response-15-quantised" | 
           EHI$Question.Key == "response-16-quantised" | EHI$Question.Key == "response-17-quantised" |
           EHI$Question.Key == "response-18-quantised" | EHI$Question.Key == "response-19-quantised" | 
           EHI$Question.Key == "response-20-quantised",]

# recode variabels
EHI$subject <- EHI$Participant.Public.ID
EHI$subject <- as.factor(EHI$subject)

# reduce data
# new data
meaningful <- c("subject", "Question Key", "Response") # select wanted columns
c<-which(names(EHI) %in% meaningful) #find colnumbers of unwanted
EHI <-EHI[,c] #remove unwanted columns
EHI <- na.omit(EHI) # remove NAs

# score responses according to the original scoring method in Oldifeld (1971)
EHI$right_hand <- 0
EHI$left_hand <- 0
for (r in 1:nrow(EHI)) {
  if(EHI$Response[r] == 1){
    EHI$right_hand[r]= 2 
  } else {
    if(EHI$Response[r] == 2){
      EHI$right_hand[r]= 1
    } else {
      if(EHI$Response[r] == 4){
        EHI$left_hand[r]= 1
      } else {
        if(EHI$Response[r] == 5){
          EHI$left_hand[r]= 2 
        } else {
          (EHI$right_hand[r]= 1) & (EHI$left_hand[r]= 1)
        }
      }
    }
  }
}

# re-reduce data
# new data
meaningful <- c("subject", "right_hand", "left_hand") # select wanted columns
c<-which(names(EHI) %in% meaningful) #find colnumbers of unwanted
EHI <-EHI[,c] #remove unwanted columns
EHI <- na.omit(EHI) # remove NAs

# find sum for left and right for each participant 
# create hand data 
nsub <-length(levels(EHI$subject))
EHI2 <- data.frame(matrix(ncol = 3, nrow = nsub))
  colnames(EHI2) <- c("subject", "right_hand", "left_hand")

myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  myrow <- myrow+1
  subname <- levels(EHI$subject)[i] # find subject ID
  myrows <- which(EHI$subject==subname) # select rows for this subject
  tmp <- data.frame(EHI[myrows,])

    EHI2$subject[myrow] <- subname # add row for subject
    EHI2$right_hand[myrow] <- sum(tmp$right_hand) # add row for hand
    EHI2$left_hand[myrow] <- sum(tmp$left_hand) # add row for number of buttons pressed
  }

# This is a summary of the EHI
for_index2 <- 
  EHI2 %>% 
  group_by(subject) %>%
  mutate(index_EHI= ((right_hand-left_hand)/(right_hand + left_hand))*100)

# now create a data frame only for index 
EHI_index <- select(for_index2, "subject", "index_EHI")
# merge with allsum
allsum <- merge(allsum, EHI_index, by= "subject")

#now plot EHI in relation to handedness (modified by DB)

eplot<-ggplot(allsum, aes(handpch,index_EHI, colour=gender)) +
  geom_jitter(width=.25,height=2)+
  ggtitle("Handedness index on the Edinburgh Handedness Inventory")
eplot
#This is not perfect, but gives and idea to check EHI is valid etc.

ggplot(allsum, aes(x=index_EHI)) + geom_density(aes(group=handpch,fill=handpch)) + theme_classic()

#NB one supposed right-hander has index_EHI around -50. 
#I will recode this one, as this seems an error.

oddcase <- intersect(which(allsum$index_EHI<0),which(allsum$handedness=='Right'))
allsum$handedness[oddcase] <- 'Left'
allsum$handpch[oddcase] <- 1
```

Next, the lexTale is preprocesssed. This is done using the weighted average for the test.

```{r lexTALE, warning=FALSE}
# read data
lexTALE <- read_csv("./data/LexTALE.csv", na = "empty")
# filter
lexTALE <- lexTALE[lexTALE$Attempt == 1,]
lexTALE <- lexTALE[lexTALE$display == "Task",]
# rename variables
lexTALE$subject <- lexTALE$Participant.Public.ID
lexTALE$accuracy <- lexTALE$Correct
lexTALE$condition <- lexTALE$ANSWER
# code as factors
lexTALE$subject <- as.factor(lexTALE$subject)
lexTALE$condition <- as.factor(lexTALE$condition)
  levels(lexTALE$condition) <- c("non-word", "word")

# create data frame
nsub <-length(levels(lexTALE$subject))
lexTale_score <- data.frame(matrix(ncol = 4, nrow = nsub))
  colnames(lexTale_score) <- c("subject", "Word", "nonWord", "lexTALE")
myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  subname <- levels(lexTALE$subject)[i] # find subject ID
  myrows <- which(lexTALE$subject==subname) # select rows for this subject
  tmp <- data.frame(lexTALE[myrows,])

    myrow <- myrow+1

    lexTale_score$subject[myrow] <- subname
    lexTale_score$Word[myrow] <- sum(tmp[tmp$condition == "word",]$accuracy)
    lexTale_score$nonWord[myrow] <- sum(tmp[tmp$condition == "non-word",]$accuracy)
    lexTale_score$lexTALE[myrow] <- ((sum(tmp[tmp$condition == "word",]$accuracy)/40*100) + (sum(tmp[tmp$condition == "non-word",]$accuracy)/20*100)) / 2

}

# plot
x<-ggplot(lexTale_score, aes(x=lexTALE, y=subject, fill=subject)) +
  geom_dotplot(binaxis='y', stackdir='center') + xlim(0, 100) +
  ggtitle("Lextale") + geom_vline(xintercept = 80, linetype="dashed", color = "black", size=1.5)
x + theme_bw() + theme(legend.position = "none") + theme(axis.text.y = element_blank())

# now create a data frame only for index 
lexTALE <- select(lexTale_score, "subject", "lexTALE")
# now merge with allsum
allsum <- merge(allsum, lexTALE, by= "subject")
# now write the full allsum 
write.csv(allsum, file = "./allsum_out.csv")

ggplot(allsum, aes(x=lexTALE)) + geom_density(aes(group=handpch,fill=handpch)) + theme_classic()

hist(allsum$lexTALE)
abline(v=80,col='red')

paste0("No. of participants below LexTALE cuttoff: ", sum(allsum$lexTALE < 80))

# remove those who didn't pass the LexTALE
allsum <- allsum[allsum$lexTALE >= 80,]
```

# Read in raw data for day 1 and 2

Before doing anything with the data, we now pull in all the data from all nodes and combine them from a sub folder in the directory. Participants, identified by random letter strings, are stacked on top of each other. We start by reading the data and cutting out unwanted information.

```{r readraw,include= FALSE, warning=FALSE}
# Set directory
for (d in 1:2){
  data_file <- "./data/combined/day1_combined.csv"
  if (d ==2){
    data_file <- "./data/combined/day2_combined.csv"
  }
  #we will assign raw data to all_dat2 (day2), but then reassign to all_dat1 on first run through loop, so we have correct data for each day
all_dat2 <- read.csv(data_file, na.strings = "NA",stringsAsFactors=F)
# cleaning - do operations that apply for all tasks
all_dat2 %>% filter(Attempt >= 1) # only answers 
# write as factors
all_dat2$subject <- as.factor(trimws(all_dat2$Participant.Public.ID))
all_dat2$RT <- as.numeric(as.character(all_dat2$Reaction.Time))
all_dat2$Trial <-as.numeric(all_dat2$Trial.Number) #useful if we want to restrict analysis to subset of trials so retain this variable
all_dat2$Task.Name <- as.factor(all_dat2$Task.Name)
all_dat2$word<-substring(all_dat2$word,66,(nchar(all_dat2$word)-4)) #remove format from written word stimuli
if (d==1)
{all_dat1 <- all_dat2}
}

all_dat2$word_OG<-NA #create dummy column just to make it easier to align columns
collist <- colnames(all_dat1)
all_dat2 <- all_dat2[,collist]

#Check for overlap in subjects between the two sessions
which(levels(all_dat1$subject) %in% levels(all_dat2$subject))
```

# Functions

This is a function that can apply to all tasks; user specifies the original file to start from, the wanted columns, and the specific task to select.

```{r genericprocessing} 
# This function just returns a subfile with just the necessary rows/cols
  procdata <- function(origfile,wanted,task,disp){
  nufile <- origfile[origfile$Task.Name==task,] #restrict consideration to this task
  nufile <- origfile[origfile$display==disp,] #restrict consideration to the correct display

  c<-which(names(origfile) %in% wanted) #find colnumbers of wanted
  #NB beware that the columns may not be in the same order as in the wanted list
  nufile <- nufile[,c]#select only wanted columns
  nufile <- na.omit(nufile) # remove NAs
  return(nufile)
}
```

The Hoaglin-Iglewicz function is used to remove outliers. We create an 'outlier' column that codes anticipations as 2, and long RT outliers as 1. All else is zero. The original RT is saved as allRT, and the RT column has NA for outliers, so they will be automatically excluded henceforth.

```{r RToutlierfunction}
remove.outliers <- function(origdat,lowcut,zcut){
# Outlier removal put in separate function, which can be applied to different task RTs.
# 
# origdat is the dataframe with data for outlier removal, mycolname is name of variable 
# for outlier removal (usually RT), lowcut is a cutoff value below which data excluded
# (typically v fast responses that are implausible, e.g. 201 ms for RT)
# and zcut is the constant in Hoaglin-Iglewicz method (usually 2.2)

# NB this function returns origdat with outlier rows marked, rather than removed
# This makes it easier to record the N outliers per subject.

# NB FUNCTION ASSUMES THAT THERE IS ARE COLUMNS CALLED subject, Correct and RT IN origfile.
# Outliers are computed by subject - NB contrary to previous version, these are defined
# across all trials, rather than separately for L and R.
origdat$allRT <- origdat$RT #we save a copy of RT, as we will be altering RT so that outliers become NA
origdat$outlier<-0
origdat$outlier[origdat$RT<lowcut] <- 2 #outlier col is coded 2 for anticipations
origdat$RT[origdat$RT<lowcut]<-NA #we exclude anticipations when computing quantiles
#Now extract correct RTs for each subject to compute limits for outliers

  for (i in 1:nsub) { # loop through subjects
  subname <- levels(origdat$subject)[i] # find subject ID
  myrows <- which(origdat$subject==subname) # select rows for this subject
  #NB myrows is NOT a consecutive series, because done in blocks
  firstrow <-min(myrows)
  tmp <- data.frame(origdat[myrows,])

  RTcorr <-tmp$RT[tmp$Correct==1] #correct RTs for this subject as a vector

 # Identify 25th and 75th quartiles, and the difference between them
          lower_quartile <- quantile(RTcorr, probs=0.25, na.rm="TRUE")
          upper_quartile <- quantile(RTcorr, probs=0.75, na.rm="TRUE")
          quartile_diff <- upper_quartile - lower_quartile
        # Outliers are defined as being below or above 2.2 times the quartile difference
          lower_limit <- lower_quartile - zcut*quartile_diff
          upper_limit <- upper_quartile + zcut*quartile_diff
        # create outlier variable
w1 <- which(origdat$RT[myrows]>upper_limit) #rows with outlier, nb RELATIVE to myrows range
origdat$outlier[myrows[w1]]<-1
}
 #For RT we just remove slow outliers, so ignore lower_limit
origdat$RT[origdat$outlier>0]<-NA #RT variable now has outliers excluded as NA
return(origdat)
}
```

This is a generic chunk of code that can be used for Chimeric faces and Rhyme Detection.

```{r maketaskdataframe, warning=FALSE}
make.df <- function(origdat,varlist,latlist){
nsub<-length(unique(origdat$subject))
# create compact data frame 
temp_dat <- data.frame(matrix(ncol = 4, nrow = nsub*2))
colnames(temp_dat) <- varlist
myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  subname <- levels(origdat$subject)[i] # find subject ID
  myrows <- which(origdat$subject==subname) # select rows for this subject
  latcol <- which(colnames(origdat) == varlist[2])
  RTcol <- which(colnames(origdat) == varlist[4])
  tmp <- data.frame(origdat[myrows,])

  for (j in 1:2) {
    myrow <- myrow+1
    w<- which(tmp[,latcol]== latlist[j]) # match on laterality
    tmp1 <- tmp[w,]
    tmp2 <- tmp1[tmp1$Correct == 1,]

    temp_dat$subject[myrow] <- subname # add row for subject
    temp_dat$side[myrow] <- latlist[j] # add row for side
    temp_dat$accurate[myrow] <- sum(tmp1$Correct,na.rm=TRUE) # add row for N accurate response
    temp_dat$p.corr[myrow] <- 100*mean(tmp1$Correct,na.rm=TRUE) # add row for %accurate responses - NB for chimeric only error is wrong emotion - measure is just a measure of choice. For RDT, can make error of selecting wrong side, so % correct is meaningful

    temp_dat$RT[myrow] <- mean(as.numeric(tmp2$RT),na.rm=TRUE) # add row for mean RT
    temp_dat$N[myrow] <- length(tmp1$RT) # add count of rows for this subject
  }
}
return(temp_dat)
}
```

This plots data and conducts t-tests for each task.

```{r plotdata_ttest, warning=FALSE}
dopirate <- function(origdat,task,measure){
# measure for each side
# plot the number of correct responses in each side
pirateplot(formula = X1 ~ side,
           data = origdat,
           main = measure,
           theme = 0,
           pal = "southpark", # southpark color palette
           bean.f.o = .0, # Bean fill
           point.o = .3, # Points
           inf.f.o = .7, # Inference fill
           inf.b.o = .8, # Inference border
           avg.line.o = 1, # Average line
           bar.f.o = .5, # Bar
           point.pch = 21,
           point.cex = .7,
           inf.method= "ci")
# plot L & R to get slope
q <- ggplot(data=origdat, aes(x=side, y=X1, group=subject, color=subject)) +
    geom_line() +
    geom_point() +
    ggtitle(paste0(task,": ",measure," by side"))
q + theme_bw() + theme(legend.position = "none")
#  t test
print(paste0(task,measure))
myt1<-t.test(origdat$X1~ origdat$side, paired = TRUE, alternative = "two.sided")
print(myt1)
}
```

Another generic function that can be used for different tasks to make a laterality index and write it to the allsum file

```{r makeLIs, warning=FALSE}
# This function modifies a copy of the allsum file and returns it

# User selects whether accuracy or RT by specifying col X1 in origdat when calling the function

# Polarity is -1 or 1 and can be set to keep all LIs positive if in a given direction
# Cutoff will create NA if level of performance is outside predetermined limits

makeLI <- function(allsums,origdat,task,mycolname,polarity,mycutoff){
# we do this for accuracy first
# reformat data frame
keeps <- c("subject", "side", "X1") #X1 holds either accuracy or RT
LIdf <- origdat[keeps]
LIdf <- spread(LIdf, side, X1)
# calculate each participants' lat index
LIdf <- 
  LIdf %>% 
  group_by(subject) %>%
  mutate(thisLI= ((Right - Left)/(Right + Left))*100)

LIdf$LI <- LIdf$thisLI*polarity #polarity can be set to reverse sign - eg for RT a large score is bad, whereas for acc it is good, so can have opp polarity

# check if accuracy or RT level is outside acceptable cutoff.
# Do this based on average for left and right
# LI will be set to NA for those outside limits
LIdf$avg <- (LIdf$Left+LIdf$Right)/2
w<-which(LIdf$avg<mycutoff)
if(length(w)>0){
LIdf$LI[w] <- NA}

# now create a data frame for lat index 
LIdf <- select(LIdf, "subject", "LI")
# now merge with allsum
allsums <- merge(allsums, LIdf, by= "subject")
lastcol<-length(colnames(allsums))
colnames(allsums)[lastcol]<-mycolname

return(allsums)
}
```

Make Z score LI.
NB this is NOT a conventional z-score based on whole sample, but rather is an individual-based z-score that indicates whether the person's L-R difference is significantly different from chance.
So you either compare their L-sided RTs and their R-sided RTs, or the proportions correct on L vs proportion correct on R, to see if there is bias away from zero in that individual.
Note that this in effect adjusts for any overall differences in accuracy and RT, as the person is compared with themselves.

```{r makeZ, warning=FALSE}
# This function modifies a copy of the allsum file and returns it

# User selects the measure by specifying col X1 in origdat when calling the function

# Polarity is -1 or 1 and can be set to keep all LIs positive if in a given direction
# Cutoff will create NA if level of performance is outside predetermined limits

makeZ <- function(allsums,origdat,task,mycolname,polarity,mycutoff){
# we do this for accuracy first
# reformat data frame
keeps <- c("subject", "side", "X1") #X1 holds either accuracy or RT
Zdf <- origdat[keeps]
Zdf <- spread(Zdf, side, X1)
# calculate each participants' lat index
Zdf <- 
  Zdf %>% 
  group_by(subject) %>%
  mutate(n= Left + Right, 
         pR= Right/n,
         pL= Left/n,
         Z= (pR-.5)/sqrt(pR*pL/n))
# remove infinite results
is.na(Zdf) <- do.call(cbind,lapply(Zdf, is.infinite))
Zdf <- na.omit(Zdf)

Zdf$Z <- Zdf$Z*polarity #polarity can be set to reverse sign - eg for RT a large score is bad, whereas for acc it is good, so can have opp polarity

# now create a data frame for lat index 
Zdf <- select(Zdf, "subject", "Z")
# now merge with allsum
allsums <- merge(allsums, Zdf, by= "subject")
lastcol<-length(colnames(allsums))
colnames(allsums)[lastcol]<-mycolname

return(allsums)
}
```

## Chimeric Face Task

This part of the R Markdown file prcoesses the "Chimeric Face Task' implmented online via Gorilla.sc which aims to assess the lateralisation of emotion recognition in a face processing task. In the task, participants are shown a chimeric face. Participants then respond by indicating the emotion that they felt was most strong expressed.

This task is near identical to that reported in Karlsson, Johnstone, and Carey (2019). Stimuli were kindly provided by Dr Michael Burt (https://www.dur.ac.uk/psychology/staff/?id=1942) and were symmetrical average images created from four male and four female faces (see Burt & Perrett, 1997; Innes et al., 2016).

```{r read_face_dat, warning=FALSE}
for (d in 1:2){ #we will process day 1 and day 2 together
  origfile <- all_dat1
  if (d==2)
  {origfile <- all_dat2}
#As with earlier reading in, we default to saving CF_dat2 (day2), but at end of the chunk we reassign to CF_day1 in the first run through the d loop
task <- "Chimeric faces"
wanted <- c("subject", "stimuli", "Response", "RT", "l_hemiface", "r_hemiface") # select wanted columns
disp <- "e_Task"
CF_dat2 <- procdata(origfile,wanted,task,disp) #use generic function (see above) to read in relevant columns

# count subjects
nsub <-length(levels(CF_dat2$subject))
print(paste('day',d,task,": subjects: ", nsub))

CF_same2 <- CF_dat2[CF_dat2$l_hemiface == CF_dat2$r_hemiface,] # create data frame for non-chimeras
CF_dat2 <- CF_dat2[CF_dat2$l_hemiface != CF_dat2$r_hemiface,] # remove items where same expression in both hemifaces
CF_dat2 <- remove.outliers(CF_dat2,200,1.65) #use generic outlier removal function
outliertable<-table(CF_dat2$subject,CF_dat2$outlier)
#print(outliertable)

# mark Ps with too many trials indicating issues with the task or cases where less than 80% of data is left
# we mark this to exclude in the next section
out2 <- subset(data.frame(outliertable), Var2== 0)
out2 <- out2[out2$Freq > 96 | out2$Freq < 76,]

if (d==1)
{CF_dat1 <- CF_dat2
 CF_same1 <- CF_same2
 out1 <- out2}
}
# bind one long list of those who have missing trials or too much data
CF_out <- rbind(out1, out2)
```

First, we judge participants' ability to correctly identify emotions using the "CF_same" dataframe. Participants are removed if their performance is below 75% due to poor ability to identify emotions that will inevitably influence performance on the Chimeric Face Task.

```{r emotions, warning=FALSE,message=FALSE}
# this will mark correct responses as 1 or 0. It will be importantfor removing participants later on. 
# NB we loop through same process for day1 and day2
for (d in 1:2){
  CF_same <- CF_same1
  if (d==2){
    CF_same <- CF_same2
  }

CF_same <- 
  CF_same %>% 
  mutate(emotion_recog= ifelse(as.character(Response) == as.character(l_hemiface), 1, 0))
# calculate participants average
emotion_mean <- aggregate(FUN= mean, data= CF_same, emotion_recog~ subject)
emotion_mean$subject <- as.factor(trimws(as.character(emotion_mean$subject)))
# add this to allsum 
allsum <- merge(allsum, emotion_mean, by= "subject")
}
mycol <- length(colnames(allsum))
colnames(allsum)[(mycol-1):mycol]<-c('Emot.recog1','Emot.recog2')

plot(jitter(allsum$Emot.recog1,10),jitter(allsum$Emot.recog2,10),pch=3,cex=.6)
abline(h=.75)
abline(v=.75)
text(.6,.8,'drop if <75% on either run',cex=.8)

#db modified to mark rather than drop cases with scores < .75
allsum$exclude <- 0 #default is to exclude

# mark the cases where there is less than 75% for identical stimuli
w <- union(which(allsum$Emot.recog1 < .75),which(allsum$Emot.recog2 < .75))
allsum$exclude[w]<-1 #code exclude = 1 if excluded because poor emot recog
# now mark those with too many or too few
allsum$exclude[allsum$subject %in% CF_out$Var1] <- 100

```

In this section, the correct sequences are marked as 1, incorrect asnwers are marked as 0. We then mark whether the correct answers were in the left or right hemiface.

```{r mark_face_correct_face, warning=FALSE}
# this will mark correct responses as 1 or 0.
# NB we again loop through same process for day1 and day2
for (d in 1:2){
  CF_dat <- CF_dat1
  if (d==2){
    CF_dat <- CF_dat2
  }
CF_dat <- 
  CF_dat %>% 
  mutate(Correct= ifelse(as.character(Response) == as.character(l_hemiface), 1, 
                         ifelse(as.character(Response) == as.character(r_hemiface), 1, 0)))
# this will mark the side responded to. NB to allow for generic functions for all tasks, we use the label 'side' rather than hemiface
CF_dat <- 
  CF_dat %>% 
  mutate(side= ifelse(as.character(Response) == as.character(l_hemiface) & Correct == 1, "Left",
                       ifelse(as.character(Response) == as.character(r_hemiface) & Correct == 1, "Right", NA)))
# code side as factor
CF_dat$side <- as.factor(CF_dat$side)

  if(d==1){
    CF_dat1 <- CF_dat}
  if (d==2){
    CF_dat2 <- CF_dat
  }
}
```

Now the data is ready, create an empty data frame and populate this. It includes columns for the participant identifier, side, and count of accurate responses for each side.

```{r makedataframe, warning=F}
#We create chim_dat files for day1 and day 2
varlist <-   c("subject", "side", "accurate", "RT") #need to be in this order, ie sub, side, acc and RR
origdat <- CF_dat1
latlist <- c("Left","Right") #names of factor levels

chim_dat1 <- make.df(CF_dat1,varlist,latlist)
chim_dat2 <- make.df(CF_dat2,varlist,latlist)

w<-which(is.na(chim_dat1$subject))
if (length(w)>0) {chim_dat1<-chim_dat1[-w,]}
w<-which(is.na(chim_dat2$subject))
if (length(w)>0) {chim_dat2<-chim_dat2[-w,]}

chim_dat1[is.na(chim_dat1)] <- 0
chim_dat2[is.na(chim_dat2)] <- 0
```

The data is visualised using the pirate plot function and t.test conducted.

```{r dosummary}
origdat <- chim_dat1
measure <- 'Accuracy'
origdat$X1 <- chim_dat1$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Chimeric faces day 1'
dopirate(origdat,task,measure)
origdat$X1 <- chim_dat1$RT
measure <- 'RT'
dopirate(origdat,task,measure)
t.test(origdat$accurate~ origdat$side)
#Now repeat for day 2
origdat <- chim_dat2
origdat$X1 <- chim_dat2$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Chimeric faces day 2'
measure <- 'Accuracy'
dopirate(origdat,task,measure)
origdat$X1 <- chim_dat2$RT
measure <- 'RT'
dopirate(origdat,task,measure)
```

Finally, a laterality index is calculated and plotted based on the equation [(R-L)/(R+L)] x 100, where L= left side and R= right side.

If the index is negative, emotion recognition is lateralised in the left hemisphere. If the index is positive, emotion recognition is lateralised in the right hemisphere

Here, a .csv is written for the chimeric face LI.

```{r domakeLI}
#We add Chimeric faces LIs to allsum for day 1 and then day2
task<-'Chimeric faces'
origdat<-chim_dat1
origdat$X1 <- origdat$accurate
mycolname<-"Day1_CF_acc_LI"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

# now do RT
origdat$X1 <- origdat$RT
mycolname<-"Day1_CF_RT_LI"
polarity <-1

#if we don't specify mycutoff, it will remain at level set for accuracy, which means it should not trap any RT values as they will all be much bigger
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

origdat<-chim_dat2
origdat$X1 <- origdat$accurate
mycolname<-"Day2_CF_acc_LI"
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
origdat$X1 <- origdat$RT
mycolname<-"Day2_CF_RT_LI"
polarity <-1
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

cor1 <- cor(allsum$Day1_CF_acc_LI,allsum$Day2_CF_acc_LI, use = "complete.obs")
plot(allsum$Day1_CF_acc_LI,allsum$Day2_CF_acc_LI,main='Chimeric LI accuracy',col=(1+allsum$exclude))
abline(v=0)
abline(h=0)
text(-50,75,paste0('r = ',round(cor1,3)))
text(-50,65,'red will be excluded',cex=.85)
cor2<-cor(allsum$Day1_CF_RT_LI,allsum$Day2_CF_RT_LI, use = "complete.obs")
plot(allsum$Day1_CF_RT_LI,allsum$Day2_CF_RT_LI,main='Chimeric LI RT',col=(1+allsum$exclude))
abline(v=0)
abline(h=0)
text(-10,22,paste0('r = ',round(cor2,3)))
text(-10,18,'red will be excluded',cex=.85)
```

Here we use the Z score version. This computes for each subject whether they have a different proportion correct for L and R

```{r domakeZ}
#We add Chimeric faces LIs to allsum for day 1 and then day2
task<-'Chimeric faces'
origdat<-chim_dat1
origdat$X1 <- origdat$N
mycolname<-"Day1_CF_acc_Z"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)

origdat<-chim_dat2
origdat$X1 <- origdat$N
mycolname<-"Day2_CF_acc_Z"
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)
w<-which(allsum$excluded == 1)
allsum$Day1_CF_acc_Z[w] <- NA #removes zscores for excluded cases
allsum$Day2_CF_acc_Z[w] <- NA

RR<-intersect(which(allsum$Day1_CF_acc_Z>1.96),which(allsum$Day2_CF_acc_Z>1.96))
LL<-intersect(which(allsum$Day1_CF_acc_Z< -1.96),which(allsum$Day2_CF_acc_Z< -1.96))

cor1 <- cor(allsum$Day1_CF_acc_Z,allsum$Day2_CF_acc_Z,use='complete.obs',method='spearman')
plot(allsum$Day1_CF_acc_Z,allsum$Day2_CF_acc_Z,main='Chimeric Z accuracy')
abline(v=0)
abline(h=0)
abline(v = 1.96, col='red')
abline(v = -1.96, col='red')
abline(h = 1.96, col='red')
abline(h = -1.96, col='red')
text(-7,7.5,paste0('Spearman r = ',round(cor1,3)),cex=.8)
text(-2,-15,paste('Those in bottom quadrant\n bounded by red are consistently\n L-lateralised, N = ',length(LL)),cex=.8)
text(5,18,paste('Those in top quadrant\n bounded by red are consistently\n R-lateralised, N = ',length(RR)),cex=.8)
#text(8,-8,'One outlier excluded',cex=.8)
#Removal of the outlier does not have much influence on rs
```

As we can see, we generally get a left side/right hemisphere advantage in our accuracy data. We would not expect a RT advantage for this type of measure, where the response is to a single stimulus with conflicting information from the 2 sides.

## Bergen Dichotic listening

This part of the R Markdown file prcoesses the "Dichotic listening task' implmented online via Gorilla.sc which aims to assess the lateralisation of receptive language lateralisation. In the task, participants hear a different CV sound in each ear. Participants respond to the sound they heard clearest

This task is near identical to that reported in Karlsson, Johnstone, and Carey (2019) and were original described in Hugdahl, Westerhausen, Alho, Medvedev, Laine, & Hämäläinen (2009). Stimuli were kindly provided by Dr Emma Karlsson (https://research.bangor.ac.uk/portal/en/researchers/emma-karlsson(17f6f06c-8aa9-4d16-93ba-4f98f285297e).html) and permission to use these stimuli was given by Professor Kenneth Hughdahl (https://www.uib.no/en/persons/Kenneth.Hugdahl).

```{r read_dichotic_dat, warning=FALSE}
for (d in 1:2){ #we will process day 1 and day 2 together
  origfile <- all_dat1
  if (d==2)
  {origfile <- all_dat2}
# lowercase for responses
origfile$left_channel <- tolower(origfile$left_channel)
origfile$right_channel <- tolower(origfile$right_channel)
# make meaningful stimuli name
origfile$stimuli <- paste0(origfile$left_channel, "-", origfile$right_channel)
#As with earlier reading in, we default to saving dichotic_dat2 (day2), but at end of the chunk we reassign to CF_day1 in the first run through the d loop
task <- "Dichotic Listening"
wanted <- c("subject", "stimuli", "Response", "RT", "left_channel", "right_channel") # select wanted columns
disp <- "e_task"
dichotic_dat2 <- procdata(origfile,wanted,task,disp) #use generic function (see above) to read in relevant columns

# count subjects
nsub <-length(levels(dichotic_dat2$subject))
print(paste('day',d,task,": subjects: ", nsub))

dichotic_same2 <- dichotic_dat2[dichotic_dat2$left_channel == dichotic_dat2$right_channel,] # create data frame for non-chimeras
dichotic_dat2 <- dichotic_dat2[dichotic_dat2$left_channel != dichotic_dat2$right_channel,] # remove items where same expression in both hemifaces
dichotic_dat2 <- remove.outliers(dichotic_dat2,200,1.65) #use generic outlier removal function
outliertable<-table(dichotic_dat2$subject,dichotic_dat2$outlier)
#print(outliertable)

# remove those with too many trials indicating issues with the task
# we mark this to exclude in the next section
out2 <- subset(data.frame(outliertable), Var2== 0)
out2 <- out2[out2$Freq > 120 | out2$Freq < 96,]

if (d==1)
{dichotic_dat1 <- dichotic_dat2
 dichotic_same1 <- dichotic_same2
 out1 <- out2}
}
# bind one long list of those who have missing trials or too much data
DL_out <- rbind(out1, out2)
```

First, we judge participants' ability to correctly identify sounds using the "dichotic_same" dataframe. Participants are marked as excluded if their performance is below 75%.

```{r soundsability, warning=FALSE,message=FALSE}
# this will mark correct responses as 1 or 0. It will be importantfor removing participants later on. 
# NB we loop through same process for day1 and day2
for (d in 1:2){
  dichotic_same <- dichotic_same1
  if (d==2){
    dichotic_same <- dichotic_same2
  }

dichotic_same <- 
  dichotic_same %>% 
  mutate(sound_recog= ifelse(as.character(Response) == as.character(left_channel), 1, 0))
# calculate participants average
sound_mean <- aggregate(FUN= mean, data= dichotic_same, sound_recog~ subject)
sound_mean$subject <- as.factor(trimws(as.character(sound_mean$subject)))
# add this to allsum 
allsum <- merge(allsum, sound_mean, by= "subject")
}
mycol <- length(colnames(allsum))
colnames(allsum)[(mycol-1):mycol]<-c('sound.recog1','sound.recog2')

# once again, mark rather than remove those with poor performance on identical stimuli
w<-union(which(allsum$sound.recog1 < .75), which(allsum$sound.recog2 < .75))
ww<-intersect(which(allsum$exclude==1),w)
#ww contains people who were excluded on both dichotic and faces. 
# code exclude = 2 for the cases where there less than 75% for identical stimuli
allsum$exclude[w]<-2 #code exclude = 2 if excluded because poor dichotic
allsum$exclude[ww]<-12 #code for those excluded on both chimeric and dichotic
# remove those with too many or missing trials
allsum$exclude[allsum$subject %in% DL_out$Var1] <- 100
# table to summarise
t <- table(allsum$exclude) 
rownames(t)<-c('include','exclude CF','exclude dich','exclude CF+dich','trials')
t
```

In this section, the correct sequences are marked as 1, incorrect asnwers are marked as 0. We then mark whether the correct answers were in the left or right ear.

```{r mark_dich_correct_ear, warning=FALSE}
# this will mark correct responses as 1 or 0.
# NB we again loop through same process for day1 and day2
for (d in 1:2){
  dichotic_dat <- dichotic_dat1
  if (d==2){
    dichotic_dat <- dichotic_dat2
  }
dichotic_dat <- 
  dichotic_dat %>% 
  mutate(Correct= ifelse(as.character(Response) == as.character(left_channel), 1, 
                         ifelse(as.character(Response) == as.character(right_channel), 1, 0)))
# this will mark the side responded to. NB to allow for generic functions for all tasks, we use the label 'side' rather than hemiface
dichotic_dat <- 
  dichotic_dat %>% 
  mutate(side= ifelse(as.character(Response) == as.character(left_channel) & Correct == 1, "Left",
                       ifelse(as.character(Response) == as.character(right_channel) & Correct == 1, "Right", NA)))
# code side as factor
dichotic_dat$side <- as.factor(dichotic_dat$side)

  if(d==1){
    dichotic_dat1 <- dichotic_dat}
  if (d==2){
    dichotic_dat2 <- dichotic_dat
  }
}
```

Now the data is ready, create an empty data frame and populate this. It includes columns for the participant identifier, side, and count of accurate responses for each side.

```{r makedataframedich, warning=F}
#We create DL_dat files for day1 and day 2
varlist <-   c("subject", "side", "accurate", "RT") #need to be in this order, ie sub, side, acc and RR
origdat <- dichotic_dat1
latlist <- c("Left","Right") #names of factor levels

DL_dat1 <- make.df(dichotic_dat1,varlist,latlist)
DL_dat2 <- make.df(dichotic_dat2,varlist,latlist)

w<-which(is.na(DL_dat1$subject))
if (length(w)>0) {DL_dat1<-DL_dat1[-w,]}
w<-which(is.na(DL_dat2$subject))
if (length(w)>0) {DL_dat2<-DL_dat2[-w,]}

DL_dat1[is.na(DL_dat1)] <- 0
DL_dat2[is.na(DL_dat2)] <- 0
```

The data is visualised using the pirate plot function and t.test conducted.

```{r dosummarydich}
origdat <- DL_dat1
measure <- 'Accuracy'
origdat$X1 <- DL_dat1$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Dichotic listening day 1'
dopirate(origdat,task,measure)
origdat$X1 <- DL_dat1$RT
measure <- 'RT'
dopirate(origdat,task,measure)
t.test(origdat$accurate~ origdat$side)
#Now repeat for day 2
origdat <- DL_dat2
origdat$X1 <- DL_dat2$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Dichotic listening day 2'
measure <- 'Accuracy'
dopirate(origdat,task,measure)
origdat$X1 <- DL_dat2$RT
measure <- 'RT'
dopirate(origdat,task,measure)
```

Finally, a laterality index is calculated and plotted based on the equation [(R-L)/(R+L)] x 100, where L= left side and R= right side.

If the index is negative, emotion recognition is lateralised in the left hemisphere. If the index is positive, emotion recognition is lateralised in the right hemisphere

Here, a .csv is written for the chimeric face LI.

```{r domakeLIdich}
#We add Dichotic LIs to allsum for day 1 and then day2
task<-'Dichotic listening'
origdat<-DL_dat1
origdat$X1 <- origdat$accurate
mycolname<-"Day1_DL_acc_LI"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

# now do RT
origdat$X1 <- origdat$RT
mycolname<-"Day1_DL_RT_LI"
polarity <-1

#if we don't specify mycutoff, it will remain at level set for accuracy, which means it should not trap any RT values as they will all be much bigger
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

origdat<-DL_dat2
origdat$X1 <- origdat$accurate
mycolname<-"Day2_DL_acc_LI"
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
origdat$X1 <- origdat$RT
mycolname<-"Day2_DL_RT_LI"
polarity <-1
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

cor1 <- cor(allsum$Day1_DL_acc_LI,allsum$Day2_DL_acc_LI, use = "complete.obs", method = "pearson")
plot(allsum$Day1_DL_acc_LI,allsum$Day2_DL_acc_LI,main='Dichotic LI accuracy',col=(1+allsum$exclude))
abline(v=0)
abline(h=0)
text(-75,75,paste0('r = ',round(cor1,3)))
cor2<-cor(allsum$Day1_DL_RT_LI,allsum$Day2_DL_RT_LI, use = "complete.obs")
plot(allsum$Day1_DL_RT_LI,allsum$Day2_DL_RT_LI,main='Dichotic LI RT',col=(1+allsum$exclude))
abline(v=0)
abline(h=0)
text(-5,15,paste0('r = ',round(cor2,3)))
```

Here we use the Z score version. This is an individualised score that directly indicates whether or not the person is reliably lateralised on one test occasion.

```{r domakeZ_dich}
#We add dichotic LIs to allsum for day 1 and then day2
task<-'Dichotic listening'
origdat<-DL_dat1
origdat$X1 <- origdat$N
mycolname<-"Day1_Dich_acc_Z"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.AP: chekced against pre-reg
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)

origdat<-DL_dat2
origdat$X1 <- origdat$N
mycolname<-"Day2_Dich_acc_Z"
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)

#convert excluded cases with code 2 or 12 to NA
w<-which(allsum$exclude>1)
allsum$Day1_Dich_acc_Z[w]<-NA
allsum$Day2_Dich_acc_Z[w]<-NA

RR<-intersect(which(allsum$Day1_Dich_acc_Z>1.96),which(allsum$Day2_Dich_acc_Z>1.96))
LL<-intersect(which(allsum$Day1_Dich_acc_Z< -1.96),which(allsum$Day2_Dich_acc_Z< -1.96))

cor1 <- cor(allsum$Day1_Dich_acc_Z,allsum$Day2_Dich_acc_Z,use='complete.obs',method='spearman')

plot(allsum$Day1_Dich_acc_Z,allsum$Day2_Dich_acc_Z,main='Dich Z accuracy',pch=allsum$handpch)
abline(v=0)
abline(h=0)
text(-7.5,15,paste0('Spearman r = ',round(cor1,3)),cex=.8)
abline(v = 1.96, col='red')
abline(v = -1.96, col='red')
abline(h = 1.96, col='red')
abline(h = -1.96, col='red')

text(20,-15,paste('Those in bottom quadrant\n bounded by red are consistently\n L-lateralised, N = ',length(LL)),cex=.8)
text(5,40,paste('Those in top quadrant\n bounded by red are consistently\n R-lateralised, N = ',length(RR)),cex=.8)

```

## Rhyme Detection Task

This portion of the R Markdown file prcoesses the "Rhyme Detection Task' implmented online via Gorilla.sc which aims to assess the lateralisation of language processing. In the task, participants are shown a word in the central visual field and different images in the left and right visual field. An arrow appears pointing to either image. Participants then respond by indicating whether the image and word rhyme.

This reads in a csv file where each row represents a response to a trial. Participants, identified by random letter strings, are stacked on top of each other. We start by reading the data and cutting out unwanted information.

```{r read_dat_RDT, warning=FALSE}
for (d in 1:2){ #same script for days 1 and 2

origfile <- all_dat1
if(d==2)
  {origfile <- all_dat2}
task <- "Rhyme Detection Task"
disp <- "Task"

# nb when creating new file, we default to name 'RDT_dat2', but this is reassigned to 'RDT_dat1' at end of processing for day 1.

# new data
wanted <- c("Trial","subject", "Correct", "RT", "ANSWER","Response") # select wanted columns
RDT_dat2 <- procdata(origfile,wanted,task,disp) #Here we reuse the procdata function that we developed for processing Chimeric faces data - just does some generic processing

# do some wrangling on the ANSWER variable to make it compatible with functions
RDT_dat2<-filter(RDT_dat2,ANSWER %in% c('left','right'))
c <- which(colnames(RDT_dat2)=="ANSWER")
colnames(RDT_dat2)[c]<-'side'
RDT_dat2$side<-as.factor(RDT_dat2$side)
  levels(RDT_dat2$side)<-c("Left","Right")
RDT_dat2<-remove.outliers(RDT_dat2,200,1.65) #numbers specify min RT and zscore for Hoaglin-Iglewicz respectively

# count subjects
nsub <-length(levels(RDT_dat2$subject))
print(paste0("Day 1 subjects: ", nsub))
outliertable<-table(RDT_dat2$subject,RDT_dat2$outlier)
#print(outliertable)

# remove those with too many trials indicating issues with the task
out2 <- subset(data.frame(outliertable), Var2== 0)
out2 <- out2[out2$Freq > 260 | out2$Freq < 208,]

if(d==1)
{RDT_dat1 <- RDT_dat2
 out1 <- out2}
}
# bind one long list of those who have missing trials or too much data
RDT_out <- rbind(out1, out2)

# once again, mark rather than remove those with poor performance on identical stimuli
allsum$exclude[allsum$subject %in% RDT_out$Var1] <- 100
# table to summarise
t <- table(allsum$exclude) 
rownames(t)<-c('include','exclude CF','exclude dich','exclude CF+dich','trials')
t
```

Now the data is ready, create an empty data frame and populate this. It includes columns for the participant identifier, rhyme, acuracy, RT (RT only for correct answers).

```{r makedataframeRDT, warning=F}
#We create files for day1 and day 2
# NB warnings created because of different Ns for time 1 and time 2, which means there are blanks in subjects
# We deal with this below by removing rows with no subject code
varlist <-   c("subject", "side", "accurate", "RT") #need to be in this order, ie sub, side, acc and RR - we will use these in the make.df function
origdat <- RDT_dat1
latlist <- c("Left","Right") #names of factor levels
RDT_df1 <- make.df(origdat,varlist,latlist)
origdat <- RDT_dat2
RDT_df2 <- make.df(origdat,varlist,latlist)

# Remove blank rows 
w<-which(is.na(RDT_df1$subject))
if (length(w)>0) {RDT_df1<-RDT_df1[-w,]}
w<-which(is.na(RDT_df2$subject))
if (length(w)>0) {RDT_df2<-RDT_df2[-w,]}

# Exclude cases with low % correct
# NB something weird with these - seem to be ones with too many trials, suggesting some problem with multiple responses
w<-which(RDT_df1$p.corr<50)
if (length(w)>0)
{RDT_df1$accurate[w]<-NA
RDT_df1$RT[w]<-NA
}
w<-which(RDT_df2$p.corr<50)
if (length(w)>=0)
{RDT_df2$accurate[w]<-NA
RDT_df2$RT[w]<-NA
}

# now remove the cases where there are NA for only one visual field
# this will remove those who responded by pressing only one button.
RDT_df1 <- na.omit(RDT_df1) %>% 
  group_by(subject) %>%
  mutate(no_rows = length(subject))
RDT_df1 <- RDT_df1[RDT_df1$no_rows == 2,]

RDT_df2 <- na.omit(RDT_df2) %>% 
  group_by(subject) %>%
  mutate(no_rows = length(subject))
RDT_df2 <- RDT_df2[RDT_df2$no_rows == 2,]
```

```{r checker, warning=FALSE}
for (d in 1:2){
origdat <-RDT_df1
if (d==2)
{origdat <- RDT_df2}
measure <- 'Accuracy'
origdat$X1 <- origdat$accurate
task <- 'Rhyme Detection'

dopirate(origdat,task,measure)
origdat$X1 <- origdat$RT
measure <- 'RT'
dopirate(origdat,task,measure)

#now create LI
origdat$X1 <- origdat$p.corr
mycolname<-"RDT_p.corr_LI"
polarity<- 1 #so LI is positive in predicted direction
mycutoff <- 10 
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
origdat$X1 <- origdat$RT
mycolname<-"RDT_RT_LI"
polarity <- (-1)
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

#Cutoff has removed those with low accuracy, but we should also remove their RT LIs
#We do this by finding those with NA in accuracy LI
w<-which(is.na(allsum$RDT_acc_LI))
if(length(w)>0)
{allsum$RDT_RT_LI[w]<-NA}
#t.test to test if laterality index differs from zero
t.test(allsum$RDT_p.corr_LI)
t.test(allsum$RDT_RT_LI)
#rename column to identify day
myncol <- length(colnames(allsum))
colnames(allsum)[(myncol-1):myncol] <- paste0('Day',d,'.',colnames(allsum)[(myncol-1):myncol])

}
cor1<-cor(allsum$Day1.RDT_p.corr_LI,allsum$Day2.RDT_p.corr_LI,use='complete.obs')
cor2<-cor(allsum$Day1.RDT_RT_LI,allsum$Day2.RDT_RT_LI,use='complete.obs')
plot(allsum$Day1.RDT_p.corr_LI,allsum$Day2.RDT_p.corr_LI,main='Rhyme detect: LI for accuracy (%correct)')
abline(v=0)
abline(h=0)
text(-3,3,paste0('r = ',round(cor1,3)))
plot(allsum$Day1.RDT_RT_LI,allsum$Day2.RDT_RT_LI,main='Rhyme detect: LI for RT')
abline(v=0)
abline(h=0)
text(-3,3,paste0('r = ',round(cor2,3)))
```

### T-test as alternative to LI for RDT

```{r do.ttest}
for (d in 1:2){
  myfile<-RDT_dat1
  if(d==2){
    myfile<-RDT_dat2}

  nsubs <- nrow(allsum)
  allsum$tlat_RDT<-NA
   allsum$plat_RDT<-NA
  for (i in 1:nsubs) {
    w<-which(myfile$subject==allsum$subject[i])
    if(length(w)>0)
    {
      myt<-t.test(myfile$RT[w]~myfile$side[w])
      allsum$tlat_RDT[i]<-myt$statistic
      allsum$plat_RDT[i]<-myt$p.value
    }
  }
   if (d==1){
     maxcol <-length(colnames(allsum))
     colnames(allsum)[(maxcol-1):maxcol]<-c('tlat_RDT_1','plat_RDT_1')
   }
   if (d==2){
     maxcol <-length(colnames(allsum))
     colnames(allsum)[(maxcol-1):maxcol]<-c('tlat_RDT_2','plat_RDT_2')
   }
}
colcount<-length(colnames(allsum))
# Remove values for cases that are excluded for low acc: do this for both times by just blitzing last 4 columns of allsum
w<-which(is.na(allsum$Day1.RDT_RT_LI))
allsum[w,(colcount-3):colcount]<-NA
w<-which(is.na(allsum$Day2.RDT_RT_LI))
allsum[w,(colcount-3):colcount]<-NA

# compare t-value with LI : use t = 1.65 as cutoff for sig lateralised
 plot(allsum$tlat_RDT_1,allsum$Day1.RDT_RT_LI)
 abline(v=1.65)
  plot(allsum$tlat_RDT_2,allsum$Day2.RDT_RT_LI)
   abline(v=1.65)
# compare t-value from day 1 and day 2 
     myr <- cor(allsum$tlat_RDT_1,allsum$tlat_RDT_2,use='complete.obs')
    plot(allsum$tlat_RDT_1,allsum$tlat_RDT_2)
    abline(v=0)
    abline(h=0)
   abline(v=1.65) 
   abline(h=1.65)
      abline(v=-1.65) 
   abline(h=-1.65)
   text(-3,4,paste0('r = ',round(myr,3)))
```

## Finger Tapping Task

This portion of the R Markdown file prcoesses the "Finger Tapping Task' implmented in Gorilla.sc which aims to quantify handedness. In the task, participants are required to press a sequence of buttons twice for each hand (Left: W, R, V, X; Right: T, U, M, B) as many times as possible within a 30 ms. There is a version that requires more precise movements.

A score of 1 is given each time that a participants completes the sequence.

This reads in a csv file where each row represents a button press. Participants, identified by random letter strings, are stacked on top of each other. We start by reading the data and cutting out unwanted information.

NOTE. The create data frame function disrupts order here.

In this section, the correct sequences are marked (e.g. pressing Y, U, J, H, in order will correspond to a score of 1).

```{r readdataFinger, warning=FALSE}

finger <- read.csv("data/combined/day1_combined.csv", na.strings = "NA")
# cleaning
finger <- finger[finger$Attempt >= 1,] # only answers 
finger <- finger[finger$display == "right hand" | finger$display == "left hand",] # only for task
# write as factors
finger$display <- as.factor(finger$display) 
finger$subject <- as.factor(finger$Participant.Public.ID)
finger$RT <- as.numeric(as.character(finger$Reaction.Time))

# new data
wanted <- c("subject", "display", "Response", "RT") # select wanted columns
c<-which(names(finger) %in% wanted) #find colnumbers of unwanted
finger <-finger[,c] #remove unwanted columns
finger <- na.omit(finger) # remove NAs

all_finger1<- finger

finger <- read.csv("data/combined/day2_combined.csv", na.strings = "NA")
# cleaning
finger <- finger[finger$Attempt >= 1,] # only answers 
finger <- finger[finger$display == "right hand" | finger$display == "left hand",] # only for task
# write as factors
finger$display <- as.factor(finger$display) 
finger$subject <- as.factor(finger$Participant.Public.ID)
finger$RT <- as.numeric(as.character(finger$Reaction.Time))

# new data
wanted <- c("subject", "display", "Response", "RT") # select wanted columns
c<-which(names(finger) %in% wanted) #find colnumbers of unwanted
finger <-finger[,c] #remove unwanted columns
finger <- na.omit(finger) # remove NAs

all_finger2<- finger

```

```{r readdataFinger2, warning=FALSE}
for (d in 1:2){ #we will process day 1 and day 2 together
  origfile <- all_finger1
  if (d==2)
  {origfile <- all_finger2}
#As with earlier reading in, we default to saving finger2 (day2), but at end of the chunk we reassign to finger1 in the first run through the d loop
finger2 <- origfile

finger2$RTbase[2:nrow(finger2)]<-finger2$RT[1:(nrow(finger2)-1)]
finger2$RTa <- finger2$RT-finger2$RTbase
#remove negative or v long RTs - can indicate when block changes etc
w<-c(which(finger2$RTa<0),which(finger2$RTa>1000))
finger2$RTa[w]<-NA
if (d==1)
{finger1 <- finger2
 }
}
```

In this section, the correct sequences are marked (e.g. pressing Y, M, I, B, in order will correspond to a score of 1).

Added exploratory analysis of SDs of RTs for keypress rather than means, but this was not particularly illuminating.

```{r scorefinger}
# We can go directly to creation of a data.frame by subjects, and use string matching to identify N correct seqs

myseq <-c('tumb','wrvx') # R, # L

subids <- levels(finger1$subject) #just look at data for those doing time1 and
nsub <- length(subids)

finger.df <- data.frame(matrix(NA,nrow=nsub,ncol=9))
colnames(finger.df)<-c('subject', 'R1', 'R2', 'L1', 'L2', 'R1sd', 'R2sd', 'L1sd', 'L2sd')
for (n in 1:nsub){
  finger.df$subject[n] <- subids[n]
  myc <- 1 #will be incremented so sums written to correct columns
  for (d in 1:2){ #each day data
    myfinger <- finger1
    if(d==2){
      myfinger<- finger2}
      tempfinger <- filter(myfinger,subject==subids[n],Response %in% c('t', 'u', 'm', 'b', 'w', 'r', 'v', 'x')) 
      #exclude rows with other stuff in Response
      #(this is not essential as nonsequences will just get ignored)

      wholeseq <-paste0(tempfinger$Response,collapse='') #make a long string of responses
      for (s in 1:2){
        myc<-myc+1 #increment column counter
        target <- myseq[s]
        finger.df[n,myc]<-str_count(wholeseq,target) 
        #count how many target sequences are in the long string that contains all responses

      }
      righth<-filter(tempfinger,display=='right hand')
      lefth<-filter(tempfinger,display=='left hand')
      startcol<-6
      if(d==2){startcol<-8}
      finger.df[n,startcol]<-sd(righth$RTa,na.rm=T)
      finger.df[n,(startcol+1)]<-sd(lefth$RTa,na.rm=T)
    }
}
# Substitute NA for any totals less than 5 - suggest using wrong keys perhaps

for (c in 2:5){
  w<-which(finger.df[,c]<5)
  if (length(w)>0){
     finger.df[w,2:9]<-NA

  }

}
#Need same data in long form for pirate plot, so reorganise
finger.left <- finger.df[,c(1,4,5,8,9)]
finger.right <- finger.df[,c(1,2,3,6,7)]
finger.left$side<-'Left'
finger.right$side <-'Right'
colnames(finger.right)<-c('subject', '1', '2', '1sd', '2sd','side')
colnames(finger.left)<-colnames(finger.right)
finger.long <-rbind(finger.left,finger.right)

```

```{r dosummary2, warning = F}
origdat <- finger.long
measure <- '1'
origdat$X1 <- origdat$`1` #we used a dummy column name X1, so we can vary the data that are used in the pirate plot etc
task <- 'day 1'
dopirate(origdat,task,measure)

origdat <- finger.long
measure <- '2'
origdat$X1 <- origdat$`2` #we used a dummy column name X1, so we can vary the data that are used in the pirate plot etc
task <- 'day 2'
dopirate(origdat,task,measure)

origdat <- finger.long
measure <- '1sd'
origdat$X1 <- origdat$`1sd` #we used a dummy column name X1, so we can vary the data that are used in the pirate plot etc
task <- 'day 1sd'
dopirate(origdat,task,measure)

origdat <- finger.long
measure <- '2sd'
origdat$X1 <- origdat$`2sd` #we used a dummy column name X1, so we can vary the data that are used in the pirate plot etc
task <- 'day 2sd'
dopirate(origdat,task,measure)
```

Finally, a handedness index is calculated and plotted. The handedness index is based on the equation [(RH-LH)/(RH+LH)] x 100, where LH= left hand and RH= right hand.

If the index is negative, the participant is categorised as left handed. If the index is positive, the participant is categorised as right handed.

In the next chunk, the LI is also added to allsum

```{r fingerLI}
task<-'Finger'
origdat<-finger.long
origdat$X1 <- origdat$`1`
mycolname<-"Day1_finger_LI"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (1)#so LI is positive in predicted direction
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

origdat$X1 <- origdat$`2`
mycolname<-"Day2_finger_LI"
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

cor1 <- cor(allsum$Day1_finger_LI,allsum$Day2_finger_LI,use='complete.obs')
plot(allsum$Day1_finger_LI,allsum$Day2_finger_LI,main='Finger Tapping LI')
abline(v=0)
abline(h=0)
text(-20,10,paste0('r = ',round(cor1,3)))
```

Here we use the Z score version.

```{r domakeZ_finger}
#We add Chimeric faces LIs to allsum for day 1 and then day2
task<-'Finger'
origdat<-finger.long
origdat$X1 <- origdat$`1`
mycolname<-"Day1_Finger_acc_Z"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (1)#so LI is positive in predicted direction
allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)

origdat<-finger.long
origdat$X1 <- origdat$`2`
mycolname<-"Day2_Finger_acc_Z"
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)

cor1 <- cor(allsum$Day1_Finger_acc_Z,allsum$Day2_Finger_acc_Z)
plot(allsum$Day1_Finger_acc_Z,allsum$Day2_Finger_acc_Z,main='Finger Z accuracy')
abline(v=0)
abline(h=0)
text(-2.5,2.5,paste0('r = ',round(cor1,3)))

write.csv(allsum, "LIs.csv")
```

## Final sample summary

Here we report the descriptives for the participants who had full datasets.

```{r Finalallsum, warning=FALSE}
# comment out to get participants with full data set
allsum <- allsum[allsum$exclude == 0,]

# first, describe age for the whole sample and plot
allsum %>%
    summarize(avg = mean(age), n = n(), sd = sd(age))
# second, show gender count for the whole population
allsum %>%
  group_by(gender) %>%
    summarize(n = n())
# third, show footedness
allsum %>%
  group_by(footedness) %>%
    summarize(n = n())
# fourth, show handedness
allsum %>%
  group_by(handedness) %>%
    summarize(n = n())
# fifth, show bilingualism
allsum %>%
  group_by(bilingual) %>%
    summarize(n = n())
# sixth, show miles
allsum %>%
  group_by(miles) %>%
    summarize(n = n())
# seventh, show porta
allsum %>%
  group_by(porta) %>%
    summarize(n = n())
# eigth, show LexTALE
allsum %>%
    summarize(avg = mean(lexTALE), n = n(), sd = sd(lexTALE))

# Now that we have the basic scores, let's look at how this looks with regards to handedness
# first, describe age for the whole sample and plot
ggplot(allsum, aes(x = handedness, y = age)) +
  geom_pirate(aes(colour = handedness), bars = FALSE) + theme_classic() + 
  ylim(10, 60) + ggtitle("Age by handedness") + ylab("Age (years)") + xlab("")
# second, show gender count for the whole population
a <- allsum %>%
  group_by(handedness) %>%
    summarize(n = n())
ggplot(data=a, aes(x=handedness, y=n, fill= handedness)) +
  geom_bar(stat="identity", alpha= .5, colour="black")+
  geom_text(aes(label=n), vjust=1.6, color="white", size=3.5)+
  theme_minimal() + ggtitle("Handedness") + 
  ylab("Count") + xlab("")
# third, show footedness
a <- allsum %>%
  group_by(handedness, footedness) %>%
    summarize(n = n())
ggplot(data=a, aes(x=handedness, y=n, fill= handedness)) +
  geom_bar(stat="identity")+
  geom_text(aes(label=n), vjust=1.6, color="white", size=3.5)+
  theme_classic() + facet_wrap(.~ footedness) + ggtitle("Footedness by handedness") + 
  ylab("Count") + xlab("")
# fourth, show bilingualism
a <- allsum %>%
  group_by(bilingual, handedness) %>%
    summarize(n = n())
ggplot(data=a, aes(x=handedness, y=n, fill= handedness)) +
  geom_bar(stat="identity")+
  geom_text(aes(label=n), vjust=1.6, color="white", size=3.5)+
  theme_classic() + facet_wrap(.~ bilingual) + ggtitle("Bilingualism by handedness") + 
  ylab("Count") + xlab("")
# fifth, show miles
a <- allsum %>%
  group_by(miles, handedness) %>%
    summarize(n = n())
ggplot(data=a, aes(x=handedness, y=n, fill= handedness)) +
  geom_bar(stat="identity")+
  geom_text(aes(label=n), vjust=1.6, color="white", size=3.5)+
  theme_classic() + facet_wrap(.~ miles) + ggtitle("Mile test by handedness") + 
  ylab("Count") + xlab("")
# six, show porta
a <- allsum %>%
  group_by(porta, handedness) %>%
    summarize(n = n())
ggplot(data=a, aes(x=handedness, y=n, fill= handedness)) +
  geom_bar(stat="identity")+
  geom_text(aes(label=n), vjust=1.6, color="white", size=3.5)+
  theme_classic() + facet_wrap(.~ porta) + ggtitle("Porta test by handedness") + 
  ylab("Count") + xlab("")
# seventh, show LexTALE
ggplot(allsum, aes(x = handedness, y = lexTALE)) +
  geom_pirate(aes(colour = handedness), bars = FALSE) + theme_classic() + 
  ylim(70, 110) + ggtitle("LexTALE by handedness") + ylab("Standardised score") + xlab("")
```

```{r corr_between_tasks}

mycols <- c('Day1_CF_acc_Z','Day2_CF_acc_Z','Day1_Dich_acc_Z','Day2_Dich_acc_Z','tlat_RDT_1','tlat_RDT_2','Day1_Finger_acc_Z','Day2_Finger_acc_Z')
colnums<- which(colnames(allsum) %in% mycols)
mycorrs<-correlate(allsum[,colnums],method='spearman')

mycorrsL <-correlate(allsum[allsum$handedness=='Left',colnums],method='spearman')
mycorrsR <-correlate(allsum[allsum$handedness=='Right',colnums],method='spearman')

#compare L and R handers
for (i in colnums){
  print(names(allsum)[i])
 print(t.test(allsum[,i]~allsum$handedness))
}

#check lateralisaiton with single group t vs zero
for (i in colnums){
  print(names(allsum)[i])
 print(t.test(allsum[,i]))
}

# corrplot the correlation matrix
mycols <- c('Day1_CF_acc_LI', 'Day1_CF_acc_Z','Day2_CF_acc_LI', 'Day2_CF_acc_Z',
            'Day1_DL_acc_LI', 'Day1_Dich_acc_Z','Day2_DL_acc_LI', 'Day2_Dich_acc_Z',
            'Day1.RDT_RT_LI', 'tlat_RDT_1', 'Day2.RDT_RT_LI', 'tlat_RDT_2',
            'Day1_finger_LI','Day1_Finger_acc_Z', 'Day2_finger_LI','Day2_Finger_acc_Z')
colnums<- which(colnames(allsum) %in% mycols)
M <- allsum[,colnums]
M <- M %>% 
  rename(
    "CF (z) 1" = Day1_CF_acc_Z,
    "CF (z) 2" = Day2_CF_acc_Z,
    "DL (z) 1" = Day1_Dich_acc_Z,
    "DL (z) 2" = Day2_Dich_acc_Z,
    "RD (t) 1" = tlat_RDT_1,
    "RD (t) 2" = tlat_RDT_2,
    "FT (z) 1" = Day1_Finger_acc_Z,
    "FT (z) 2" = Day2_Finger_acc_Z,
    "CF (LI) 1" = Day1_CF_acc_LI,
    "CF (LI) 2" = Day2_CF_acc_LI,
    "DL (LI) 1" = Day1_DL_acc_LI,
    "DL (LI) 2" = Day2_DL_acc_LI,
    "RD (LI) 1" = Day1.RDT_RT_LI,
    "RD (LI) 2" = Day2.RDT_RT_LI,
    "FT (LI) 1" = Day1_finger_LI,
    "FT (LI) 2" = Day2_finger_LI)
# PLOT TO SEE IF DIFFERENT FROM ZERO
# rearrange df
N <- M %>% select("CF (LI) 1", "CF (LI) 2", "CF (z) 1", "CF (z) 2",
             "DL (LI) 1", "DL (LI) 2", "DL (z) 1", "DL (z) 2",
             "RD (LI) 1", "RD (LI) 2", "RD (t) 1", "RD (t) 2",
             "FT (LI) 1", "FT (LI) 2", "FT (z) 1", "FT (z) 2")
# rearrange for pirate plot
# set participant ID
N$ID <- 1:nrow(N)
# format to long
data_long <- tidyr::gather(N, task, index, `CF (LI) 1`:`FT (z) 2`, factor_key=TRUE)
# add task identifer, session, etc
data_long <- data_long %>%
    mutate(group= if_else(task == "CF (LI) 1" | task == "CF (LI) 2" | task == "DL (LI) 1" | task == "DL (LI) 2" |
                          task == "RD (LI) 1" | task == "RD (LI) 2" | task == "FT (LI) 1" | task == "FT (LI) 2",
                          "LI", "t/z"),
           session= if_else(task == "CF (LI) 1" | task == "CF (z) 1" |
                            task == "DL (LI) 1" | task == "DL (z) 1" |
                            task == "RD (LI) 1" | task == "RD (t) 1" |
                            task == "FT (LI) 1" | task == "FT (z) 1", "1", "2"),
           task2= if_else(task == "CF (LI) 1" | task == "CF (LI) 2" | task == "CF (z) 1" | task == "CF (z) 2", "Chimeric faces",
                  if_else(task == "DL (LI) 1" | task == "DL (LI) 2" | task == "DL (z) 1" | task == "DL (z) 2", "Dichotic listening",
                  if_else(task == "RD (LI) 1" | task == "RD (LI) 2" | task == "RD (t) 1" | task == "RD (t) 2", "Rhyme decision",
                  "Finger Tapping"))))
# PLOT CORRELATION BETWEEN DAYS
# create matrix
M<-cor(M)
# removed odds and evens on each axis for plotting
dat <- M
toDeleteA <- seq(2, nrow(M), 2)
toDeleteB <- seq(1, ncol(M), 2)
dat <- dat[ toDeleteA , toDeleteB ]
dat=dat[,order(ncol(dat):1)]
# plot
ggcorrplot::ggcorrplot(dat, lab = TRUE, 
                       ggtheme = ggplot2::theme_minimal,
                       colors = c("#6D9EC1", "white", "#E46726"),
                       show.legend = FALSE, digits = 2, 
                       title = "Correlation matrix for LIs on the four behavioural laterality tasks")

# correlate each
# dichotic
DL_scatter <- ggplot(allsum, aes(x= Day1_DL_acc_LI, y= Day2_DL_acc_LI)) +
  geom_hline(yintercept=0, color = "black") +
  geom_vline(xintercept=0, color = "black") +
  geom_point(aes(color= handedness)) +
  geom_rug(aes(color= handedness)) +
  theme_classic() +
  ylab("Dichotic listening LI (day 2)") + 
  xlab("Dichotic listening LI (day 1)") +
  xlim(-100, 100) + ylim(-100, 100) +
  annotate("text", x = 90, y = 90, label = "paste(italic(R), \" = .72\")", parse = TRUE) +
  theme(legend.position = c(.15, .90))
# Set relative size of marginal plots (main plot 10x bigger than marginals)
ggExtra::ggMarginal(DL_scatter, groupColour = TRUE, groupFill = TRUE)
# chimeric
CF_scatter <- ggplot(allsum, aes(x= Day1_CF_acc_LI, y= Day2_CF_acc_LI)) +
  geom_hline(yintercept=0, color = "black") +
  geom_vline(xintercept=0, color = "black") +
  geom_point(aes(color= handedness)) +
  geom_rug(aes(color= handedness)) +
  theme_classic() +
  ylab("Chimeric faces LI (day 2)") + 
  xlab("Chimeric faces LI (day 1)") +
  xlim(-100, 100) + ylim(-100, 100) +
  annotate("text", x = 90, y = 90, label = "paste(italic(R), \" = .88\")", parse = TRUE) +
  theme(legend.position = c(.15, .90))
# Set relative size of marginal plots (main plot 10x bigger than marginals)
ggExtra::ggMarginal(CF_scatter, groupColour = TRUE, groupFill = TRUE)
# rhyme
rhyme_scatter <- ggplot(allsum, aes(x= Day1.RDT_RT_LI, y= Day2.RDT_RT_LI)) +
  geom_hline(yintercept=0, color = "black") +
  geom_vline(xintercept=0, color = "black") +
  geom_point(aes(color= handedness)) + 
  geom_rug(aes(color= handedness)) +
  theme_classic() +
  ylab("Rhyme decision LI (day 2)") + 
  xlab("Rhyme decision LI (day 1)") +
  annotate("text", x = 8, y = 8, label = "paste(italic(R), \" = .63\")", parse = TRUE) +
  theme(legend.position = c(.15, .90))
# Set relative size of marginal plots (main plot 10x bigger than marginals)
ggExtra::ggMarginal(rhyme_scatter, groupColour = TRUE, groupFill = TRUE)
# finger
finger_scatter <- ggplot(allsum, aes(x= Day1_finger_LI, y= Day2_finger_LI)) +
  geom_hline(yintercept=0, color = "black") +
  geom_vline(xintercept=0, color = "black") +
  geom_point(aes(color= handedness)) +
  geom_rug(aes(color= handedness)) +
  theme_classic() +
  ylab("Finger tapping LI (day 2)") + 
  xlab("Finger tapping LI (day 1)") +
  annotate("text", x = 18, y = 18, label = "paste(italic(R), \" = .89\")", parse = TRUE) +
  theme(legend.position = c(.15, .90))
# Set relative size of marginal plots (main plot 10x bigger than marginals)
ggExtra::ggMarginal(finger_scatter, groupColour = TRUE, groupFill = TRUE)
```

**References:**

- Burt, D. M., & Perrett, D. I. (1997). Perceptual asymmetries in judgements of facial attractiveness, age, gender, speech and expression. Neuropsychologia, 35(5), 685–693.
- Hoaglin, D. C., & Iglewicz, B. (1987). Fine-tuning some resistant rules for outlier labelling. Journal of the American Statistical Association, 82(400), 1147-1149. 
- Hugdahl, K., Westerhausen, R., Alho, K., Medvedev, S., Laine, M., & Hämäläinen, H. (2009). Attention and cognitive control: unfolding the dichotic listening story. Scandinavian journal of psychology, 50(1), 11-22.
- Innes, B. R., Burt, D. M., Birch, Y. K., & Hausmann, M. (2016). A leftward bias however you look at it: Revisiting the emotional chimeric face task as a tool for measuring emotion lateralization. Laterality: Asymmetries of Body, Brain and Cognition, 21(4-6), 643–661.
- Karlsson, E. M., Johnstone, L. T., & Carey, D. P. (2019). The depth and breadth of multiple perceptual asymmetries in right handers and non-right handers. Laterality: Asymmetries of Body, Brain and Cognition, 24(6), 707-739.