### 2020-05-24 battery-preprocessing_v5_DB.Rmd
Full analysis of RDT task. It is looking bad on all fronts - not very lateralised, not very reliable, no relation to handedness. Pairwise analysis of laterality by items looks even worse. Medians rather than mean RT checked; doesn't improve things.

---
title: "Online Laterality Battery Pre-Processing - short version"
author: "Adam Parker + Dorothy Bishop"
date: "updated May 21st 2020 (now notebook format)"
output: html_notebook
---
23 May 2020: Created subject number for those included in final analysis. CF outlier detection needed updating to ensure only correct trials used for outliers - Correct had not been coded at that point in prior version
22 May 2020: AP made changes to N trials needed for inclusion  
NB. 2 changes Apr 22nd by DB: 
a) all data now in a Data folder, so when reading need to specify that; 
b) *demographics* data frame renamed to *allsum*.

This R Markdown file processes the various components of the Online Laterality Battery which is implemented online via Gorilla.sc.

This version updated on 1st Feb 2020 to make script more reproducible and compact by using functions for repeated operations.
Adapted in April 2020 for processing dataset on 300 people

The script pulls in csv files and processes them so that pre-registered analyses can be conducted (pre-registered on the OSF: ). First, questionnaires are analyses followed by chimeric, dichotic, visual half field and tapping tasks. This is done for time 1 and time 2. A "full" laterality index output is then appended to a file with the demographics. This is then used in subsequent R markdown scripts.

This script analyses the following: 
1) Questionnaires
  - demographics
  - Miles test
  - Porta test
  - Edinburgh Handedness Inventory
  - lexTALE
2) Chimeric face task
3) Dichotic listening
4) Rhyme detection task
5) Key Pressing Task (finger dexterity)
And then produces summary of demog and results for the final sample

NB. Just also putting here relevant references that need to be looked at!
Bourne on RT in chimeric faces: https://pubmed.ncbi.nlm.nih.gov/18050003/

# Questionnaires

This R Markdown processes the following questionnaires from the Online battery:

- demographics
- Miles test
- Porta test
- Edinburgh Handedness Inventory
- lexTALE

Once processed, a file is made.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#NB for original data setwd("~/Dropbox/Online LDT/300 Ps")
# libraries 
library(readr)
library(yarrr)
library(dplyr)
library(tidyr)
library(car)
library(ggplot2)
library(knitr)
require(readr)  # for read_csv()
require(dplyr)  # for mutate()
require(purrr)  # for map(), reduce()
library(GGally)
require(stringr) # string matching
library(ggpirate)
library(corrr) #easier correlations!
library(beeswarm) #just a kind of plot Dorothy likes
```

First, the demographics is processed.This gives us a participant subject, age, gender, footedness, and handedness.

```{r demograhics, warning=FALSE}
#Reading from OSF - this only works if files are public

demo_dat <- read.csv('https://osf.io/7smj6/download',stringsAsFactor=F)

# relabel 
demo_dat$subject <- demo_dat$Participant.Public.ID
demo_dat$subject <- as.factor(demo_dat$subject)
# count and print subjects
nsub <-length(levels(demo_dat$subject))
print(paste0("subjects: ", nsub))

# create data frame
allsum <- data.frame(matrix(ncol = 7, nrow = nsub))
  colnames(allsum) <- c("subject", "age", "gender", "footedness", "handedness", "bilingual", "strongest.language")
myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  subname <- levels(demo_dat$subject)[i] # find subject subject
  myrows <- which(demo_dat$subject==subname) # select rows for this subject
  tmp <- data.frame(demo_dat[myrows,])

    myrow <- i

    allsum$subject[myrow] <- subname
    allsum$age[myrow] <- tmp[tmp$Question.Key == "Age",]$Response
    allsum$gender[myrow] <- tmp[tmp$Question.Key == "Gender",]$Response
    allsum$footedness[myrow] <- tmp[tmp$Question.Key == "footedness",]$Response
    allsum$handedness[myrow] <- tmp[tmp$Question.Key == "categorical_hand",]$Response
    allsum$bilingual[myrow] <- tmp[tmp$Question.Key == "bilingual",]$Response
    allsum$strongest.language[myrow] <- tolower(tmp[tmp$Question.Key == "bilingual-text",]$Response)

    allsum$bilingual <- as.character(allsum$bilingual)
    allsum$bilingual <- gsub('. My strongest language is:', '', allsum$bilingual)
}

#Check ages
range(allsum$age)
# Here we summarise the age broken down by handedness and gender. This will be important for any write up. 
#### NOTE. THIS IS NOT REPRESENTATIVE OF THE FINAL DATASET
# age
allsum %>%
    group_by(handedness, gender) %>%
    summarize(avg = mean(age), n = n(), sd = sd(age))

#create a numerical code for handedness - can be used in plots to distinguish handedness groups (eg if pch set to handpch, get circles and triangles)
allsum$handpch <-1  #numeric code for plotting handedness
allsum$handpch[allsum$handedness=='Left'] <-2
```

Now we import and append the Porta and Miles Test to the allsum data frame.

```{r sightednesss, warning=FALSE}
porta_dat <- read.csv("https://osf.io/zmqkt/download",stringsAsFactors = F)
miles_dat <-  read.csv("https://osf.io/ekqsp//download",stringsAsFactors = F)

miles_dat$test_subject <- "miles"
porta_dat$test_subject <- "porta"
# merge the two
sight_dat <- rbind(miles_dat, porta_dat)
# relabel 
sight_dat$subject <- sight_dat$Participant.Public.ID
sight_dat$subject <- as.factor(sight_dat$subject)

# count and print subjects
nsub <-length(levels(sight_dat$subject))
print(paste0("subjects: ", nsub))

# create data frame
sightedness <- data.frame(matrix(ncol = 3, nrow = nsub))
  colnames(sightedness) <- c("subject", "miles", "porta")
myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  subname <- levels(sight_dat$subject)[i] # find subject subject
  myrows <- which(sight_dat$subject==subname) # select rows for this subject
  tmp <- data.frame(sight_dat[myrows,])

    myrow <- i

    sightedness$subject[myrow] <- subname
    sightedness$miles[myrow] <- tmp[tmp$Question.Key == "response-2",]$Response
    sightedness$porta[myrow] <- tmp[tmp$Question.Key == "response-2",]$Response
  }

# now merge sightedness and allsum
allsum <- merge(allsum, sightedness, by= "subject")
head(allsum) #just to check all looking OK
```

Now we add the EHI. Here we score and plot the Edinburgh handedness inventory (Oldfield, 1971) in its short form.

```{r EHI, warning=FALSE}

EHI <- read.csv("https://osf.io/kfdma/download",stringsAsFactors = F)

# select only responses
EHI <- EHI[EHI$Question.Key == "response-2-quantised" | EHI$Question.Key == "response-3-quantised" | 
           EHI$Question.Key == "response-4-quantised" | EHI$Question.Key == "response-5-quantised" |
           EHI$Question.Key == "response-6-quantised" | EHI$Question.Key == "response-7-quantised" | 
           EHI$Question.Key == "response-8-quantised" | EHI$Question.Key == "response-9-quantised" |
           EHI$Question.Key == "response-10-quantised" | EHI$Question.Key == "response-11-quantised" | 
           EHI$Question.Key == "response-12-quantised" | EHI$Question.Key == "response-13-quantised" |
           EHI$Question.Key == "response-14-quantised" | EHI$Question.Key == "response-15-quantised" | 
           EHI$Question.Key == "response-16-quantised" | EHI$Question.Key == "response-17-quantised" |
           EHI$Question.Key == "response-18-quantised" | EHI$Question.Key == "response-19-quantised" | 
           EHI$Question.Key == "response-20-quantised",]

# recode variabels
EHI$subject <- EHI$Participant.Public.ID
EHI$subject <- as.factor(EHI$subject)

# reduce data
# new data
meaningful <- c("subject", "Question Key", "Response") # select wanted columns
c<-which(names(EHI) %in% meaningful) #find colnumbers of unwanted
EHI <-EHI[,c] #remove unwanted columns
EHI <- na.omit(EHI) # remove NAs

# score responses according to the original scoring method in Oldfield (1971)
EHI$right_hand <- 0
EHI$left_hand <- 0
for (r in 1:nrow(EHI)) {
  if(EHI$Response[r] == 1){
    EHI$right_hand[r]= 2 
  } else {
    if(EHI$Response[r] == 2){
      EHI$right_hand[r]= 1
    } else {
      if(EHI$Response[r] == 4){
        EHI$left_hand[r]= 1
      } else {
        if(EHI$Response[r] == 5){
          EHI$left_hand[r]= 2 
        } else {
          (EHI$right_hand[r]= 1) & (EHI$left_hand[r]= 1)
        }
      }
    }
  }
}

# re-reduce data
# new data
meaningful <- c("subject", "right_hand", "left_hand") # select wanted columns
c<-which(names(EHI) %in% meaningful) #find colnumbers of unwanted
EHI <-EHI[,c] #remove unwanted columns
EHI <- na.omit(EHI) # remove NAs

# find sum for left and right for each participant 
# create hand data 
nsub <-length(levels(EHI$subject))
EHI2 <- data.frame(matrix(ncol = 3, nrow = nsub))
  colnames(EHI2) <- c("subject", "right_hand", "left_hand")

for (i in 1:nsub) { # loop through subjects
  myrow <- i
  subname <- levels(EHI$subject)[i] # find subject subject
  myrows <- which(EHI$subject==subname) # select rows for this subject
  tmp <- data.frame(EHI[myrows,])

    EHI2$subject[myrow] <- subname # add row for subject
    EHI2$right_hand[myrow] <- sum(tmp$right_hand) # sum for R hand
    EHI2$left_hand[myrow] <- sum(tmp$left_hand) # sum for L hand
  }

EHI2$index_EHI <- 100*(EHI2$right_hand-EHI2$left_hand)/(EHI2$right_hand+EHI2$left_hand)

# now create a data frame only for index 
# merge with allsum
allsum$index_EHI <- NA #initialise
for (i in 1:nrow(allsum)){
  thissub <- allsum$subject[i]
  w<-which(EHI2$subject==thissub)
 allsum$index_EHI[i] <- EHI2$index_EHI[w]
}

#now plot EHI in relation to handedness (modified by DB)

eplot<-ggplot(allsum, aes(handpch,index_EHI, colour=gender)) +
  geom_jitter(wsubjectth=.25,height=2)+
  ggtitle("Handedness index on the Edinburgh Handedness Inventory")
eplot
#This is not perfect, but gives and subjectea to check EHI is valid etc.

ggplot(allsum, aes(x=index_EHI)) + geom_density(aes(group=handpch,fill=handpch))

#NB two supposed right-handers have index_EHI around -50. 
#I will recode these, as this seems an error.

oddcase <- intersect(which(allsum$index_EHI<0),which(allsum$handedness=='Right'))
allsum$handedness[oddcase] <- 'Left'
allsum$handpch[oddcase] <- 1
```

Next, the lexTale is preprocesssed. This is done using the weighted average for the test.

```{r lexTALE, warning=FALSE}
# read data
lexTALE <- read.csv("https://osf.io/h4btp/download",stringsAsFactors = F)

# filter
lexTALE <- lexTALE[lexTALE$Attempt == 1,]
lexTALE <- lexTALE[lexTALE$display == "Task",]
# rename variables
lexTALE$subject <- lexTALE$Participant.Public.ID
lexTALE$accuracy <- lexTALE$Correct
lexTALE$condition <- lexTALE$ANSWER
# code as factors
lexTALE$subject <- as.factor(lexTALE$subject)
lexTALE$condition <- as.factor(lexTALE$condition)
  levels(lexTALE$condition) <- c("non-word", "word")

# create data frame
lexTale_score <- data.frame(matrix(ncol = 4, nrow = nsub))
  colnames(lexTale_score) <- c("subject", "Word", "nonWord", "lexTALE")

for (i in 1:nsub) { # loop through subjects
  subname <- levels(lexTALE$subject)[i] # find subject subject
  myrows <- which(lexTALE$subject==subname) # select rows for this subject
  tmp <- data.frame(lexTALE[myrows,])

    myrow <- i

    lexTale_score$subject[myrow] <- subname
    lexTale_score$Word[myrow] <- sum(tmp[tmp$condition == "word",]$accuracy)
    lexTale_score$nonWord[myrow] <- sum(tmp[tmp$condition == "non-word",]$accuracy)
    lexTale_score$lexTALE[myrow] <- ((sum(tmp[tmp$condition == "word",]$accuracy)/40*100) + (sum(tmp[tmp$condition == "non-word",]$accuracy)/20*100)) / 2

}

# plot
x<-ggplot(lexTale_score, aes(x=lexTALE, y=subject, fill=subject)) +
  geom_dotplot(binaxis='y', stackdir='center') + xlim(0, 100) +
  ggtitle("Lextale") + geom_vline(xintercept = 80, linetype="dashed", color = "black", size=1.5)
x + theme_bw() + theme(legend.position = "none") + theme(axis.text.y = element_blank())

# now create a data frame only for index 
lexTALE <- select(lexTale_score, "subject", "lexTALE")
# now merge with allsum
allsum <- merge(allsum, lexTALE, by= "subject")
# now write the full allsum 
write.csv(allsum, file = "./allsum_out.csv")

hist(allsum$lexTALE)
abline(v=80,col='red')

paste0("No. of participants below LexTALE cuttoff: ", sum(allsum$lexTALE < 80))

# remove those who dIDn't pass the LexTALE
allsum <- allsum[allsum$lexTALE >= 80,]
```

# Read in raw data for day 1 and 2

Before doing anything with the data, we now pull in all the data from OSF and combine them. Participants, identified by random letter strings, are stacked on top of each other. We start by reading the data and cutting out unwanted information.  
NB This step is quite slow (slower if reading from OSF than from drive).

```{r readraw,include= FALSE, warning=FALSE}
# Set directory
for (d in 1:2){
  data_file <- "https://osf.io/mhk29/download" #day 1 data

  if (d ==2){
    data_file <-  "https://osf.io/ebm6s/download"
  }
  #we will assign raw data to all_dat2 (day2), but then reassign to all_dat1 on first run through loop, so we have correct data for each day
all_dat2 <- read.csv(data_file, na.strings = "NA",stringsAsFactors=F)
# cleaning - do operations that apply for all tasks
all_dat2 %>% filter(Attempt >= 1) # only answers 
# write as factors
all_dat2$subject <- as.factor(all_dat2$Participant.Public.ID)
all_dat2$RT <- as.numeric(as.character(all_dat2$Reaction.Time))
all_dat2$Trial <-as.numeric(all_dat2$Trial.Number) #useful if we want to restrict analysis to subset of trials so retain this variable
all_dat2$Task.Name <- as.factor(all_dat2$Task.Name)
all_dat2$word<-substring(all_dat2$word,66,(nchar(all_dat2$word)-4)) #remove format from written word stimuli
if (d==1)
{all_dat1 <- all_dat2}
}

all_dat2$word_OG<-NA #create dummy column just to make it easier to align columns
collist <- colnames(all_dat1)
all_dat2 <- all_dat2[,collist]
```

Create a simple numerical ID to make life easier when screening data if we need to identify subjects whose data is odd in some way. Ensure this is consistent across all_dat1 and all_dat2 and allsum.
We first rename the current subject variable so we can reuse this name with the numbered ID for all future computations.
Again, this is rather slow because done in a loop, which is probably not very efficient!

```{r makesubject}
#Use allsum as the basis for the check
#First rename all subject to origID
allsum <- allsum %>% rename(origID = subject) 
all_dat1 <- all_dat1 %>% rename(origID = subject) 
all_dat2 <- all_dat2 %>% rename(origID = subject) 
all_dat1$subject <- NA
all_dat2$subject <- NA
for (n in 1:nrow(allsum)){
  allsum$subject[n] <- n
  w<-which(all_dat1$origID == allsum$origID[n])
  all_dat1$subject[w] <-n
  w<-which(all_dat2$origID == allsum$origID[n])
  all_dat2$subject[w] <-n
}
head(allsum)
```

# Functions

This is a function that can apply to all tasks; user specifies the original file to start from, the wanted columns, and the specific task to select.

```{r genericprocessing} 
# This function just returns a subfile with just the necessary rows/cols
  procdata <- function(origfile,wanted,task,disp){
  nufile <- origfile[origfile$Task.Name==task,] #restrict consideration to this task
  nufile <- origfile[origfile$display==disp,] #restrict consideration to the correct display

  c<-which(names(nufile) %in% wanted) #find colnumbers of wanted
  #NB beware that the columns may not be in the same order as in the wanted list
  nufile <- nufile[,c]#select only wanted columns
  nufile <- na.omit(nufile) # remove NAs
  return(nufile)
}
```

The Hoaglin-Iglewicz function is used to remove outliers. We create an 'outlier' column that codes anticipations as 2, and long RT outliers as 1. All else is zero. The original RT is saved as allRT, and the RT column has NA for outliers, so they will be automatically excluded henceforth.

Outlier removal put in separate function, which can be applied to different task RTs.  
Origdat is the dataframe with data for outlier removal, mycolname is name of variable  
For outlier removal (usually RT), lowcut is a cutoff value below which data excluded (typically v fast responses that are implausible, e.g. 201 ms for RT)
 and zcut is the constant in Hoaglin-Iglewicz method (usually 2.2).
 Here the focus is on just one tail of the distribution (slow RTs) so we use 1.65 as zcut.

```{r RToutlierfunction}
detect.outliers <- function(origdat,lowcut,zcut){

# NB this function returns origdat with outlier rows marked, rather than removed
# This makes it easier to record the N outliers per subject.

# NB FUNCTION ASSUMES THAT THERE IS ARE COLUMNS CALLED subject, Correct and RT IN origfile.
# Outliers are computed by subject - NB contrary to previous version, these are defined
# across all trials, rather than separately for L and R.
origdat$allRT <- origdat$RT #we save a copy of RT, as we will be altering RT so that outliers become NA
origdat$outlier<-0
origdat$outlier[origdat$RT<lowcut] <- 2 #outlier col is coded 2 for anticipations
origdat$RT[origdat$RT<lowcut]<-NA #we exclude anticipations when computing quantiles
#Now extract correct RTs for each subject to compute limits for outliers

  for (i in 1:nsub) { # loop through subjects
  subname <- origdat$subject[i] # find subject 
  myrows <- which(origdat$subject==subname) # select rows for this subject
  #NB myrows is NOT a consecutive series, because done in blocks
  firstrow <-min(myrows)
  tmp <- data.frame(origdat[myrows,])

  RTcorr <-tmp$RT[tmp$Correct==1] #correct RTs for this subject as a vector

 # IDentify 25th and 75th quartiles, and the difference between them
          lower_quartile <- quantile(RTcorr, probs=0.25, na.rm="TRUE")
          upper_quartile <- quantile(RTcorr, probs=0.75, na.rm="TRUE")
          quartile_diff <- upper_quartile - lower_quartile
        # Outliers are defined as being below or above 2.2 times the quartile difference
          lower_limit <- lower_quartile - zcut*quartile_diff
          upper_limit <- upper_quartile + zcut*quartile_diff
        # create outlier variable
w1 <- which(origdat$RT[myrows]>upper_limit) #rows with outlier, nb RELATIVE to myrows range
origdat$outlier[myrows[w1]]<-1
}
 #For RT we just remove slow outliers, so ignore lower_limit
origdat$RT[origdat$outlier>0]<-NA #RT variable now has outliers excluded as NA
return(origdat)
}
```

This is a generic chunk of code that can be used for Chimeric faces and Rhyme Detection.

```{r maketaskdataframe, warning=FALSE}
make.df <- function(origdat,varlist,latlist,nloop){
nsub<-length(unique(origdat$subject))
# create compact data frame ; one row for each subject, subjects stacked, L above R
temp_dat <- data.frame(matrix(ncol = 4, nrow = nsub*nloop))
colnames(temp_dat) <- varlist
myrow <- 0 # start row counter
for (i in 1:nsub) { # loop through subjects
  subname <- i # find subject 
  myrows <- which(origdat$subject==subname) # select rows for this subject
  latcol <- which(colnames(origdat) == varlist[2])
  RTcol <- which(colnames(origdat) == varlist[4])
  tmp <- data.frame(origdat[myrows,]) #all trials from this subject

  for (j in 1:nloop) { #make row for means for each side for this subject
    myrow <- myrow+1
    w<- which(tmp[,latcol]== latlist[j]) # match on laterality
    tmp1 <- tmp[w,]
    tmp2 <- tmp1[tmp1$Correct == 1,]

    temp_dat$subject[myrow] <- subname # add row for subject
    temp_dat$side[myrow] <- latlist[j] # add row for sIDe
    temp_dat$accurate[myrow] <- sum(tmp1$Correct,na.rm=TRUE) # add row for N accurate response
    temp_dat$p.corr[myrow] <- 100*mean(tmp1$Correct,na.rm=TRUE) # add row for %accurate responses - NB for chimeric only error is wrong emotion - measure is just a measure of choice. For RDT, can make error of selecting wrong sIDe, so % correct is meaningful

    temp_dat$RT[myrow] <- mean(as.numeric(tmp2$RT),na.rm=TRUE) # add row for mean RT
    #NB for RDT task, also looked at medians, but it did not make much difference, presumably because v slow responses already omitted via outlier routine
    temp_dat$N[myrow] <- length(tmp1$RT) # add count of rows for this subject
  }
  #If there are no selections/trials for one side, then exclude L and R
  # if(temp_dat$N[myrow==0]||temp_dat$N[(myrow-1)==0]){
  #   temp_dat$RT[(myrow-1):myrow]<-NA
  # }
}
return(temp_dat)
}
```

This plots data and conducts t-tests for each task.

```{r plotdata_ttest, warning=FALSE}
dopirate <- function(origdat,task,measure){
# measure for each sIDe
# plot the number of correct responses in each sIDe
  myhead<-paste0(task,": ",measure)
pirateplot(formula = X1 ~ side,
           data = origdat,
           main = myhead,
           theme = 0,
           pal = "southpark", # southpark color palette
           bean.f.o = .0, # Bean fill
           point.o = .3, # Points
           inf.f.o = .7, # Inference fill
           inf.b.o = .8, # Inference border
           avg.line.o = 1, # Average line
           bar.f.o = .5, # Bar
           point.pch = 21,
           point.cex = .7,
           inf.method= "ci")
# plot L & R to get slope
q <- ggplot(data=origdat, aes(x=side, y=X1, group=as.factor(subject), color=subject)) +
    geom_line() +
    geom_point() +
    ggtitle(paste0(task,": ",measure," by side"))
q + theme_bw() + theme(legend.position = "none")
#  t test
print(paste0(task,measure))
myt1<-t.test(origdat$X1~ origdat$side, paired = TRUE, alternative = "two.sided")
print(myt1)
}
```

Another generic function that can be used for different tasks to make a laterality index and write it to the allsum file

```{r makeLIs, warning=FALSE}
# This function modifies a copy of the allsum file and returns it

# User selects whether accuracy or RT by specifying col X1 in origdat when calling the function

# Polarity is -1 or 1 and can be set to keep all LIs positive if in a given direction
# Cutoff will create NA if level of performance is outside predetermined limits

makeLI <- function(allsum,origdat,task,mycolname,polarity,mycutoff){
# we do this for accuracy first
# reformat data frame so left and right are side by side
keeps <- c("subject", "side", "X1") #X1 holds either accuracy or RT
Lsubjectf <- origdat[keeps]
Lsubjectf <- spread(Lsubjectf, side, X1)
# calculate each participants' lat index

Lsubjectf$thisLI <- 100*(Lsubjectf$Right - Lsubjectf$Left)/(Lsubjectf$Right + Lsubjectf$Left)

Lsubjectf$LI <- Lsubjectf$thisLI*polarity #polarity can be set to reverse sign - eg for RT a large score is bad, whereas for acc it is good, so can have opp polarity

# check if accuracy or RT level is outside acceptable cutoff.
# Do this based on average for left and right
# LI will be set to NA for those outside limits
Lsubjectf$avg <- (Lsubjectf$Left+Lsubjectf$Right)/2
w<-which(Lsubjectf$avg<mycutoff)
if(length(w)>0){
Lsubjectf$LI[w] <- NA}

# now create a data frame for lat index 
Lsubjectf <- select(Lsubjectf, "subject", "LI")
# now merge with allsum
# First check if this column already exists and delete it if so
w<-which(colnames(allsum)==mycolname)
if(length(w)>0){
allsum<-allsum[,-w]}

#now add this LI
allsum <- merge(allsum, Lsubjectf, by= "subject",all.x=T)
#Need all.x = T to retain all original subjects from allsum, even if no match
lastcol<-length(colnames(allsum))
colnames(allsum)[lastcol]<-mycolname

return(allsum)
}
```

Make Z score LI.
NB this is NOT a conventional z-score based on whole sample, but rather is an individual-based z-score that indicates whether the person's L-R difference is significantly different from chance.
So you either compare their L-sided RTs and their R-sided RTs, or the proportions correct on L vs proportion correct on R, to see if there is bias away from zero in that individual.
Note that this in effect adjusts for any overall differences in accuracy and RT, as the person is compared with themselves.

```{r makeZ, warning=FALSE}
# This function modifies a copy of the allsum file and returns it

# User selects the measure by specifying col X1 in origdat when calling the function

# Polarity is -1 or 1 and can be set to keep all LIs positive if in a given direction
# Cutoff will create NA if level of performance is outside predetermined limits

makeZ <- function(allsum,origdat,task,mycolname,polarity,mycutoff){
# we do this for accuracy first
# reformat data frame
keeps <- c("subject", "side", "X1") #X1 holds either accuracy or RT
Zdf <- origdat[keeps]
Zdf <- spread(Zdf, side, X1)
# calculate each participants' z score for laterality 
Zdf <- 
  Zdf %>% 
  group_by(subject) %>%
  mutate(n= Left + Right, 
         pR= Right/n,
         pL= Left/n,
         Z= (pR-.5)/sqrt(pR*pL/n))
# remove infinite results
is.na(Zdf) <- do.call(cbind,lapply(Zdf, is.infinite))
Zdf <- na.omit(Zdf)

Zdf$Z <- Zdf$Z*polarity #polarity can be set to reverse sign - eg for RT a large score is bad, whereas for acc it is good, so can have opp polarity

# now create a data frame for lat index 
Zdf <- select(Zdf, "subject", "Z")
# now merge with allsum
# First check if this column already exists and delete it if so
w<-which(colnames(allsum)==mycolname)
if(length(w)>0){
allsum<-allsum[,-w]} #delete cols that already exist with that name

allsum <- merge(allsum, Zdf, by= "subject",all.x=T)
lastcol<-length(colnames(allsum))
colnames(allsum)[lastcol]<-mycolname

return(allsum)
}
```

## Chimeric Face Task

This part of the notebook processes the "Chimeric Face Task' implmented online via Gorilla.sc which aims to assess the lateralisation of emotion recognition in a face processing task. In the task, participants are shown a chimeric face. Participants then respond by indicating the emotion that they felt was most strong expressed.

This task is near identical to that reported in Karlsson, Johnstone, and Carey (2019). Stimuli were kindly provided by Dr Michael Burt (https://www.dur.ac.uk/psychology/staff/?subject=1942) and were symmetrical average images created from four male and four female faces (see Burt & Perrett, 1997; Innes et al., 2016). Stimuli for each combination of anger, disgust, happy and sad (8 left-first, 8 right-first) plus 8 of each emotion where both sides the same the same to check overall perception. Total of 96 chimeric stimuli presented in each session. We will exclude those who are inaccurate when both faces the same (as in preregistration). Responses to chimeric faces are later coded as whether they correspond to emotion on L or R. Note that if the reported emotion does not match either side, then trial will not be counted in LI

```{r read_face_dat, warning=FALSE}
for (d in 1:2){ #we will process day 1 and day 2 together
  origfile <- all_dat1
  if (d==2)
  {origfile <- all_dat2}
#As with earlier reading in, we default to saving CF_dat2 (day2), but at end of the chunk we reassign to CF_day1 in the first run through the d loop
task <- "Chimeric faces"
wanted <- c("subject", "stimuli", "Response", "RT", "l_hemiface", "r_hemiface","Correct") # select wanted columns - nb added Correct here
disp <- "e_Task"
CF_dat_all <- procdata(origfile,wanted,task,disp) #use generic function (see above) to read in relevant columns

#Before detecting outliers need to add a column indicating if response was correct. Although we read in 'Correct' it is not accurate!

w<-union(which(CF_dat_all$Response==CF_dat_all$l_hemiface),
         which(CF_dat_all$Response==CF_dat_all$r_hemiface))
CF_dat_all$Correct[w]<-1

# count subjects
nsub <-length(unique(CF_dat_all$subject)) #previously we used the ID code and had to count levels of the factor; now with numbers we just use unique
print(paste('day',d,task,": subjects: ", nsub))

CF_same2 <- CF_dat_all[CF_dat_all$l_hemiface == CF_dat_all$r_hemiface,] # create data frame for non-chimeras
CF_dat2 <- CF_dat_all[CF_dat_all$l_hemiface != CF_dat_all$r_hemiface,] # create df for chimeras: remove items where same expression in both hemifaces
if (d==1)
{CF_dat1 <- CF_dat2
 CF_same1 <- CF_same2}
}
```

The chunk below detects outliers, and codes them as 1 if the RT is outside the range specified in Hoaglin-Iglewicz, and as 2 if the RT is < 200. Outliertable shows each subject with Ns with different outlier codes. Some people have 0-1 trials - these seem to be ones where the task was skipped for some reason - I have checked and they seem to have no data.

```{r CFoutliers}
CF_dat1 <- detect.outliers(CF_dat1,200,1.65) #use generic outlier detection function. The parameters are lowcut (200 ms cutoff for anticipations) and then the value used for Hoaglin Iglewicz (1.65)
outliertable1<-table(CF_dat1$subject,CF_dat1$outlier)
CF_dat2 <- detect.outliers(CF_dat2,200,1.65) #use generic outlier detection function. The parameters are lowcut (200 ms cutoff for anticipations) and then the value used for Hoaglin Iglewicz (1.65)
outliertable2<-table(CF_dat2$subject,CF_dat2$outlier)
#Most have 96 accepted trials in CF_dat2. A few have 1-2 - these seem to be ones that had RTs < 200, and may have been responding rapidly at random.

```

First, we judge participants' ability to correctly identify emotions using the "CF_same" dataframe. Participants are flagged if their performance is below 75% as this indicates poor ability to identify emotions, which will inevitably influence performance on the Chimeric Face Task.

```{r emotions, warning=FALSE,message=FALSE}
# this will mark correct responses as 1 or 0. It will be importantfor removing participants later on. 
# NB we loop through same process for day1 and day2
for (d in 1:2){
  CF_same <- CF_same1
  CF_diff <- CF_dat1
  if (d==2){
    CF_same <- CF_same2
    CF_diff <- CF_dat2
  }

  # removed code to compute correct, as already done above
  # calculate participants average
  #update: do this for chimeric as well as Same trials
  emot_corr_Same <- aggregate(FUN= mean, data= CF_same, Correct~ subject)
  # add this to allsum 
  colnames(emot_corr_Same)[2]<-'Emot_corr_same'
  allsum <- merge(allsum, emot_corr_Same, by= "subject")
  emot_corr_Diff <- aggregate(FUN= mean, data= CF_diff, Correct~ subject)
  # add this to allsum 
  colnames(emot_corr_Diff)[2]<-'Emot_corr_diff'
  allsum <- merge(allsum, emot_corr_Diff, by= "subject",all.x=T)
}

w<-which(colnames(allsum) %in% c("Emot.corr.same1","Emot.corr.diff1","Emot.corr.same2","Emot.corr.diff2"))
if (length(w)>0){
  allsum<-allsum[,-w] #remove any columns already created with these names
}
mycol <- length(colnames(allsum))

colnames(allsum)[(mycol-3):mycol]<-c("Emot.corr.same1","Emot.corr.diff1","Emot.corr.same2","Emot.corr.diff2")
```

```{r excludelowCFacc}
#db modified to mark rather than drop cases with scores < .75
allsum$exclude <- 0 #default is to exclude

# mark the cases where there is less than 75% for identical stimuli
w <- union(which(allsum$Emot.corr.same1 < .75),which(allsum$Emot.corr.same2 < .75))
allsum$exclude[w]<-1 #code exclude = 1 if excluded because poor emot recog
print(paste('N flagged for exclusion because < 75% on at least one trial:',length(w)))

```

```{r plotCFacc}
plot(jitter(allsum$Emot.corr.same1,10),jitter(allsum$Emot.corr.same2,10),pch=3,cex=.6,xlab='Nonchimeric emotion recog 1',ylab='Nonchimeric emotion recog 2',col=(1+allsum$exclude))
abline(h=.75)
abline(v=.75)
text(.6,.8,'red: exclude if <75% on either run',cex=.8)

# Just for interest look also at % correct on chimeric.
# Presumably the conflicting information makes this harder so we don't use this for exclusion
plot(jitter(allsum$Emot.corr.diff1,10),jitter(allsum$Emot.corr.diff2,10),pch=3,cex=.6,xlab='Chimeric emotion recog 1',ylab='Chimeric emotion recog 2',col=(1+allsum$exclude),main='Proportion correct for chimeric (not used for exclusion)')
abline(h=.75)
abline(v=.75)
text(.6,.95,'Red excluded, <75% on same')

```

In this section, we mark whether the correct answers were in the left or right hemiface.

```{r mark_face_correct_face, warning=FALSE}
# this will mark correct responses as 1 or 0.
# NB we again loop through same process for day1 and day2
for (d in 1:2){
  CF_dat <- CF_dat1
  if (d==2){
    CF_dat <- CF_dat2
  }

# this will mark the side responded to. NB to allow for generic functions for all tasks, we use the label 'side' rather than hemiface
CF_dat <- 
  CF_dat %>% 
  mutate(side= ifelse(as.character(Response) == as.character(l_hemiface) & Correct == 1, "Left",
                       ifelse(as.character(Response) == as.character(r_hemiface) & Correct == 1, "Right", "X")))

#Added coding of 'X' for responses of emotion not shown in either side

# code side as factor
CF_dat$side <- as.factor(CF_dat$side)

  if(d==1){
    CF_dat1 <- CF_dat}
  if (d==2){
    CF_dat2 <- CF_dat
  }
}
#sanity check: looks fine. X means response was emotion of neither side
 table(CF_dat1$side)
 table(CF_dat2$side)
```

Now the data is ready, create an empty data frame and populate this. It includes columns for the participant identifier, side, and count of accurate responses for each side.

```{r makedataframe, warning=F}
#We create chim_dat files for day1 and day 2
varlist <-   c("subject", "side", "accurate", "RT") #need to be in this order, ie sub, side, acc and RR
latlist <- c("Left","Right") #names of factor levels
nloop=2 #same as levels of latlist
chim_dat1 <- make.df(CF_dat1,varlist,latlist,nloop)
chim_dat2 <- make.df(CF_dat2,varlist,latlist,nloop)

# w<-which(is.na(chim_dat1$subject))
# if (length(w)>0) {chim_dat1<-chim_dat1[-w,]}
# w<-which(is.na(chim_dat2$subject))
# if (length(w)>0) {chim_dat2<-chim_dat2[-w,]}

```

The data is visualised using the pirate plot function and t.test conducted.

```{r dosummary}
#find included cases
w<-which(allsum$exclude==0)
inclsubs <- allsum$subject[w]

#First look at accuracy, Day 1
origdat <- chim_dat1[(chim_dat1$subject %in% inclsubs),]
measure <- 'Accuracy'
origdat$X1 <- origdat$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc with the dopirate function
task <- 'Chimeric faces day 1'
dopirate(origdat,task,measure)
#Same function will do RT next
origdat$X1 <- origdat$RT
measure <- 'RT'
dopirate(origdat,task,measure)
t.test(origdat$accurate~ origdat$side)
#Now repeat for day 2
#-------------------------
#We need to add a step to exclude accuracy and RT if fewer than 3 trials in one condition. This only affects 2nd session data, so this code is specific to that file
w<-which(chim_dat2$N<2)
dropsub <- chim_dat2$subject[w]
chim_dat2$accurate[chim_dat2$subject %in% dropsub]<-NA
chim_dat2$RT[chim_dat2$subject %in% dropsub]<-NA
#-------------------------
origdat <- chim_dat2[(chim_dat2$subject %in% inclsubs),]
origdat$X1 <- origdat$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Chimeric faces day 2'
measure <- 'Accuracy'
dopirate(origdat,task,measure)

origdat$X1 <- origdat$RT
measure <- 'RT'
dopirate(origdat,task,measure)
```

Next, a laterality index is calculated and plotted based on the equation [(R-L)/(R+L)] x 100, where L= left side and R= right side.

If the index is negative, emotion recognition is lateralised in the left hemisphere. If the index is positive, emotion recognition is lateralised in the right hemisphere

Here, a .csv is written for the chimeric face LI.

```{r domakeLI}
#We add Chimeric faces LIs to allsum for day 1 and then day2
task<-'Chimeric faces'
origdat<-chim_dat1
origdat$X1 <- origdat$accurate
mycolname<-"Day1_CF_acc_LI"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

# now do RT
origdat$X1 <- origdat$RT
mycolname<-"Day1_CF_RT_LI"
polarity <-1

#if we don't specify mycutoff, it will remain at level set for accuracy, which means it should not trap any RT values as they will all be much bigger
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

origdat<-chim_dat2
origdat$X1 <- origdat$accurate
mycolname<-"Day2_CF_acc_LI"
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
origdat$X1 <- origdat$RT
mycolname<-"Day2_CF_RT_LI"
polarity <-1
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

cor1 <- cor(allsum$Day1_CF_acc_LI,allsum$Day2_CF_acc_LI,use='complete.obs')
plot(allsum$Day1_CF_acc_LI,allsum$Day2_CF_acc_LI,main='Chimeric LI accuracy',col=(1+allsum$exclude))
abline(v=0)
abline(h=0)
text(-50,75,paste0('r = ',round(cor1,3)))
myn <- nrow(allsum)
text(-50,85,paste0('N = ',myn))
cor2<-cor(allsum$Day1_CF_RT_LI,allsum$Day2_CF_RT_LI,use='complete.obs')
plot(allsum$Day1_CF_RT_LI,allsum$Day2_CF_RT_LI,main='Chimeric LI RT',col=(1+allsum$exclude)) #this would plot excluded as red, but they are now excluded at earlier stage so don't show up
abline(v=0)
abline(h=0)
text(-10,20,paste0('r = ',round(cor2,3)))

text(-10,23,paste0('N = ',myn))
#text(-10,18,'red will be excluded',cex=.85)
```

Here we use the Z score version. This computes for each subject whether they have a different proportion correct for L and R

```{r domakeZ}
#We add Chimeric faces LIs to allsum for day 1 and then day2
task<-'Chimeric faces'
origdat<-chim_dat1
origdat$X1 <- origdat$N
mycolname<-"Day1_CF_acc_Z"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)

origdat<-chim_dat2
origdat$X1 <- origdat$N
mycolname<-"Day2_CF_acc_Z"
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)
w<-which(allsum$exclude == 1)
allsum$Day1_CF_acc_Z[w] <- NA #removes zscores for excluded cases
allsum$Day2_CF_acc_Z[w] <- NA

RR<-intersect(which(allsum$Day1_CF_acc_Z>1.96),which(allsum$Day2_CF_acc_Z>1.96))
LL<-intersect(which(allsum$Day1_CF_acc_Z< -1.96),which(allsum$Day2_CF_acc_Z< -1.96))

cor1 <- cor(allsum$Day1_CF_acc_Z,allsum$Day2_CF_acc_Z,use='complete.obs',method='spearman')
plot(allsum$Day1_CF_acc_Z,allsum$Day2_CF_acc_Z,main='Chimeric Z accuracy')
abline(v=0)
abline(h=0)
abline(v = 1.96, col='red')
abline(v = -1.96, col='red')
abline(h = 1.96, col='red')
abline(h = -1.96, col='red')
text(-7,7.5,paste0('Spearman r = ',round(cor1,3)),cex=.8)
text(-2,-15,paste('Those in bottom quadrant\n bounded by red are consistently\n L-lateralised, N = ',length(LL)),cex=.8)
text(5,18,paste('Those in top quadrant\n bounded by red are consistently\n R-lateralised, N = ',length(RR)),cex=.8)
#text(8,-8,'One outlier excluded',cex=.8)
#Removal of the outlier does not have much influence on rs

#Can also do sanity check: the Chimeric Z accuracy should be very highly correlated with the laterality index, but on a different scale

plot(allsum$Day1_CF_acc_LI,allsum$Day1_CF_acc_Z,main='Day 1: LI vs zlat')
plot(allsum$Day2_CF_acc_LI,allsum$Day2_CF_acc_Z,main='Da7 2: LI vs zlat')
```

As we can see, we generally get a left side/right hemisphere advantage in our accuracy data. We would not expect a RT advantage for this type of measure, where the response is to a single stimulus with conflicting information from the 2 sides. Nevertheless, there is a RT advantage - this is consistent with Bourne:Laterality 2008 Jan;13(1):92-103. doi: 10.1080/13576500701754315. (just read abstract so far).

## Does CF laterality relate to handedness?
Chunk below indicates there is a significant difference, though plots show much overlap.

```{r CFhanded}

t.test(allsum$Day1_CF_acc_Z~allsum$handpch)
t.test(allsum$Day2_CF_acc_Z~allsum$handpch)
beeswarm(Day1_CF_acc_Z~handedness, data = allsum, 
  pch = 16, cex=.5,col = rainbow(8),
  main = 'Day 1')
beeswarm(Day2_CF_acc_Z~handedness, data = allsum, 
  pch = 16, cex=.5,col = rainbow(8),
  main = 'Day 2')
```
## Next Question: Does CF laterality relate to task accuracy
Just looks at correlations and plot scattergram.

The scattergram shapes are a bit unusual - not sure if this is just because high levels of accuracy are more common. Does suggest high levels of lateralisation in either direction may be more accurate?
```{r CFlat_acc}
cor1 <- round(cor(allsum$Day1_CF_acc_Z,allsum$Emot.corr.diff1,use="complete.obs"),3)
plot(allsum$Day1_CF_acc_Z,allsum$Emot.corr.diff1)
text(15,.6,paste0('r = ',cor1))
cor2 <- round(cor(allsum$Day2_CF_acc_Z,allsum$Emot.corr.diff2,use="complete.obs"),3)
plot(allsum$Day2_CF_acc_Z,allsum$Emot.corr.diff2)
text(15,.6,paste0('r = ',cor2))

#These are correlations with unsigned z-scores: significant in large sample, but not much going on here
cor1a <- round(cor(abs(allsum$Day1_CF_acc_Z),allsum$Emot.corr.diff1,use="complete.obs"),3)
cor2a <- round(cor(abs(allsum$Day2_CF_acc_Z),allsum$Emot.corr.diff1,use="complete.obs"),3)
```

## Bergen Dichotic listening

This part of the notebook prcoesses the "Dichotic listening task' implmented online via Gorilla.sc which assesses the lateralisation of receptive language. In the task, participants hear a different CV syllable in each ear. Participants report the syllable they heard clearest

This task is near identical to that reported in Karlsson, Johnstone, and Carey (2019) and was original described in Hugdahl, Westerhausen, Alho, Medvedev, Laine, & Hämäläinen (2009). Stimuli were kindly provided by Dr Emma Karlsson (https://research.bangor.ac.uk/portal/en/researchers/emma-karlsson(17f6f06c-8aa9-4d16-93ba-4f98f285297e).html) and permission to use these stimuli was given by Professor Kenneth Hugdahl (https://www.uib.no/en/persons/Kenneth.Hugdahl).

Much of the processing parallels that of chimeric faces.

```{r read_dichotic_dat, warning=FALSE}
for (d in 1:2){ #we will process day 1 and day 2 together
  origfile <- all_dat1
  if (d==2)
  {origfile <- all_dat2}
# lowercase for responses
origfile$left_channel <- tolower(origfile$left_channel)
origfile$right_channel <- tolower(origfile$right_channel)
# make meaningful stimuli name
origfile$stimuli <- paste0(origfile$left_channel, "-", origfile$right_channel)
#As with earlier reading in, we default to saving dichotic_dat2 (day2), but at end of the chunk we reassign to CF_day1 in the first run through the d loop
task <- "Dichotic Listening"
wanted <- c("subject", "stimuli", "Response", "RT", "left_channel", "right_channel") # select wanted columns
disp <- "e_task"
dichotic_all <- procdata(origfile,wanted,task,disp) #use generic function (see above) to read in relevant columns

dichotic_all$Correct <- 0 #default is error
w<-union(which(dichotic_all$Response==dichotic_all$left_channel),which(dichotic_all$Response==dichotic_all$right_channel))
dichotic_all$Correct[w] <-1
print(paste0('Day ',d,': total Error (0) and Correct(1)'))
print(table(dichotic_all$Correct))
# count subjects
nsub <-length(unique(dichotic_all$subject))
print(paste('day',d,task,": subjects: ", nsub))

dichotic_same2 <- dichotic_all[dichotic_all$left_channel == dichotic_all$right_channel,] # create data frame for trials with both ears same
dichotic_dat2 <- dichotic_all[dichotic_all$left_channel != dichotic_all$right_channel,] # remove items where same expression in both hemifaces
if (d==1)
{dichotic_dat1 <- dichotic_dat2
 dichotic_same1 <- dichotic_same2}
}
```

Now remove outliers, using same function as for CF.
```{r dichotic.outliers}
dichotic_dat1 <- detect.outliers(dichotic_dat1,200,1.65) #use generic outlier removal function
outliertable1<-table(dichotic_dat1$subject,dichotic_dat1$outlier)
dichotic_dat2 <- detect.outliers(dichotic_dat2,200,1.65) #use generic outlier removal function
outliertable2<-table(dichotic_dat2$subject,dichotic_dat2$outlier)
#print(outliertable)

```
Inspection of outliertable1 and outliertable2 shows most are data as expected, with 120 trials. Outliers not v numerous. Occasional cases where test does not seem to be done, just one trial recorded.

Now, we judge participants' ability to correctly identify sounds using the "dichotic_same" dataframe. Participants are marked as excluded if their performance is below 75%.

```{r soundsability, warning=FALSE,message=FALSE}
# We consider how many errors there are when both ears have same sound. 
# NB we loop through same process for day1 and day2
for (d in 1:2){
  dichotic_same <- dichotic_same1
   dichotic_diff<- dichotic_dat1
  if (d==2){
    dichotic_same <- dichotic_same2
     dichotic_diff <- dichotic_dat2
  }

  # calculate participants average
  sound_mean <- aggregate(FUN= mean, data= dichotic_same, Correct~ subject)
  dichotic_acc <- aggregate(FUN= mean, data= dichotic_diff, Correct~ subject)
  # add this to allsum

  # Check if sound.recog columns already exist (ie if this chunk was run previously)
  w <- which(colnames(allsum) %in% c('sound.recog1','sound.recog2','dich_acc1','dich_acc2'))
  if (length(w)>0){
    allsum<-allsum[,-w]
  }

  allsum <- merge(allsum, sound_mean, by= "subject",all.x=T)
  allsum <- merge(allsum, dichotic_acc, by= "subject",all.x=T)
}

mycol <- length(colnames(allsum))
colnames(allsum)[(mycol-3):mycol]<-c('sound.recog1','sound.recog2','dichotic.acc1','dichotic.acc2')

# once again, mark rather than remove those with poor performance on identical stimuli
w<-union(which(allsum$sound.recog1 < .75), which(allsum$sound.recog2 < .75))
ww<-intersect(which(allsum$exclude==1),w)
#ww contains people who were excluded on both dichotic and faces. 
# code exclude = 2 for the cases where there less than 75% for identical stimuli
allsum$exclude[w]<-2 #code exclude = 2 if excluded because poor dichotic
allsum$exclude[ww]<-12 #code for those excluded on both chimeric and dichotic
t <- table(allsum$exclude) 
rownames(t)<-c('include','exclude CF','exclude dich','exclude CF+dich')
t
```

In this section, the correct sequences are marked as 1, incorrect asnwers are marked as 0. We then mark whether the correct answers were in the left or right ear.

```{r mark_dich_correct_ear, warning=FALSE}
# this will mark correct responses as 1 or 0.
# NB we again loop through same process for day1 and day2
for (d in 1:2){
  dichotic_dat <- dichotic_dat1
  if (d==2){
    dichotic_dat <- dichotic_dat2
  }

# this will mark the side responded to. NB to allow for generic functions for all tasks, we use the label 'side' rather than hemiface
dichotic_dat <- 
  dichotic_dat %>% 
  mutate(side= ifelse(as.character(Response) == as.character(left_channel) & Correct == 1, "Left",
                       ifelse(as.character(Response) == as.character(right_channel) & Correct == 1, "Right", NA)))
# code side as factor
dichotic_dat$side <- as.factor(dichotic_dat$side)

  if(d==1){
    dichotic_dat1 <- dichotic_dat}
  if (d==2){
    dichotic_dat2 <- dichotic_dat
  }
}
```

Now the data is ready, create an empty data frame and populate this. It includes columns for the participant identifier, side, and count of accurate responses for each side.

```{r makedataframedich, warning=F}
#We create DL_dat files for day1 and day 2
varlist <-   c("subject", "side", "accurate", "RT") #need to be in this order, ie sub, side, acc and RR
origdat <- dichotic_dat1
latlist <- c("Left","Right") #names of factor levels
nloop=2 #same as levels of latlist
DL_dat1 <- make.df(dichotic_dat1,varlist,latlist,nloop)
DL_dat2 <- make.df(dichotic_dat2,varlist,latlist,nloop)

w<-which(is.na(DL_dat1$subject))
if (length(w)>0) {DL_dat1<-DL_dat1[-w,]}
w<-which(is.na(DL_dat2$subject))
if (length(w)>0) {DL_dat2<-DL_dat2[-w,]}

DL_dat1[is.na(DL_dat1)] <- 0
DL_dat2[is.na(DL_dat2)] <- 0
```

The data is visualised using the pirate plot function and t.test conducted.

```{r dosummarydich}
includesubs <- which(allsum$exclude<2) #only plot/t-test included cases
origdat <- DL_dat1[DL_dat1$subject %in% includesubs,]
measure <- 'Accuracy'
origdat$X1 <- origdat$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Dichotic listening day 1'
dopirate(origdat,task,measure)
origdat$X1 <- origdat$RT
measure <- 'RT'
dopirate(origdat,task,measure)
t.test(origdat$accurate~ origdat$side)
#Now repeat for day 2
origdat <-  DL_dat2[DL_dat2$subject %in% includesubs,]
origdat$X1 <- origdat$accurate #we used a dummy column name so we can vary the data that are used in the pirate plot etc
task <- 'Dichotic listening day 2'
measure <- 'Accuracy'
dopirate(origdat,task,measure)
origdat$X1 <- origdat$RT
measure <- 'RT'
dopirate(origdat,task,measure)
```

Finally, a laterality index is calculated and plotted based on the equation [(R-L)/(R+L)] x 100, where L= left side and R= right side.

If the index is negative, speech recognition is lateralised in the left hemisphere. If the index is positive, speech recognition is lateralised in the right hemisphere

Here, a .csv is written for the dichotic LI.

```{r domakeLidich}
#We add Dichotic LIs to allsum for day 1 and then day2
task<-'Dichotic listening'
origdat<-DL_dat1
origdat$X1 <- origdat$accurate
mycolname<-"Day1_DL_acc_LI"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

# now do RT
origdat$X1 <- origdat$RT
mycolname<-"Day1_DL_RT_LI"
polarity <-1

#if we don't specify mycutoff, it will remain at level set for accuracy, which means it should not trap any RT values as they will all be much bigger
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

origdat<-DL_dat2
origdat$X1 <- origdat$accurate
mycolname<-"Day2_DL_acc_LI"
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
origdat$X1 <- origdat$RT
mycolname<-"Day2_DL_RT_LI"
polarity <-1
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

w<-which(allsum$exclude==2) #substitute NA for those who were excluded
ncol<-length(colnames(allsum))
allsum[w,(ncol-3):ncol]<-NA
cor1 <- cor(allsum$Day1_DL_acc_LI,allsum$Day2_DL_acc_LI,use='complete.obs',method='spearman')
plot(allsum$Day1_DL_acc_LI,allsum$Day2_DL_acc_LI,main='Dichotic LI accuracy',col=(1+allsum$exclude))
abline(v=0)
abline(h=0)
text(-75,75,paste0('r = ',round(cor1,3)))
cor2<-cor(allsum$Day1_DL_RT_LI,allsum$Day2_DL_RT_LI,use='complete.obs',method='spearman')
plot(allsum$Day1_DL_RT_LI,allsum$Day2_DL_RT_LI,main='Dichotic LI RT',col=(1+allsum$exclude))
abline(v=0)
abline(h=0)
text(-5,15,paste0('r = ',round(cor2,3)))

#Plot of LI shows that two participants reported only from one ear (acc LI = 100).
#Although we did not pre-register this, it would make sense to exclude these, who are clear outliers, doing the task quite differently from others.

w<-which(allsum$Day1_DL_RT_LI== (-100))
allsum$Day1_DL_RT_LI[w] <- NA
allsum$Day1_DL_acc_LI[w] <- NA
allsum$Day2_DL_RT_LI[w] <- NA
allsum$Day2_DL_acc_LI[w] <- NA

allsum$exclude[w]<-2 #mark as excluded for dichotic

#Redo the correlation and plots
cor1 <- cor(allsum$Day1_DL_acc_LI,allsum$Day2_DL_acc_LI,use='complete.obs',method='spearman')
plot(allsum$Day1_DL_acc_LI,allsum$Day2_DL_acc_LI,main='Dichotic LI accuracy, after 2 exclusions',col=(1+allsum$exclude))
abline(v=0)
abline(h=0)
text(-75,75,paste0('r = ',round(cor1,3)))
cor2<-cor(allsum$Day1_DL_RT_LI,allsum$Day2_DL_RT_LI,use='complete.obs',method='spearman')
plot(allsum$Day1_DL_RT_LI,allsum$Day2_DL_RT_LI,main='Dichotic LI RT, after 2 exclusions',col=(1+allsum$exclude))
abline(v=0)
abline(h=0)
text(-5,15,paste0('r = ',round(cor2,3)))

```

Here we use the Z score version. This is an individualised score that directly indicates whether or not the person is reliably lateralised on one test occasion.

```{r domakeZ_dich}
#We add dichotic LIs to allsum for day 1 and then day2
task<-'Dichotic listening'
origdat<-DL_dat1
origdat$X1 <- origdat$N
mycolname<-"Day1_Dich_acc_Z"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)

origdat<-DL_dat2
origdat$X1 <- origdat$N
mycolname<-"Day2_Dich_acc_Z"
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)

#convert excluded cases with code 2 or 12 to NA
w<-which(allsum$exclude>1)
allsum$Day1_Dich_acc_Z[w]<-NA
allsum$Day2_Dich_acc_Z[w]<-NA

RR<-intersect(which(allsum$Day1_Dich_acc_Z>1.96),which(allsum$Day2_Dich_acc_Z>1.96))
LL<-intersect(which(allsum$Day1_Dich_acc_Z< -1.96),which(allsum$Day2_Dich_acc_Z< -1.96))

cor1 <- cor(allsum$Day1_Dich_acc_Z,allsum$Day2_Dich_acc_Z,use='complete.obs',method='spearman')

plot(allsum$Day1_Dich_acc_Z,allsum$Day2_Dich_acc_Z,main='Dich Z accuracy',pch=allsum$handpch)
abline(v=0)
abline(h=0)
text(-7.5,15,paste0('Spearman r = ',round(cor1,3)),cex=.8)
abline(v = 1.96, col='red')
abline(v = -1.96, col='red')
abline(h = 1.96, col='red')
abline(h = -1.96, col='red')

text(20,-15,paste('Those in bottom quadrant\n bounded by red are consistently\n L-lateralised, N = ',length(LL)),cex=.8)
text(5,40,paste('Those in top quadrant\n bounded by red are consistently\n R-lateralised, N = ',length(RR)),cex=.8)

```
## Does dichotic laterality relate to handedness?
Chunk below indicates there is a significant difference, but v small effect size.

```{r dichotic_handed}
partsum <- allsum[allsum$subject %in% includesubs,]

t.test(partsum$Day1_Dich_acc_Z~partsum$handpch)
t.test(partsum$Day2_Dich_acc_Z~partsum$handpch)
beeswarm(Day1_Dich_acc_Z~handedness, data = partsum, 
  pch = 16, cex=.5,col = rainbow(8),
  main = 'Day 1')
beeswarm(Day2_Dich_acc_Z~handedness, data = partsum, 
  pch = 16, cex=.5,col = rainbow(8),
  main = 'Day 2')
```
## Next Question: Does Dichotic laterality relate to task accuracy
Just looks at correlations and plot scattergram.
No hint of any relationship.

```{r Cdichotic_acc}
cor1 <- round(cor(partsum$Day1_Dich_acc_Z,partsum$dichotic.acc1,use="complete.obs"),3)
plot(partsum$Day1_Dich_acc_Z,partsum$dichotic.acc1)
text(15,.6,paste0('r = ',cor1))
cor2 <- round(cor(partsum$Day2_Dich_acc_Z,partsum$dichotic.acc2,use="complete.obs"),3)
plot(partsum$Day2_Dich_acc_Z,partsum$dichotic.acc2)
text(15,.6,paste0('r = ',cor2))

```

## Rhyme Detection Task

This portion of the notebook prcoesses the "Rhyme Detection Task' implemented online via Gorilla.sc, which aims to assess the lateralisation of language processing. This is a new task devised for this project. In the task, participants are shown a word in the central visual field and two images in the left and right visual field. The task is to judge which of the two images has a name that rhymes with the central word.

 We start by reading the data and cutting out unwanted information. Here we are including item-specific information as it may be used for later analysis.

```{r read_dat_RDT, warning=FALSE}
for (d in 1:2){ #same script for days 1 and 2

  origfile <- all_dat1
  if(d==2)
  {origfile <- all_dat2}
  task <- "Rhyme Detection Task"
  disp <- "Task"

  # nb when creating new file, we default to name 'RDT_dat2', but this is reassigned to 'RDT_dat1' at end of processing for day 1.

  # new data
  wanted <- c("Trial","subject", "Correct", "RT", "ANSWER","Response","word","Left_image","Right_image") # select wanted columns
  RDT_dat2 <- procdata(origfile,wanted,task,disp) #Here we reuse the procdata function that we developed for processing Chimeric faces data - just does some generic processing

  # do some wrangling on the ANSWER variable to make it compatible with functions
  RDT_dat2<-filter(RDT_dat2,ANSWER %in% c('left','right','no rhyme'))
  c <- which(colnames(RDT_dat2)=="ANSWER")
  colnames(RDT_dat2)[c]<-'side'
  RDT_dat2$side<-as.factor(RDT_dat2$side)
  levels(RDT_dat2$side)<-c("left","no rhyme","right")

  #get rid of .png extensions
  RDT_dat2$Left_image<-tools::file_path_sans_ext(RDT_dat2$Left_image)
   RDT_dat2$Right_image<-tools::file_path_sans_ext(RDT_dat2$Right_image)

   RDT_dat2$picpair <- paste0(RDT_dat2$Left_image,'_',RDT_dat2$Right_image)
   w<-which(RDT_dat2$side=='right')
   RDT_dat2$picpair[w]<- paste0(RDT_dat2$Right_image[w],'_',RDT_dat2$Left_image[w])

   RDT_dat2$item <- paste0(RDT_dat2$word,'_',RDT_dat2$picpair) 
   #Check all OK
   t<-table(RDT_dat2$item,RDT_dat2$side)
   t
   write.csv(t,'items.csv')

   #Reveals one item miscoded, so we will correct it here.
   w<-which(RDT_dat2$item=='bite_Door_Kite')
   temp<-RDT_dat2[w,]
   #This shows that 'no response' was treated as correct, when it should have been 'right'

   w2 <- intersect(w,which(RDT_dat2$Response=='right'))
   RDT_dat2$Correct[w2]<-1
   RDT_dat2$item[w2]<-'bite_Kite_Door'
   RDT_dat2$side[w2]<-'right'

   #Also reveals occasional trials where Response shows 'Could not set screen ...' These were coded as errors, so need removing
   s <- "Could not set image to requested size.  Either the participants screen or the containing zone is too small."
   w<-which (RDT_dat2$Response==s)
   RDT_dat2<-RDT_dat2[-w,]

  # count subjects
nsub <-length(unique(RDT_dat2$subject))
print(paste0("Day ",d," subjects: ", nsub))
if(d==1)
  {RDT_dat1 <- RDT_dat2}
}
```
Now use generic function to remove outliers.

```{r rdt_outliers}
RDT_dat1<-detect.outliers(RDT_dat1,200,1.65) #numbers specify min RT and 
RDT_dat2<-detect.outliers(RDT_dat2,200,1.65) #numbers specify min RT and zscore for Hoaglin-Iglewicz respectively

outliertable1<-table(RDT_dat1$subject,RDT_dat1$outlier)
outliertable2<-table(RDT_dat2$subject,RDT_dat2$outlier)

```

Scrutiny of outliertable indicates a few problematic cases with far fewer than 260 trials in total. Seems related to glitches in Gorilla.
Will need to exclude these

Now the data is ready, create an empty data frame and populate this. It includes columns for the participant identifier, rhyme, acuracy, RT (RT only for correct answers).

```{r makedf_rhyme}
#We create chim_dat files for day1 and day 2
varlist <-   c("subject", "side", "accurate", "RT") #need to be in this order, ie sub, side, acc and RR
latlist <- c("left","no rhyme","right") #names of factor levels
nloop<-length(latlist)
RDT_df1 <- make.df(RDT_dat1,varlist,latlist,nloop)
RDT_df2 <- make.df(RDT_dat2,varlist,latlist,nloop)

#Scrutiny of these reveal a few cases who missed all 'no rhyme' cases.
#Qu of whether to remove these: think so - this was included to ensure thoughtful responding.
```

Next exclude those who never get 'no rhyme' items correct, and those with less than 75% correct responses on either side

```{r RDT_exclude}
wn <-unique(which(RDT_df1$accurate ==0),which(RDT_df2$accurate ==0)) #this catches those who were always wrong on 'no rhyme' trials, and possibly others where presentation messed up for technical reasons

w1<-which(RDT_df1$p.corr < 75)
w2<-which(RDT_df2$p.corr < 75)

exRDT<-unique(c(RDT_df1$subject[w1],RDT_df2$subject[w2],RDT_df2$subject[wn]))

allsum$exRDT <- 0
allsum$exRDT[allsum$subject %in% exRDT]<-1 #decided best to have separate column for exclusion for each task
w<-which(allsum$exRDT==0)  
incRDT <- allsum$subject[w]
ninc <- length(incRDT)
print(paste0('RDT: number remaining after exclusions = ',ninc))
```

Comparison shows that there is significant difference in RT. Accuracy is close to ceiling and does not show side difference (but it would be a bit odd for this task, - unlike CF and dichotic, there is no competing information and only one answer is correct).
```{r rdtpirate}
for (d in 1:2){
origdat <-RDT_df1[RDT_df1$subject %in% incRDT,]
if (d==2)
{origdat <-RDT_df2[RDT_df2$subject %in% incRDT,]}
origdat <-origdat[origdat$side %in% c('left','right'),]
measure <- 'Accuracy'
origdat$X1 <- origdat$p.corr
task <- 'Rhyme Detection'

dopirate(origdat,task,measure)
origdat$X1 <- origdat$RT
measure <- 'RT'
dopirate(origdat,task,measure)
}
```

```{r RDT.LI}
#now create LI
for (d in 1:2){
origdat <-RDT_df1[RDT_df1$subject %in% incRDT,]
if (d==2)
{origdat <-RDT_df2[RDT_df2$subject %in% incRDT,]}
origdat$X1 <- origdat$p.corr
origdat$side <- as.factor(origdat$side)
levels(origdat$side) <-c('Left','No rhyme','Right') #function assumes capitalised
mycolname<-"RDT_p.corr_LI"
polarity<- 1 #so LI is positive in predicted direction
mycutoff <- 75
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
origdat$X1 <- origdat$RT
mycolname<-"RDT_RT_LI"
polarity <- (-1)
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

#Cutoff has removed those with low accuracy, but we should also remove their RT LIs
#We do this by finding those with NA in accuracy LI
w<-which(is.na(allsum$RDT_acc_LI))
if(length(w)>0)
{allsum$RDT_RT_LI[w]<-NA}
#t.test to test if laterality index differs from zero
partsum <- allsum[allsum$exRDT==0,]
print('Testing for difference of single mean from zero')
t.test(partsum$Day1.RDT_p.corr_LI)
t.test(partsum$Day1.RDT_RT_LI)
t.test(partsum$Day2.RDT_p.corr_LI)
t.test(partsum$Day2.RDT_RT_LI)
#rename column to identify day
myncol <- length(colnames(allsum))
colnames(allsum)[(myncol-1):myncol] <- paste0('Day',d,'.',colnames(allsum)[(myncol-1):myncol])

}
cor1<-cor(partsum$Day1.RDT_p.corr_LI,partsum$Day2.RDT_p.corr_LI,use='complete.obs')
cor2<-cor(partsum$Day1.RDT_RT_LI,partsum$Day2.RDT_RT_LI,use='complete.obs')
plot(partsum$Day1.RDT_p.corr_LI,partsum$Day2.RDT_p.corr_LI,main='Rhyme detect: LI for accuracy (%correct)')
abline(v=0)
abline(h=0)
text(-3,3,paste0('r = ',round(cor1,3)))
plot(partsum$Day1.RDT_RT_LI,partsum$Day2.RDT_RT_LI,main='Rhyme detect: LI for RT')
abline(v=0)
abline(h=0)
text(-3,9,paste0('r = ',round(cor2,3)))
```
## Does RDT relate to handedness?
Answer is convincing nope!

```{r RDThand}
t.test(Day1.RDT_RT_LI~handpch,data=partsum)
t.test(Day2.RDT_RT_LI~handpch,data=partsum)
```
### T-test as alternative to LI for RDT

The LI results are not encouraging - reliability well below acceptable level, and although RT shows lateralised bias, it is tiny. Also, many subjects dropped.
Previously I also did a t-test that was analogous to the z-test for CF/dichotic, i.e. showed if the individual had a significant RT difference between L and R stimuli. But that is v highly correlated with LI. The only way to make it more sensitive would be to try a pairwise t-test that compares RTs for the same items presented on L and R. So the next bit of code is designed to compute that.

```{r RDTaggregate}
#remove 'no rhyme' trials
for (d in 1:2){
  thisdat<-RDT_dat2[RDT_dat2$side %in% c('left','right'),]
  if(d==1){
    thisdat<-RDT_dat1[RDT_dat1$side %in% c('left','right'),]
  }

  thisdat <- thisdat[thisdat$outlier==0,] #remove outliers
  aggdata <- aggregate(thisdat$RT,by=list(thisdat$side,thisdat$item,thisdat$subject),
                       FUN=mean,na.rm=TRUE)
  colnames(aggdata)<-c('side','item','subject','RT')
  mysubs<-unique(aggdata$subject)
  if(d==1){
  pairedt <- data.frame(matrix(NA,nrow=length(mysubs),ncol=5))
  colnames(pairedt) <- c('subject','t1','p1','t2','p2')
  }
  nsubs<-length(mysubs)
  for (i in 1:nsubs){
    myt<-NA
    myp<-NA
    thisdat <- aggdata[aggdata$subject==mysubs[i],]
    pairedt$subject[i]<- mysubs[i]
    #check there are data for both left and right on all items
    checkdat <-table(thisdat$item)
    w<-which(checkdat<2)
    if(length(w)==0){
      t<-t.test(thisdat$RT[thisdat$side=='left'],thisdat$RT[thisdat$side=='right'],paired=T)
      myt<-t$statistic
      myp<-t$p.value
    }
    if(d==1){
      pairedt$t1[i] <- myt
      pairedt$p1[i] <-myp
    }
    if(d==2){
      pairedt$t2[i] <-myt
      pairedt$p2[i] <-myp
    }
  }
}
cor(pairedt$t1,pairedt$t2,use='complete.obs')
```

So the paired items analysis gives a totally nonsignificant correlation between time 1 and time 2. Alas, the RDT task has to be ditched!

## Finger Tapping Task

This portion of the R Markdown file prcoesses the "Finger Tapping Task' implmented in Gorilla.sc which aims to quantify handedness. In the task, participants are required to press a sequence of buttons twice for each hand (Left: W, R, V, X; Right: T, U, M, B) as many times as possible within a 30 ms. There is a version that requires more precise movements.

A score of 1 is given each time that a participants completes the sequence.

This reads in a csv file where each row represents a button press. Participants, identified by random letter strings, are stacked on top of each other. We start by reading the data and cutting out unwanted information.

NOTE. The create data frame function disrupts order here.

In this section, the correct sequences are marked (e.g. pressing Y, U, J, H, in order will correspond to a score of 1).

```{r readdataFinger, warning=FALSE}

finger <- read.csv("day1_combined.csv", na.strings = "NA")

# cleaning
finger <- finger[finger$Attempt >= 1,] # only answers 
finger <- finger[finger$display == "right hand" | finger$display == "left hand",] # only for task
# write as factors
finger$display <- as.factor(finger$display) 
finger$subject <- as.factor(finger$Participant.Public.ID)
finger$RT <- as.numeric(as.character(finger$Reaction.Time))

# new data
wanted <- c("subject", "display", "Response", "RT") # select wanted columns
c<-which(names(finger) %in% wanted) #find colnumbers of unwanted
finger <-finger[,c] #remove unwanted columns
finger <- na.omit(finger) # remove NAs

all_finger1<- finger

finger <- read.csv("day2_combined.csv", na.strings = "NA")

# cleaning
finger <- finger[finger$Attempt >= 1,] # only answers 
finger <- finger[finger$display == "right hand" | finger$display == "left hand",] # only for task
# write as factors
finger$display <- as.factor(finger$display) 
finger$subject <- as.factor(finger$Participant.Public.ID)
finger$RT <- as.numeric(as.character(finger$Reaction.Time))

# new data
wanted <- c("subject", "display", "Response", "RT") # select wanted columns
c<-which(names(finger) %in% wanted) #find colnumbers of unwanted
finger <-finger[,c] #remove unwanted columns
finger <- na.omit(finger) # remove NAs

all_finger2<- finger

```

```{r readdataFinger2, warning=FALSE}
for (d in 1:2){ #we will process day 1 and day 2 together
  origfile <- all_finger1
  if (d==2)
  {origfile <- all_finger2}
#As with earlier reading in, we default to saving finger2 (day2), but at end of the chunk we reassign to finger1 in the first run through the d loop
finger2 <- origfile

finger2$RTbase[2:nrow(finger2)]<-finger2$RT[1:(nrow(finger2)-1)]
finger2$RTa <- finger2$RT-finger2$RTbase
#remove negative or v long RTs - can indicate when block changes etc
w<-c(which(finger2$RTa<0),which(finger2$RTa>1000))
finger2$RTa[w]<-NA
if (d==1)
{finger1 <- finger2
 }
}
```

In this section, the correct sequences are marked (e.g. pressing Y, M, I, B, in order will correspond to a score of 1).

Added exploratory analysis of SDs of RTs for keypress rather than means, but this was not particularly illuminating.

```{r scorefinger}
# We can go directly to creation of a data.frame by subjects, and use string matching to identify N correct seqs

myseq <-c('tumb','wrvx') # R, # L

subids <- levels(finger1$subject) #just look at data for those doing time1 and
nsub <- length(subids)

finger.df <- data.frame(matrix(NA,nrow=nsub,ncol=9))
colnames(finger.df)<-c('subject', 'R1', 'R2', 'L1', 'L2', 'R1sd', 'R2sd', 'L1sd', 'L2sd')
for (n in 1:nsub){
  finger.df$subject[n] <- subids[n]
  myc <- 1 #will be incremented so sums written to correct columns
  for (d in 1:2){ #each day data
    myfinger <- finger1
    if(d==2){
      myfinger<- finger2}
      tempfinger <- filter(myfinger,subject==subids[n],Response %in% c('t', 'u', 'm', 'b', 'w', 'r', 'v', 'x')) 
      #exclude rows with other stuff in Response
      #(this is not essential as nonsequences will just get ignored)

      wholeseq <-paste0(tempfinger$Response,collapse='') #make a long string of responses
      for (s in 1:2){
        myc<-myc+1 #increment column counter
        target <- myseq[s]
        finger.df[n,myc]<-str_count(wholeseq,target) 
        #count how many target sequences are in the long string that contains all responses

      }
      righth<-filter(tempfinger,display=='right hand')
      lefth<-filter(tempfinger,display=='left hand')
      startcol<-6
      if(d==2){startcol<-8}
      finger.df[n,startcol]<-sd(righth$RTa,na.rm=T)
      finger.df[n,(startcol+1)]<-sd(lefth$RTa,na.rm=T)
    }
}
# Substitute NA for any totals less than 5 - suggest using wrong keys perhaps

for (c in 2:5){
  w<-which(finger.df[,c]<5)
  if (length(w)>0){
     finger.df[w,2:9]<-NA

  }

}
#Need same data in long form for pirate plot, so reorganise
finger.left <- finger.df[,c(1,4,5,8,9)]
finger.right <- finger.df[,c(1,2,3,6,7)]
finger.left$side<-'Left'
finger.right$side <-'Right'
colnames(finger.right)<-c('subject', '1', '2', '1sd', '2sd','side')
colnames(finger.left)<-colnames(finger.right)
finger.long <-rbind(finger.left,finger.right)

```

```{r dosummary2, warning = F}
origdat <- finger.long
measure <- '1'
origdat$X1 <- origdat$`1` #we used a dummy column name X1, so we can vary the data that are used in the pirate plot etc
task <- 'day 1'
dopirate(origdat,task,measure)

origdat <- finger.long
measure <- '2'
origdat$X1 <- origdat$`2` #we used a dummy column name X1, so we can vary the data that are used in the pirate plot etc
task <- 'day 2'
dopirate(origdat,task,measure)

origdat <- finger.long
measure <- '1sd'
origdat$X1 <- origdat$`1sd` #we used a dummy column name X1, so we can vary the data that are used in the pirate plot etc
task <- 'day 1sd'
dopirate(origdat,task,measure)

origdat <- finger.long
measure <- '2sd'
origdat$X1 <- origdat$`2sd` #we used a dummy column name X1, so we can vary the data that are used in the pirate plot etc
task <- 'day 2sd'
dopirate(origdat,task,measure)
```

Finally, a handedness index is calculated and plotted. The handedness index is based on the equation [(RH-LH)/(RH+LH)] x 100, where LH= left hand and RH= right hand.

If the index is negative, the participant is categorised as left handed. If the index is positive, the participant is categorised as right handed.

In the next chunk, the LI is also added to allsum

```{r fingerLI}
task<-'Finger'
origdat<-finger.long
origdat$X1 <- origdat$`1`
mycolname<-"Day1_finger_LI"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (1)#so LI is positive in predicted direction
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

origdat$X1 <- origdat$`2`
mycolname<-"Day2_finger_LI"
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

cor1 <- cor(allsum$Day1_finger_LI,allsum$Day2_finger_LI,use='complete.obs')
plot(allsum$Day1_finger_LI,allsum$Day2_finger_LI,main='Finger Tapping LI')
abline(v=0)
abline(h=0)
text(-20,10,paste0('r = ',round(cor1,3)))
```

Here we use the Z score version.

```{r domakeZ_finger}
#We add Chimeric faces LIs to allsum for day 1 and then day2
task<-'Finger'
origdat<-finger.long
origdat$X1 <- origdat$`1`
mycolname<-"Day1_Finger_acc_Z"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (1)#so LI is positive in predicted direction
allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)

origdat<-finger.long
origdat$X1 <- origdat$`2`
mycolname<-"Day2_Finger_acc_Z"
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)

cor1 <- cor(allsum$Day1_Finger_acc_Z,allsum$Day2_Finger_acc_Z)
plot(allsum$Day1_Finger_acc_Z,allsum$Day2_Finger_acc_Z,main='Finger Z accuracy')
abline(v=0)
abline(h=0)
text(-2.5,2.5,paste0('r = ',round(cor1,3)))

write.csv(allsum, "LIs.csv")
```

## Final sample summary

Here we report the descriptives for the participants who had full datasets.

```{r Finalallsum, warning=FALSE}
# first, describe age for the whole sample and plot
allsum %>%
    summarize(avg = mean(age), n = n(), sd = sd(age))
# second, show gender count for the whole population
allsum %>%
  group_by(gender) %>%
    summarize(n = n())
# third, show footedness
allsum %>%
  group_by(footedness) %>%
    summarize(n = n())
# fourth, show handedness
allsum %>%
  group_by(handedness) %>%
    summarize(n = n())
# fifth, show bilingualism
allsum %>%
  group_by(bilingual) %>%
    summarize(n = n())
# sixth, show miles
allsum %>%
  group_by(miles) %>%
    summarize(n = n())
# seventh, show porta
allsum %>%
  group_by(porta) %>%
    summarize(n = n())
# eigth, show LexTALE
allsum %>%
    summarize(avg = mean(lexTALE), n = n(), sd = sd(lexTALE))

# Now that we have the basic scores, let's look at how this looks with regards to handedness
# first, describe age for the whole sample and plot
ggplot(allsum, aes(x = handedness, y = age)) +
  geom_pirate(aes(colour = handedness), bars = FALSE) + theme_classic() + 
  ylim(10, 60) + ggtitle("Age by handedness") + ylab("Age (years)") + xlab("")
# second, show gender count for the whole population
a <- allsum %>%
  group_by(gender, handedness) %>%
    summarize(n = n())
ggplot(data=a, aes(x=handedness, y=n, fill= handedness)) +
  geom_bar(stat="identity")+
  geom_text(aes(label=n), vjust=1.6, color="white", size=3.5)+
  theme_classic() + facet_wrap(.~ gender) + ggtitle("Gender by handedness") + 
  ylab("Count") + xlab("")
# third, show footedness
a <- allsum %>%
  group_by(handedness, footedness) %>%
    summarize(n = n())
ggplot(data=a, aes(x=handedness, y=n, fill= handedness)) +
  geom_bar(stat="identity")+
  geom_text(aes(label=n), vjust=1.6, color="white", size=3.5)+
  theme_classic() + facet_wrap(.~ footedness) + ggtitle("Footedness by handedness") + 
  ylab("Count") + xlab("")
# fourth, show bilingualism
a <- allsum %>%
  group_by(bilingual, handedness) %>%
    summarize(n = n())
ggplot(data=a, aes(x=handedness, y=n, fill= handedness)) +
  geom_bar(stat="identity")+
  geom_text(aes(label=n), vjust=1.6, color="white", size=3.5)+
  theme_classic() + facet_wrap(.~ bilingual) + ggtitle("Bilingualism by handedness") + 
  ylab("Count") + xlab("")
# fifth, show miles
a <- allsum %>%
  group_by(miles, handedness) %>%
    summarize(n = n())
ggplot(data=a, aes(x=handedness, y=n, fill= handedness)) +
  geom_bar(stat="identity")+
  geom_text(aes(label=n), vjust=1.6, color="white", size=3.5)+
  theme_classic() + facet_wrap(.~ miles) + ggtitle("Mile test by handedness") + 
  ylab("Count") + xlab("")
# six, show porta
a <- allsum %>%
  group_by(porta, handedness) %>%
    summarize(n = n())
ggplot(data=a, aes(x=handedness, y=n, fill= handedness)) +
  geom_bar(stat="identity")+
  geom_text(aes(label=n), vjust=1.6, color="white", size=3.5)+
  theme_classic() + facet_wrap(.~ porta) + ggtitle("Porta test by handedness") + 
  ylab("Count") + xlab("")
# seventh, show LexTALE
ggplot(allsum, aes(x = handedness, y = lexTALE)) +
  geom_pirate(aes(colour = handedness), bars = FALSE) + theme_classic() + 
  ylim(70, 110) + ggtitle("LexTALE by handedness") + ylab("Standardised score") + xlab("")
```

```{r corr_between_tasks}
mycols <- c('Day1_CF_acc_Z','Day2_CF_acc_Z','Day1_Dich_acc_Z','Day2_Dich_acc_Z','tlat_RDT_1','tlat_RDT_2','Day1_Finger_acc_Z','Day2_Finger_acc_Z')
colnums<- which(colnames(allsum) %in% mycols)
mycorrs<-correlate(allsum[,colnums],method='spearman')

mycorrsL <-correlate(allsum[allsum$handedness=='Left',colnums],method='spearman')
mycorrsR <-correlate(allsum[allsum$handedness=='Right',colnums],method='spearman')

#compare L and R handers

for (i in colnums){
  print(names(allsum)[i])
 print(t.test(allsum[,i]~allsum$handedness))
}

#check lateralisaiton with single group t vs zero
for (i in colnums){
  print(names(allsum)[i])
 print(t.test(allsum[,i]))
}

```

**References:**

- Burt, D. M., & Perrett, D. I. (1997). Perceptual asymmetries in judgements of facial attractiveness, age, gender, speech and expression. Neuropsychologia, 35(5), 685–693.
- Hoaglin, D. C., & Iglewicz, B. (1987). Fine-tuning some resistant rules for outlier labelling. Journal of the American Statistical Association, 82(400), 1147-1149. 
- Hugdahl, K., Westerhausen, R., Alho, K., Medvedev, S., Laine, M., & Hämäläinen, H. (2009). Attention and cognitive control: unfolding the dichotic listening story. Scandinavian journal of psychology, 50(1), 11-22.
- Innes, B. R., Burt, D. M., Birch, Y. K., & Hausmann, M. (2016). A leftward bias however you look at it: Revisiting the emotional chimeric face task as a tool for measuring emotion lateralization. Laterality: Asymmetries of Body, Brain and Cognition, 21(4-6), 643–661.
- Karlsson, E. M., Johnstone, L. T., & Carey, D. P. (2019). The depth and breadth of multiple perceptual asymmetries in right handers and non-right handers. Laterality: Asymmetries of Body, Brain and Cognition, 24(6), 707-739.