### 2020-07-30 online_laterality_reliability.Rmd
runs with knitr

---
title: 'Assessing reliability of an online behavioural laterality battery: a pre-registered
  study.'
output:
  word_document: default
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
---

<!--- updated 20/7/20 to change 'select' to 'dplyr::select' to avoid conflicts-->
Adam J. Parker

Zoe V. J. Woodhead

Paul A. Thompson

Dorothy V. M. Bishop

Department of Experimental Psychology, University of Oxford, Anna Watts Building, Radcliffe Observatory Quarter, Woodstock Road, Oxford, OX2 6GG.

<!--- not sure how corresponding authorship works in OSCCI lab --->
<!--- I'm fine with you in this role--->
*Correspondence regarding this article should be addressed to Adam J. Parker, Department of Experimental Psychology, University of Oxford, Anna Watts Building, Woodstock Road, Oxford, OX2 6GG. Email: adam.parker@psy.ox.ac.uk.

\newpage
## Abstract

Studies of cerebral lateralisation often involve participants completing a series of perceptual tasks under laboratory conditions. This has constrained the number of participants recruited in such studies. Online testing can potentially allow for much larger sample sizes, but limits the amount of experimental control that is feasible. Here we considered whether online testing could give valid and reliable results on four tasks: a rhyme decision visual half-field task, a dichotic listening task, a chimeric faces task, and a finger tapping task to assess manual preference. We recruited 392 participants, oversampling left-handers, who completed the battery twice. Three of the tasks showed evidence of both validity and reliability, insofar as they showed hemispheric advantages in the expected direction and test-retest reliability of at least .65. The rhyme judgement task was less satisfactory. We also found that left-handers were more likely to depart from typical lateralisation. Contrary to prediction, lateralisation across tasks was not significantly correlated. We conclude that three of the tasks, dichotic listening, chimeric faces and finger-tapping, show considerable promise for online evaluation of cerebral lateralisation, but the results do not support the notion of a common underlying cause of lateralisation for different cognitive tasks.

*Keywords*: cerebral lateralisation, hemispheric asymmetry, language, online behaviour research methods, reliability

\newpage
## Introduction

Hemispheric specialisation, or lateralisation, is a ubiquitous characteristic of the human brain, yet many questions remain unanswered as to its causes and consequences. Much of the ambiguity in the published literature may arise from unreliable measures of lateralisation and underpowered studies. Without well-powered experiments that use reliable tasks, it becomes difficult to address many research questions regarding the relationship between laterality indices and their stability over time. With respect to the relationship between indices, unless a task is reliable, the absence of a correlation cannot be interpreted as evidence for independence of function (Bishop, 2017). Reliability can either be assessed within a test session (e.g. split-half reliability, from correlations between odd and even items) or from repeated assessments on different occasions, test-retest reliability. The latter is generally preferred as it includes all test items and will incorporate fluctuations in performance due to changes in state or test conditiions from one occasion to another. Assessment of test-retest reliability, however, requires large samples of people who undergo repeated testing, and this is seldom feasible for in-person testing. Online testing allows for recruitment of much larger samples, but raises questions about test validity when opportunities for experimental control are reduced. Here we introduce an online behavioural laterality battery for which we established the test-retest reliability using a sample of 392 participants. The behavioural tasks included were: a novel rhyme decision task, a dichotic listening task (Hugdahl et al., 2009), a chimeric faces task (Burt & Perret, 1997), and a novel finger tapping task. Validity of the tasks was assessed in terms of whether lateralisation was observed in the predicted direction, based on prior literature with in-person testing on analogous tasks. We also examined two pre-registered questions: (1) do extreme left-handers show a higher rate of atypical laterality than right handers on each behavioural laterality task? (2) do the laterality indices from different tasks correlate?

### Behavioural laterality tasks and their reliability.

Behavioural laterality tasks have long been used to establish hemispheric dominance for certain cognitive functions. Rather consistently, results from visual half-field and dichotic listening paradigms support the view that the left hemisphere plays the dominant role in language processing (Hugdahl, 2000; Ocklenburg & Güntürkün, 2018). Here we review how each paradigm has been implemented to study language lateralisation and what is known about reliability and validity.

#### Visual half-field paradigm

In this paradigm, visual stimuli are briefly presented to the left or right of a central fixation stimulus. The visual pathways from the left and right sides of space to visual cortex are crossed for non-foveal viewing, and so the right eye projects directly to the left hemisphere, and the left eye to the right hemisphere. The corpus callosum allows for transfer between hemispheres, but the rationale of the method is that there will be a processing advantage for language stimuli that are directly projected to the language-dominant hemisphere. This method has been used since the 1950s (Mishkin & Forgays, 1952). Bourne (2006) reviewed the literature, which consistently shows advantageous processing for letter strings presented in the right, relative to the left, visual field in most people.

Despite wide reporting of a right visual field advantage at the population level,  the validity of visual half-field tasks has been questioned. Krach, Chen, and Hartje (2006) compared the laterality indices (LIs) produced during a visual half-field experiment with those produced using functional Transcranial Doppler Sonography (fTCD). The visual half-field task consisted of lexical decisions made towards abstract nouns presented in the left- and right visual-fields. During the experiment, the blood flow to left and right hemispheres of 58 participants was measured. In addition, blood flow was measured during a word generation task — the gold standard task for assessing hemispheric dominance for language. Despite showing the expected right visual field advantage on the visual half-field task, the behavioural LIs produced on the visual half-field paradigm did not correlate with the fTCD LIs produced during word generation (r= .18) or the visual half-field task itself (r -.004). The lack of validity here likely reflects the poor reliability of the visual field task as split-half reliability was very low (RT spilt-half: r= .14; accuracy split-half: r= .20). This illustrates the difficulty of drawing conclusions about associations between laterality indices when reliability of measurement is so low.

In recent years it has become clear that the reliability and validity of visual half-field tasks to identify the language-dominant hemisphere is largely dependent on several methodological aspects. For example, Hunter and Brysbaert (2008) noted stronger and more stable visual-field differences when stimuli were presented bilaterally relative to those where only one stimulus was presented either in the LVF or in the RVF (Boles, 1987, 1990, 1994; see also Iacoboni & Zaidel, 1996). Brysbaert and Hunter argued that this is the result of stimuli competing during bilateral presentation, enhancing the advantage to the stimulus presented in the language dominant hemisphere. Reliability and validity also depend on there being a sufficient number of trials and brief stimulus presentation to prevent eye movements. When accounting for these methodological aspects, the validity of visual half-field tasks seems promising. For example, Hunter and Brysbaert (2008) used two visual half-field tasks, word naming (192 trials) and picture naming (160) in which stimuli were bilaterally presented and masked at offset. The two tasks correlated .8 (95% CI .598-.956) in a sample of 26 left-handers. Encouragingly, the LIs from these visual half-field tasks also correlated positively with laterality from functional Magnetic Resonance Imaging (fMRI) on a word generation task in 10 left-handers. However, the small sample size means that the confidence intervals around these estimates are very wide: word naming, r= .63 (95% CI = .001-.902), picture naming, r= .77 (95% CI = .272-.943).  <!---I removed Hausmann et al in interests of length, as they didn't report reliability--->

#### Dichotic listening

Simultaneous presentation of different auditory stimuli to left and right ears for immediate recall was a method originally used in studies of attention, but it became evident that there was usually an advantage for verbal stimuli presented to the right ear (Kimura, 1967). Both ears project to both hemispheres, but the weaker ipsilateral pathways appear to be suppressed when the two sides are in competition, and so the right ear, with a contralateral projection to the left hemisphere, will have an advantage in tasks that involve language processing.

Over the years, various dichotic listening paradigms have been used (Bryden, 1988; Westerhausen, 2019). As with the visual half-field paradigm, the test-retest reliability of dichotic listening paradigms has been far from optimal (Voyer, 1998; Kelley & Littenberg, 2019). However, recent data from Westerhausen and Samuelsen (2020) indicated that when designed well, the dichotic listening paradigm can be reliable. First, based on a review of the published literature, Westerhausen and Samuelsen outlined eight design features that should be used in an optimal dichotic listening paradigm and explain their justification for each. The criteria are as follows: (1) stop consonant-vowel (CV) syllables as stimulus materials, (2) pair only CV stimuli from the same voicing category, (3) Alternating trials of voiced and unvoiced stimulus pairs, (4) free-recall instruction, (5) single-stimulus pair per trial, (6) paradigm length of 120 dichotic trials, (7) all stimulus pairs are presented to the two sides in identical frequency, and (8) includes binaural (homonymic) trials. Putting these criteria into practice, Westerhausen and Samuelsen administered a modified dichotic listening paradigm using both a verbal and manual response format to 50 right-handers. Interestingly, the full paradigm yielded a test-retest reliability of r= .93 for verbal responses and r= .91 for manual responses. Thus, the dichotic listening task provides a promising paradigm for those wanting to investigate language lateralisation without reliance on visual stimuli.

#### Chimeric faces

In studies of face processing, it is possible to use a modification of the visual half-field paradigm in which single stimuli are viewed, composed of vertically split chimeric faces, where each half displays a different emotion or identity. Our focus here is on emotion detection in chimeric faces, where a bias is found towards reporting the emotion shown in the left hemiface.

The reliability of the chimeric faces tasks seem promising. Early work by Levy et al (1983) using two vertically presented chimeric faces, found split-half reliability of .93. The validity of the chimeric faces task has been demonstrated using behavioural and fMRI data, but in a task of recognising facial identity rather than emotion. In Yovel, Tambini, and Brandman’s (2008) study, participants performed a behavioural experiment where they had to recognize centrally presented chimeric faces, which presented different identities in the right- and left-visual-field. It is encouraging to note that the behavioural measure significantly correlated with the LI calculated based on activation in the fusiform face area (r= .49). However, the sample size used here was small: with 17 participants, the 95% confidence interval around the observed correlation of .49 ranged from .012 to .786.

### Behavioural laterality tasks and (a)typical lateralisation

The studies reviewed so far have indicated consistent population biases in lateralisation on behavioural tasks, but also individual differences. One factor that has been associated with the degree of lateralisation is handedness. Neuroimaging data indicates that for language tasks these differences in lateralisation reflect the contribution of a subset of strong left-handers who are right language dominant, rather than indicating that this relationship operates on a continuum (Mazoyer et al., 2014). This association can be used to validate behavioural measures of lateralisation: we would expect to see departure from typical lateralisation in strong left-handers.

With regards to visual half-field techniques for studying language lateralisation, there is evidence to suggest that the visual field advantage varies with handedness (Elias, Bulman-Fleming, & McManus, 1999; Weekes, Capetillo-Cunliffe, Rayman, Iacoboni, & Zaidel, 1999). However, a relatively well powered study, involving 100 participants conducted by Willemin et al. (2016) reported no hint of any effect of handedness, when coded as left and right, or when divided into left (N = 19), mixed (N = 24) and right-handed (N = 57) groups, on the visual field advantage during a lexical decision task. This result illustrates the recurring challenge of interpreting null results in this field: do they reflect a true lack of relationship, poor test reliability, inadequate power to detect small effects, or specific methodological factors? We may note that Willemin et al. asked participants to respond via manual response, whereas other studies have involved an active speech production component, which could influence task lateralisation.

On dichotic listening paradigms, the clear right ear (i.e. left hemisphere) advantage that is noted at the population level (see Hugdahl et al., 2009, and Westerhausen, 2019, for a review) is reduced in left-handers (Bethmann, Tempelmann, De Bleser, Scheich, & Brechmann, 2007; Bryden, 1970, 1988; Hugdahl, 2003). More recently, Karlsson, Johnstone, and Carey (2019) employed a consonant-vowel version of the dichotic listening task where participants reported which one of six syllables that they heard the clearest. In study 1, 87% of right-handers and 78% of non-right-handers showed a right ear advantage. For study 2, 85% of right-handers and 79% of non-right-handers showed a right ear advantage.

For chimeric faces, which typically show a left visual field advantage there again appears to be a greater departure from the typical pattern of laterality for left-handers (e.g. Gilbert & Bakan, 1973; Levy et al., 1983; Rozkowski & Snelbecker, 1982). This was also studied in the study by Karlsson et al. (2019). When viewing a single face with a different emotion shown in either hemiface, 75% of right-handers and 60% of non-right-handers showed a leftwards bias. For the second chimeric faces task that involved two chimeric faces arranged vertically, a similar pattern emerged. 73% of right-handers and 57% of non-right-handers showing a leftwards bias.

From the above reviewed studies, it appears that handedness is related to laterality across a range of tasks. This relationship points towards a general laterality factor which governs lateralisation across tasks and behaviours. This has been argued in several hypotheses (Behrmann & Plaut, 2015; Cook, 1984; Hellige, 1983; Kosslyn, 1987). Common to all hypotheses is the assumption that there would be a correlation between the degree of laterality across complementary tasks. While studies do report significant correlations between measures of lateralisation, these are often small (Badzakova-Trajkov, Haberling, Roberts, & Corballis, 2010). Furthermore, there are a large number of studies which report no correlation between laterality indices (Bryden, Hecaen, & Deagostini, 1983; Rosch, Bishop, & Badcock, 2012; Whitehouse & Bishop, 2009; van der Haegen & Brysbaert, 2018). However, it has not been possible to clearly unpick whether the null results indicate that there are multiple independent lateral biases across tasks, or whether there is a more prosaic explanation in terms of poor reliability or validity of tasks.

### The current study

The principal goal of the current study was to evaluate the reliability and validity of an online behavioural laterality battery. Specifically, we examined whether it is possible to observe the usual lateral biases at the population level in tests of language and emotion processing, and whether it is possible to achieve a test-retest reliability of .65 or above using online behaviour research methods. In addition, for tasks where satisfactory reliability was achieved, we considered how many trials were needed to achieve this. The tasks included in the battery were a novel rhyme decision visual half-field task, a consonant-vowel dichotic listening tasks, a chimeric faces visual half-field task, and a novel finger tapping task. These behavioural tasks were completed twice on two different occasions.

Our first set of predictions concerned validity of the tasks as indicators of cerebral lateralisation. Based on previous work, we predicted that at the population level: (1) the laterality indices for Rhyme Decision would show a right visual field advantage (indicating left hemisphere dominance for language processing); (2) the laterality indices for the Dichotic Listening task would show a right ear advantage (indicating left hemisphere dominance for language); and (3) the laterality indices for Chimeric Faces would show a left visual field advantage (indicating right hemisphere dominance for face or emotion processing). The Finger Tapping task was designed to provide a quantitive measure of degree of handedness, and was expected to show a right-hand advantage in the overall sample.

Our second prediction related to handedness and atypical lateralisation. We predicted that strong left-handers (i.e. those with laterality quotient of -90 or less on the Edinburgh Handedness Inventory) would be more likely to show atypical laterality (i.e. laterality in the opposite direction to our population level predictions) than right-handers. This prediction followed from observations by Mazoyer et al (2014, p. 1) who studied a large group of left- and right-handers using fMRI to measure language laterality. They concluded “concordance of hemispheric dominance for hand and for language occurs barely above the chance level, except in a group of rare individuals (less than 1% in the general population) who exhibit strong right hemisphere dominance for both language and their preferred hand”.

Our final prediction related to the inter-relationships between laterality indices. With online testing, there is the opportunity to test inter-relationships between laterality indices with an adequately powered sample using measures of known reliability. Prior studies have been inconclusive because it has been difficult gather large amounts of data to establish test-retest reliability. For instance, van der Haegen and Brysbaert (2018), found that language laterality did not correlate with face laterality in visual half-field tasks. However, their study was restricted to left-handers, who may be atypical in this regard, and they noted that lack of association was hard to interpret because it may be due to unreliability of measurement. In the preregistration, we predicted that for right-handers, a model (a) where the two language tasks are correlated with each other but not with the chimeric faces will provide better fit to the data than a model (b) in which laterality indices are correlated between the language and chimeric face tasks, or a base model (c) where each task is independently lateralised. For left-handers, we anticipated that that model (c) may be a better fit to the data, indicating that laterality is dissociated across the different tasks.

## Method

This project was pre-registered on the Open Science Framework (OSF) prior to the start of data collection. The registration form along with the analysis scripts and anonymised data can be found on the OSF at https://osf.io/qsxcv/.

### Participants

As our ultimate goal is to use these measures in individual differences research, we aimed to optimise the task to achieve observed test-retest reliability of at least .65. Our initial power analysis showed that to achieve sufficient precision around the estimate of reliability, we would require 300 participants. Out of the 300 participants, we originally specified that 80 would be left-handers. This would give us power to apply our planned model-fitting analysis to left-handers as a separate group.

In practice, our final sample exceeded 300, because we discarded participants from individual tasks for failing to meet prespecified criteria (see below). In total, 441 native English speakers were initially recruited through Prolific Academic. Participants had normal or corrected-to-normal vision and reported no history of neurological disease (e.g. head injury, brain tumour, stroke) or significant hearing loss/impairment.

This study was approved by the Medical Sciences Inter-Divisional Research Ethics Committee (reference number: R66638/RE002) and conducted in accordance with the principles of the Declaration of Helsinki. Participants gave informed consent and received compensation for their participation.

### Online behavioural laterality battery

The battery consisted of eight tasks and was administered via Gorilla (Anwyl-Irvine, Massonnié, Flitton, Kirkham, & Evershed, 2020), a cloud-based tool for collecting data in the behavioural sciences. Each task is described below.

#### Basic demographics questionnaire

The basic demographics questionnaire asked participants to report handedness, years in education, and whether they were bilingual. We also asked participants a question about footedness: which foot do you normally use to step up on a ladder/step? Participants could answer "left", "right", or "don't know".

#### Tests of ocular dominance

Adapted versions of the Miles test (Miles, 1929) and the Porta test (Porac & Coren, 1976) were used to determine each participant’s eye dominance in central gaze (i.e., when looking straight ahead). Each test classified participants as being either left or right eye dominant. In practice, the two tests agreed perfectly, so we report here just the Miles test.

#### Edinburgh Handedness Inventory

The Edinburgh Handedness Inventory (EHI: Oldfield, 1971) was administered to quantify handedness on a continuum. For 10 everyday activities, participants indicated their preferred hand use on a 5-step scale (right hand strongly preferred, right hand preferred, no preference, left hand preferred, left hand strongly preferred). The EHI was scored in the standard way, which involves combining information about direction and exclusiveness of hand preference. For simple contrast between left- and right-handers, we treated indices greater than 0 as right handed, and indices less than 0 as left handed. For analysis of extreme left-handers, we focussed on left-handers with an index of -90 or less, which corresponds to near-exclusive use of the left hand.

#### LexTALE

The Test for Advanced Learners of English (Lemhöfer & Broersma, 2012) was used to assess participants’ vocabulary knowledge. Participants saw 60 letter strings and had to indicate which word they knew. Of the 60 strings, 40 were actual English words and 20 were non-words. LexTALE scores were corrected the unequal proportion of words and non-word by averaging the percentages corrected for these two item types (e.g., ((number of words correct/40 x 100) + (number of nonwords correct/20 x 100))/2). Accordingly, the LexTALE scores ranged from 0 to 100.

#### Rhyme decision task

The rhyme decision task was developed for this project as a novel visual half-field technique which requires participants to decide whether a foveally presented letter string rhymed with the name of an image shown in either the left or right hemifield. With online presentation, a particular concern is potential for eye movements away from fixation. We relied on brief bilateral presentation, which has guaranteed adequate control for eye movements in previous VHF experiments with verbal naming (Beaumont, 1982). Studies that directly monitored eye movements in this setting reported failures of fixation on only 0.5% of trials (Geffen, Bradshaw & Nettleton, 1972, but see also Bourne, 2006). The design we used ensured that participants had to start the trial fixating centrally in order to read the target word. Pictures of familiar objects were presented for rhyme matching to the written word, with the goal of ensuring that the participant would have to covertly name the picture, thereby engaging left-hemisphere speech processing.

##### Materials

Twenty-six monosyllabic words were presented in foveal vision. Each word was 3-4 letters in length (*mean*= 3.8; *SD*= 0.40) and had an average Zipf frequency [log10(frequency per million words) + 3] of 4.4 (*SD*= 0.72) when using the SUBTLEX-UK database (van Heuven, Mandera, Keuleers, & Brysbaert, 2014). Each foveal word was paired with an image that corresponded to one of 26 monosyllabic words which had a mean Zipf frequency of 4.5 (*SD*= 0.48). For example, the word “bite” was paired with the image “kite” and the word “more” was paired with the image “door”. Each word-image pair was then paired with another word-image pair for which there was no inter-pair rhyme (i.e. bite-kite was paired with more-door). This resulted in 13 word-image couplets. For each of the foveally presented words in a given word-image couplet, the corresponding images could appear in either the LVF or RVF. Thus, for each of the 13 couplets, four combinations were possible. For the pairing of bite-kite and more-door the following stimuli were created: kite – bite – door, door – bite – kite, door – more – kite, kite – more – door. In total, there were 52 unique stimuli. On a given trial, each image was displayed at 2.5 deg. eccentricity of the foveal word (see Figure 1).

![Figure 1. Schematic illustration of a trial on the rhyme decision task. In the example, the word “bite” is presented along with the two images “kite” and “door”. When making the correct judgement, participants would respond with by pressing ‘S’ to indicate that the rhyme was on the left.](Picture1.png)

##### Procedure

The trials were preceded by a familiarisation phase in which each of the stimulus pictures was presented together with its written name. Then 260 stimuli (52 unique stimuli each shown 5 times) were displayed across four blocks (65 items per block). Each of the four blocks also contained five items for which there was no rhyme. For example, the word “moon” was displayed with the images for kite and door: kite – moon – door. These foil items were included in attempt to prevent participants from preferentially responding to items in either the LVF or RVF. These items were then pseudo-randomised such that identical image pairs could not appear one after the other.

Each trial began with a fixation cross presented centrally for 800 ms. Participants were instructed to keep their eyes on the central fixation point. Then a foveally presented word would appear for 200 ms. Word stimuli were presented in black lowercase Tahoma font on a white background. Font was scaled to 250% equating to 28pt font. The foveally presented word was replaced with bilaterally presented images at approximately 2.5 deg. of visual angle either side of fixation. The eccentricity of images was maintained using Gorilla’s scaling tool. The bilaterally presented images were then presented for 150 ms. At stimulus offset, participants indicated whether the image in the LVF or RVF rhymed with the foveally presented word. Participants were instructed to press ‘S’ to indicate that the foveally presented word rhyme with the image in the LVF and ‘K’ if it rhymed with the image in the RVF. If participants did not detect a rhyme, they were instructed to press the space bar. The response triggered the next trial. Accuracy and response time were recorded.

#### Dichotic listening task

The consonant-vowel dichotic stimuli were taken from a dichotic listening task that was developed to work as a smartphone app, and shown to give high levels of test-retest reliability (Bless et al., 2013). These stimuli were the same as those used by Karlsson et al (2019) and were used with the kind permission of Professor Kenneth Hugdahl.

##### Materials

The consonant-vowel stimuli were formed by pairing six stop-consonants (/b/, /d/, /g/, /p/, /t/, /k/) with the vowel /a/. Each of the six consonant-vowel stimuli (/ba/, /da/, /ga/, /pa/, /ta/, /ka/) were paired and played in each sound channel (e.g. /ba/-/da/), resulting in 36 pairings including homonyms.

##### Procedure

As the dichotic listening task demands the use of stereo-headphones, participants were screened at the start of each session. Screening consisted of two tasks. The first involved participants judging which of three pure tones is quietest, and is described by Woods, Siegel, Traer, and McDermott (2017) in full detail. Sound discrimination is designed to be easy to achieve when using headphones but difficult over loudspeakers due to phase-cancellation. Participants were removed from the study if they failed to correctly discriminate the quietest sound on at least five out of six trials, indicating that they were unlikely to be using headphones. The second check involve participants hearing a sound presented in only one of the two channels. Participants were removed if they failed to correctly identify which channel the sound was played on five out of six trials, indicating that they were not using stereo-headphones.

Participants completed four blocks of the dichotic listening task. In each block, participants heard each of the 36 stimuli once. Across the four blocks, they responded to 144 stimuli. Participants were instructed to wear headphones when completing the dichotic listening task. We included two headphone checks to ensure that participants were using stereo-headphones, which we describe further in the General Procedure. At the start of each trial, participants saw a fixation cross for 250 ms, after which a stimulus was played. After hearing each stimulus, participants were instructed to report the syllable they heard or, if it seemed like they heard two different sounds, the one they heard best or most clearly by clicking a button corresponding to one of the six syllables. The response triggered the start of the next trial.

#### Chimeric faces task

The chimeric faces stimuli were kindly provided by Dr Michael Burt, and have been reported elsewhere (Burt & Perrett, 1997; Innes, Burt, Birch, & Hausmann, 2016; Karlsson et al., 2019).

##### Materials

The faces consisted of symmetrical averaged images created from four male and four female faces. Four emotional expressions were used: anger, disgust, happiness, and sadness. The faces were split vertically down the middle and paired so that one emotive hemiface was attached to another and blended at the midline (see Figure 2 for an example). These were paired in all possible combinations creating 16 individual stimuli. Four stimuli showed identical emotions in each hemiface.

![Figure 2. An example stimulus from the chimeric faces task. In the example, “happiness” is shown on the left and “sadness” on the right. For a participant who is right lateralised, they might be expected to have stronger emotion processing for the left hemiface. As such, they will likely report seeing “happiness”.](chim_example.png)

##### Procedure

Participants completed four blocks of the chimeric faces task. In each block, participants viewed each stimulus twice. Thus, they viewed 32 stimuli per block and 128 in total. At the start of each trial, participants fixated a cross for 1000 ms. A chimeric face was then displayed for 400 ms. Participants then reported the emotion seen in the face by clicking a button corresponding to one of the four emotions. The response triggered the start of the next trial.

#### Finger tapping task

The Finger Tapping task is a novel behavioural task that we designed as a quantitative index of handedness. Participants were required to either press the ‘W’, ‘R’, ‘V’, and ‘X’ keys with their left hand or the ‘Y’, ‘I’, ‘M’, ‘B’ keys with their right hands in sequence as many times as possible within a 30 second window. Participants completed four blocks of the finger tapping task in total. Each block contained required participants to complete the task for both their dominant and non-dominant hand.

### General Procedure

The current study implemented a test-retest, within-subjects design across two sessions. During session 1, participants completed the full battery of tasks. Self -report measures were completed first, followed by the behavioural laterality tasks. During session 2, participants completed only the behavioural laterality tasks. Hence, each participant provided data for the rhyme decision, dichotic listening, chimeric faces, and finger tapping tasks twice. Each session lasted approximately 45 minutes and 35 minutes respectively.

When completing the behavioural tasks participants were instructed to sit approximately 50 cm from the screen. Gorilla’s scaling tool was utilised such that stimuli size and eccentricity was maintained across tasks and participants. Stimuli from each task were repeated across four blocks. Within each block participants completed 70 trials of the rhyme detection task (R); 36 trials of the dichotic listening task (D); 32 trials of the chimeric face task (C); and 2 trials of the finger tapping task (F) for both their dominant and non-dominant hand. The order of tasks with each block were determined as follows RDCF, DCFR, CFRD, FRDC, and the order of tasks remained consistent across blocks for each participant. For each task, trial order was randomised, except for the finger tapping task where participants responded with their dominant hand followed by their non-dominant hand.

### Data analysis

```{r read_dat, echo- FALSE, include= FALSE}
library(ggplot2)
library(tidyr)
library(dplyr)
library(knitr)
library(kableExtra)
library(ggpubr)
library(corrr)
library(flextable)
library(cocor)
options(knitr.kable.NA = '') #should suppress NA in tables
# read in data from csv
#LIs <- read.csv('https://osf.io/x5jd2/download',stringsAsFactors = F)
#We will replace with read from OSF, but better stick with LIs to ensure final version is the one we use 
allsum <- read.csv('./data/allsum.csv',stringsAsFactors = F)

```

```{r prepdat,echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
# remove those without day 2 to make LIs file - we will just use this
# reformat NA
allsum$day2.complete <- ifelse(is.na(allsum$day2.complete), 0, allsum$day2.complete)
LIs <- allsum[allsum$day2.complete == 1,]

# remove
LIs_c <- LIs[LIs$day2.complete == 1,]

# lextale out df
LIs_no <- LIs[LIs$lexOut != "keep",]
# now remove them from LI
LIs <- LIs[LIs$lexOut == "keep",]
# table of counts for footed and sigh
short_dat <- dplyr::select(LIs, "subject", "handedness", "footedness", "miles", "porta")
long_dat <- tidyr::gather(short_dat, task, response, footedness:porta, factor_key=TRUE)
# make table
tab <- as.data.frame(table(long_dat$handedness, long_dat$task, long_dat$response))
tab <- tab %>% 
  rename(
    Handedness = Var1,
    Task = Var2,
    Response= Var3,
    Count= Freq
    )
split <- ggplot(data=tab, aes(x=Response, y=Count, fill=Handedness)) +
  geom_bar(stat="identity", position=position_dodge()) + facet_grid(.~Task) +
    geom_text(aes(label=Count), vjust=-0.3, position = position_dodge(0.9), size=3.5)+
  theme_classic() + theme(legend.position = "top") + ylab("Count") + xlab("") + ylim(0,220)
ggsave(
  "handed_sight.png",
  plot = split,
  width = 6, height = 4,
  dpi = 300
)
#Add a category of handedness that distinguishes extreme left from other left-handers
#In preregistration, extreme left defined as EHI laterality -90 or less.
LIs$handed3 <- LIs$handedness
w<-which(LIs$index_EHI<(-89))
LIs$handed3[w]<-'Extreme.left'
```

Out of 441 participants recruited via Prolific, `r nrow(LIs_c)` (`r round(nrow(LIs_c)/441*100, 1)`%) completed both sessions. We then removed `r nrow(LIs_no)` participants (`r round(nrow(LIs_no)/nrow(LIs_c)*100, 1)`%) from our final sample as their lexTALE score was below 80, indicating that a participant’s level of vocabulary knowledge was below the cut off for lower advanced proficiency. This left a sample of `r nrow(LIs)` participants.

##### Data cleaning

For each task we adopted the following data cleaning procedures. Applying each of these procedures meant that the number of participants left in the final analyses could vary by task. First, for all tasks, we removed trials in which response times were shorter than 200 ms. Response times were log-transformed to reduce skew, as specified in our pre-registration. Outlier reaction times were identified using Hoaglin and Iglewicz’s (1987) procedure, whereby outliers are defined as response times that are 1.65 times the difference between the first and third quantiles (1.65 * (Q3-Q1)) below or above the first and third quantile values (e.g: lower limit = Q1 – 1.65 x (Q3-Q1); upper limit = Q3 + 1.65 x (Q3-Q1)). Our differences criterion was set to 1.65 instead of the typical 2.2 because we expected to obtain asymmetric distributions for response time, with most outliers in one tail of the distribution. The Hoaglin and Iglewicz procedure was applied to each participant's data separately. After applying each data trimming procedure, participants left with less than 80% of trials were removed.

For the chimeric faces task, we removed participants if they scored lower than 75% on trials in which an identical emotion was shown in each hemiface, and for the dichotic listening task, we removed those who scored less than 75% correct on trials where the same sound was played to each ear.  These data trimming procedures left data for `r nrow(LIs)-nrow(LIs[LIs$exclude== 1 | LIs$exclude== 12,])` participants on the chimeric faces `r nrow(LIs)-nrow(LIs[LIs$exclude== 2 | LIs$exclude== 12,])` participants on the dichotic listening task. An additional 7 cases had to be removed because of a technical error that meant they had less than 80 trials presented.

For the rhyme decision task, we removed participants who scored less than 50% accuracy in either visual field. Furthermore, we removed participants who scored less than 75% accuracy on the catch trials (trials in which there was no rhyme). These trimming procedures left data for `r nrow(LIs[LIs$exRDT== 0,])` participants on this task. We considered recruiting further participants to make up the numbers, but this seemed uneconomic as we already had more than 300 cases for the other tasks, and we could tell by this point that the test-retest reliability of the rhyme detection task was unsatisfactory; With `r nrow(LIs[LIs$exRDT== 0,])` participants, the test-retest reliability was *r*= `r round(cor(LIs[LIs$exRDT== 0,]$Day1.RDT_RT_LI, LIs[LIs$exRDT== 0,]$Day2.RDT_RT_LI, use = "complete.obs"),2)`, which is smaller than the *r*= 0.65 that we pre-registered as our criterion for a reliable task. We plan to evaluate modified versions of this task in future, but decided to stop data collection at this point and analysed the `r nrow(LIs)` participants with data on other tasks.

For the finger-tapping task we added a data-cleaning step that had not been pre-registered, to remove participants where there were more than 20 keypresses across left and right hands that did not correspond to the  intended sequence. This could reflect that the individual had forgotten the target sequence, or was not co-operating with instructions, but scrutiny of cases where this occurred suggested that it usually arose because of technical problems. For instance, some individuals had repeated recording of the same keypress at short intervals, suggestive of a sticky keyboard. The cutoff of 20 was arbitrarily selected, and was successful in excluding cases with abnormally low scores in one session.

The final group of `r nrow(LIs)` participants (`r nrow(LIs[LIs$gender== "Female",])` female) had a mean age of `r round(mean(LIs$age, na.rm= TRUE), 1)` years (*SD*= `r round(sd(LIs$age, na.rm= TRUE), 2)` and included `r nrow(LIs[LIs$handedness== "Left",])` left handers (`r nrow(LIs[LIs$handedness== "Left" & LIs$gender== "Female",])` female). `r nrow(LIs[LIs$handedness== "Right" & LIs$bilingual== "Yes",])` right-handers were bilingual and `r nrow(LIs[LIs$handedness== "Left" & LIs$bilingual== "Yes",])` left-handers were bilingual.

##### Methods for calcuating laterality indices, reliability, and population level analysis

We preregistered that we would calculate LIs in two ways: (1) in the traditional fashion (100 x (Right – Left)/(Right + Left)); (2) a standardized LI that involved computing a z-score that represented, for each individual, a comparison between left- and right-sided responses. In fact, for RT data we specified a t-score, but t and z converge when sample size exceeds 30, and for simplicity we refer to these as LI_z scores. These LI_z scores are very highly correlated with the traditional LI measures (correlations typically exceed .95), but can be used to provide a common metric for RT and accuracy measures when identifying individuals who are significantly lateralised. For example, using a conventional 2-sided alpha of .05, we can categorise individuals with an absolute LI_z greater than 1.96 as significantly lateralised.

For the rhyme task, we pre-registered that we would calculate LIs based on log response time (RT) data, as accuracy during piloting was at ceiling. To obtain LI_z we ran a *t*-test for each participant where accurate log RTs were the dependent variable and visual field was the independent variable.

For the dichotic listening and chimeric faces, we specified that LI calculated would be based on trials in which participants correctly identified a stimulus. The count of correct responses that corresponded to each side was used to generate an accuracy LI in two ways: (1) in the traditional fashion (100 x (Right – Left)/(Right + Left)); (2) as a *z*-statistic for each participant, using z = (pR-.5)/sqrt(pR x pL/n), where pR is the proportion of correct responses in the RVF, pL is proportion of L responses in the LVF, and n is total LVF and RVF responses. This method can yield extreme z-scores (for instance, if 83/89 responses are to the right, then z = 15.92), so scores were censored to be in the range -5 to +5.  Finally, LIs for the finger tapping task were calculated in the same manner,  using the number of sequences generated by the left and right hand as opposed to the proportion of correct responses in each visual field.

To address hypothesis 1 and examine the laterality of each task at the population level, we conducted a series of one-sample *t*-tests for each task using session 1 data.

For the reliability assessment, we checked LIs for normality using a Shapiro-Wilk test. As most LIs were non-normal, we used a series of Spearman’s correlations to examine the correlation of coefficients for session 1 versus session 2. For tasks where the test-retest reliability was above .75 with the existing test length, we reanalysed the data using just the first N trials, reducing N until reliability becomes unsatisfactory (below *r*= .65). This was of interest in indicating the most efficient way to administer these tasks in future, while retaining good reliability. However, the full data (including all trials) was used for all subsequent analyses.

##### Atypical laterality

To address hypothesis 2, that extreme left-handers will be more likely to show atypical laterality than right-handers, we categorised participants as extreme left-handers if they scored less than -90 and right-handers if they scored more than 90 on the Edinburgh Handedness Inventory. Our preregistration specified that atypicality would first be defined according to whether the participant's LI_z was significantly lateralised (with alpha = .05) in the opposite direction from the majority. If there were less than 20 cases defined as atypical using this method, we planned to adjust the cut-off on that task until we acquired 20 atypical participants.

For Rhyme Decision and Chimeric Faces task, participants were categorised into four groups: typical on both; atypical on Rhyme Decision only; atypical on Chimeric Faces only, and atypical on both. A chi-square test was then used to test for association with the number of people in each of the four groups and the handedness classification from the Edinburgh Handedness Inventory. This was then be repeated for the Dichotic Listening and the Chimeric Faces tasks. A Bonferroni correction was applied to the outcome of each of the two chi-squared tests.

##### Relationship between language and non-language LIs

We examined the relationship between LIs on the Rhyme Decision, Dichotic Listening, and Chimeric faces by modelling covariances. This allowed us to compare (1) a model where the two language tasks are correlated with each other but not with the chimeric faces, (2) a model where all LIs are correlated, and (3) a base model where the LIs are unrelated. This was completed separately for right- and left-handers using both session 1 and session 2 data.

## Results

Lateral preference characteristics of the sample after exclusions as noted above is shown in Table 1. As participants self-reported eye dominance was always consistent between the Miles and Porta tests, we only report data from the Miles test.

```{r simplifiedlatdescriptives, echo- FALSE, include= FALSE}
#Miles and Porta are identical so only report one.
short_dat$cat3<-'R'
short_dat$cat3[short_dat$handedness=='Left']<-'L'
short_dat$temp<-'R'
short_dat$temp[short_dat$footedness=='Left']<-'L'
short_dat$temp[short_dat$footedness=="Don't know"]<-'X'
short_dat$cat3<-paste0(short_dat$cat3,short_dat$temp)
short_dat$temp<-'R'
short_dat$temp[short_dat$miles=='Left']<-'L'
short_dat$temp[short_dat$miles=="Don't know"]<-'X'
short_dat$cat3<-paste0(short_dat$cat3,short_dat$temp)
t<-table(short_dat$cat3)
lattab<-data.frame(matrix(NA,nrow=6,ncol=5))
colnames(lattab)<-c('Foot.Eye','R.handers.N','R.handers.%','L.handers.N','L.handers.%')
lattab$Foot.Eye<-c('RR','RL','LR','LL','Other','Total')
lattab$R.handers.N[6]<-sum(t[9:16])
lattab$R.handers.N[1:4]<-as.vector(t[c(13,12,10,9)])
lattab$R.handers.N[5]<-lattab$R.handers.N[6]-sum(lattab$R.handers.N[1:4])
lattab$L.handers.N[6]<-sum(t[1:8])
lattab$L.handers.N[1:4]<-as.vector(t[c(5,4,2,1)])
lattab$L.handers.N[5]<-lattab$L.handers.N[6]-sum(lattab$L.handers.N[1:4])
sumR<-lattab$R.handers.N[6]
sumL<-lattab$L.handers.N[6]
lattab$`R.handers.%`<-round(100*lattab$R.handers.N/sumR,1)
lattab$`L.handers.%`<-round(100*lattab$L.handers.N/sumL,1)
lattab$`R.handers.%`[6]<-NA
lattab$`L.handers.%`[6]<-NA
```

Table 1 <!---New table - lattab to be added in nice format--->
Count of left- and right-handers who report left or right on self-reported footedness, and eyedness (Miles test). 
```{r Tab1,echo=FALSE, comment=FALSE, warning=FALSE}
flextable(lattab)
```
*Note*. RR: right foot, right eye; RL: right foot, left eye; LR: left foot, right eye; LL: left foot, left eye. 'Other' inlcudes those who responded 'Don't know'.

```{r computepcorrRDT, echo=FALSE, comment=FALSE, warning=FALSE}
#revised this, as p.corr values now in allsum
RDT.short.dat<-allsum[allsum$exRDT==0,]
tempRDT <- dplyr::select(RDT.short.dat,RDT1.pcorrL,RDT1.pcorrR,RDT2.pcorrL,RDT2.pcorrR)
allmeanpcorr <- mean(colMeans(tempRDT,na.rm=T))
allsdpcorr <- sd(c(tempRDT$RDT1.pcorrL,tempRDT$RDT1.pcorrR,tempRDT$RDT2.pcorrL,tempRDT$RDT2.pcorrR),na.rm=T)
```

### Significance of lateralisation on each task

For a preliminary impression of the data, Figure 3 shows scatterplots indicating scores on left and right sides for each task, with handedness colour-coded. For each task, mean for each side is shown as vertical and horizontal dotted lines, and equal performance on the two sides corresponds to the sloping red dotted line: thus an impression of whether the task is lateralised can be obtained by considering whether points cluster around the sloping lines symmetrically.

```{r altplotraw,echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}

makerawplot <- function(tempLR,mysession,myname,mylim){
  colnames(tempLR)<-c('Left','Right','Handedness')
  tempLR<-tempLR[!is.na(tempLR$Handedness), ] #remove rows with missing data
  tempLR$Handedness<-as.factor(tempLR$Handedness)
  droplevels(tempLR$Handedness)
  levels(tempLR$Handedness)=c('Right','Left')
  #myalpha<-c(1,.75) #haven't been able to adjust cols or pch this way. alpha works but one is always v faint regardless of setting
  meanL<-mean(tempLR$Left,na.rm=T)
  meanR<-mean(tempLR$Right,na.rm=T)
  mytitle<-paste0(myname,"\n Session",mysession)
  myg <- ggplot(tempLR, aes(x=Left, y=Right,col=Handedness)) + geom_point(size=.5) +
    xlim(mylim)+ylim(mylim)+
    labs(title=mytitle)+
    geom_abline(intercept = 0, slope = 1, color="red", 
                linetype="dotted")+
    geom_hline(yintercept=meanR,linetype='dotted')+geom_vline(xintercept=meanL,linetype='dotted')+
    theme(legend.position = "none") + theme_classic() +
    theme(plot.title = element_text(size = 12)
    )
  myg<-myg+ scale_color_manual(values=c("black","red"))
  return(myg)
}
```

```{r dorawplotfunction,echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}

myname<-'Rhyme judge RT'
tempbit<-LIs[LIs$exRDT==0,]
tempLR<-dplyr::select(tempbit,RDTlog1.Lraw,RDTlog1.Rraw,handpch)
#convert back to raw time
tempLR[,1:2]<-exp(tempLR[,1:2])
session<-1
mylim<-c(500,1500)
myg1<-makerawplot(tempLR,session,myname,mylim)

tempLR<-dplyr::select(LIs[LIs$exRDT==0,],RDTlog2.Lraw,RDTlog2.Rraw,handpch)
tempLR[,1:2]<-exp(tempLR[,1:2]) #convert from log back to original scale for plot
session<-2
myg2<-makerawplot(tempLR,session,myname,mylim)

myname<-'Dichotic accuracy'
tempLR<-dplyr::select(LIs[LIs$exclude<2,],DL1.Lraw,DL1.Rraw,handpch)
session<-1
mylim<-c(0,100)
myg3<-makerawplot(tempLR,session,myname,mylim)
tempLR<-dplyr::select(LIs[LIs$exclude<2,],DL2.Lraw,DL2.Rraw,handpch)
session<-2
myg4<-makerawplot(tempLR,session,myname,mylim)

myname<-'Chimeric faces accuracy'
#LIs exclude of 1 or 12 indicates excluded from CF; retain exclude 0,2
tempLR<-dplyr::select(LIs[LIs$exclude %in% c(0,2),],CF1.Lraw,CF1.Rraw,handpch)
session<-1
mylim<-c(0,100)
myg5<-makerawplot(tempLR,session,myname,mylim)
tempLR<-dplyr::select(LIs[LIs$exclude %in% c(0,2),],CF2.Lraw,CF2.Rraw,handpch)
session<-2
myg6<-makerawplot(tempLR,session,myname,mylim)

myname<-'N finger sequences'
tempLR<-dplyr::select(LIs,finger.L1,finger.R1,handpch)
session<-1
mylim<-c(50,180)
myg7<-makerawplot(tempLR,session,myname,mylim)
tempLR<-dplyr::select(LIs,finger.L2,finger.R2,handpch)
session<-2
myg8<-makerawplot(tempLR,session,myname,mylim)

ggarrange(myg1,myg2,myg3,myg4,myg5,myg6,myg7,myg8, ncol = 2, nrow = 4,common.legend=T,legend="bottom")
ggsave(
  "./plots/all.raw.png",
  width = 5, height = 8,
  dpi = 300
)
```

![Figure 3. Left vs right scores for sessions 1 and 2 for the behavioural laterality tasks. The diagonal dotted line shows point of equality: points above this are higher for R than L, and points below are higher for L than R. Mean values for R and L are shown as horizontal and vertical dotted lines. ](all.raw.png)

For the rhyme task, accuracy was high (*M* = `r round(allmeanpcorr,1)`%, *SD* = `r round(allsdpcorr,1)` %),and only correct responses were used for computing mean log RTs (shown transformed back to original ms scale in the plot). For dichotic listening and chimeric faces a correct response can be given from either side; it follows that there will be a negative correlation between the two sides reflecting the fact that the number of selections on the left will be inversely related to the number of correct responses on the right. For these tasks it is evident by inspection that most responses for dichotic listening cluster above the dotted line, indicating a right ear advantage, whereas most responses on chimeric faces cluster below the dotted line, indicating a left hemi-face advantage. Finally, we show the number of correct sequences with each hand on the finger-tapping task. Here, a strong impact of handedness is evident on inspection, with left-handers generally more accurate with the left hand, and right-handers more accurate with the right hand.

The distribution of LIs can be used to quantify departures from symmetry. Recall that for all tasks, we computed LIs in the typical way: (100 x (Right – Left)/(Right + Left)), and that because of crossed pathways from hands, ears and eyes to brain, a positive LI indicates left lateralisation and a negative LI indicates right lateralisation at the brain level. <!---I flipped this for consistency with Hugdahl and others, as it was clear from emails that they found it confusing to have left-brain as negative--->. Table 2 shows these mean LIs, as well a one-sample t-test comparing the mean LI with zero, and results from a Shapiro-Wilks test for departure from normality. It can be seen that all measures were significantly lateralised in the predicted direction at the population level, with left-hemisphere superiority for rhyme judgement, dichotic listening and finger tapping, and right-hemisphere superiority for chimeric faces. Most distributions departed significantly from normality.

<!---for ease of interpretation I have chopped up the original big table so we can look at different aspects one step at a time; i.e. overall distribution of LIs (different from zero / skew); then the standardized LIs; then the test-retest.--->

```{r make.tab2,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
#Make a table to summarise LI means/distribution stats for 4 tasks
#This will gradually get populated as we go through each task
#Cells for each task are filled in course of plots.function chunk
LItable <- data.frame(matrix(NA,nrow=8,ncol=7))
colnames(LItable)<-c('Task','N','Mean','SD','One sample t','One sample p','Normality p')
LItable[,1]<-c('Rhyme Decision 1','Rhyme Decision 2','Dichotic 1','Dichotic 2','Chimeric Faces 1','Chimeric Faces 2','Finger Tapping 1','Finger Tapping 2')

keynames <- c("Day1.RDT_RT_LI"  ,"Day2.RDT_RT_LI" ,"Day1_DL_acc_LI" , "Day2_DL_acc_LI",   "Day1_CF_acc_LI", "Day2_CF_acc_LI","Day1_finger_LI", "Day2_finger_LI"   )
for (i in 1:length(keynames)){
  keycol<-which(names(LIs)==keynames[i]) #find this column
#exclude appropriately
   if (i<7){
  thisdata <- LIs[LIs$exclude %in% c(0,1),]}
  if (i<5){
  thisdata <- LIs[LIs$exclude %in% c(0,2),]}
    if (i<3){
  thisdata <- LIs[LIs$exRDT==0,]}
  temp<-thisdata[!is.na(thisdata[,keycol]),keycol]

  LItable[i,2]<-length(temp)
  LItable[i,3]<-round(mean(temp),2)
  LItable[i,4]<-round(sd(temp),2)
  LItable[i,5]<-round(t.test(temp)$statistic,2)
  pstring <- round(t.test(temp)$p.value,3)
  ifelse(pstring<.001,pstring<-'<0.001',pstring<-char(pstring))

  LItable[i,6]<-pstring
  pstring <- round(shapiro.test(temp)$p.value,3)
  ifelse(pstring<.001,pstring<-'<0.001',pstring<-as.character(pstring))

  LItable[i,7]<-pstring

}
```

Table 2.Mean LIs and results from one-sample t-tests and Shapiro-Wilks tests for departure from normality
```{r Tab2,echo=FALSE, comment=FALSE, warning=FALSE}
tab2 <- flextable(LItable)
autofit(tab2, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
```

Table 3 shows data from the LI_z measures, which make it easier to compare different tasks on the same scale. For each task, those with LI_z scores greater than 1.96 are categorised as left-hemisphere dominant, and those with LI_z scores less than -1.96 are categorised as right-hemisphere dominant. If there were no population bias, we would expect around 2.5% of individuals to meet each of these criteria. It can be seen that for dichotic listening and chimeric faces, around 45% of individuals have hemispheric dominance in the predicted direction and around 9% have hemispheric dominance in the opposite direction. The proportions showing significant laterality are lower for the rhyme decision task (around 17% left-hemisphere dominant, and 5% right-hemisphere dominant), and for finger-tapping, the proportions are only slightly greater than would be expected by chance, although there is a suggestion of increasing laterality on that task across the two sessions.

```{r make.tab3,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
#Make a table to summarise LI means/distribution stats for 4 tasks this time using LI_z.
#This will gradually get populated as we go through each task
#Cells for each task are filled in course of plots.function chunk
LIztable <- data.frame(matrix(NA,nrow=8,ncol=6))
colnames(LIztable)<-c('Task','N','Mean','SD','L-lateralised %','R-lateralised %')
LIztable[,1]<-c('Rhyme Decision 1','Rhyme Decision 2','Dichotic 1','Dichotic 2','Chimeric Faces 1','Chimeric Faces 2','Finger Tapping 1','Finger Tapping 2')

keynames <- c("zlat_RDT_1"  ,"zlat_RDT_2" ,"Day1_Dich_acc_Z" , "Day2_Dich_acc_Z",   "Day1_CF_acc_Z", "Day2_CF_acc_Z","Day1_Finger_acc_Z", "Day2_Finger_acc_Z"   )
for (i in 1:length(keynames)){
  keycol<-which(names(LIs)==keynames[i]) #find this column
   if (i<7){
  thisdata <- LIs[LIs$exclude %in% c(0,1),]}
  if (i<5){
  thisdata <- LIs[LIs$exclude %in% c(0,2),]}
    if (i<3){
  thisdata <- LIs[LIs$exRDT==0,]}
  temp<-thisdata[!is.na(thisdata[,keycol]),keycol]

  LIztable[i,2]<-length(temp)
  LIztable[i,3]<-round(mean(temp),2)
  LIztable[i,4]<-round(sd(temp),2)
  LIztable[i,5]<-round(100*length(which(temp>1.96))/length(temp),1)
  LIztable[i,6]<-round(100*length(which(temp<(-1.96)))/length(temp),1)
}
```

Table 3.Mean LI_z and results from one-sample t-tests and Shapiro-Wilks tests for departure from normality
```{r Tab3,echo=FALSE, comment=FALSE, warning=FALSE}
tab3 <- flextable(LIztable)
autofit(tab3, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
```

### Test-retest reliability

The scatterplots depicting test-retest reliability are shown in Figure 4, with Spearman test-retest correlations in Table 4. These were based on LI_z values, but are essentially identical to values from unscaled LI values.  Confidence intervals were estimated using the *spearman.ci()* function from the RVAideMemoire package (version 0.9-75; Hervé, 2020) with 10,000 iterations.  For all but the rhyme decision task, we obtained good test-retest reliability that exceeded our pre-registered threshold of *r*= .65.

```{r make.combotablex,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
#Make a table to summarise results for all 4 tasks
#This will gradually get populated as we go through each task
#Cells for each task are filled in course of plots.function chunk
combotable <- data.frame(matrix(NA,nrow=4,ncol=3))
colnames(combotable)<-c('Task','Spearman r','95% CI')
combotable[,1]<-c('Rhyme Decision' ,'Dichotic Listening','Chimeric Faces','Finger Tapping ')
```

```{r plots.function, echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
# Adam's code here turned into generic function
# Also saving the scatterplots as png.
# Also computes spearmans with CI and returns these from function
densplots <- function(dat, time1,time2,name1,name2,scattername,y1, y2, x1, x2, ax, ay, remove, title){
  # plots
  dens1<-ggplot(dat, aes(x=time1))+geom_density(linetype="dashed")+theme_classic()+ggtitle(name1)
  dens1 <- dens1+ geom_vline(xintercept = 0)

  # day 2
  # plot density
  dens2<-ggplot(dat, aes(x=time2))+geom_density(linetype="dashed")+theme_classic()+ggtitle(name2)
  dens2 <- dens2+ geom_vline(xintercept = 0)

  # conduct Spearman's wth bootstrapped CI
  myspear<-RVAideMemoire::spearman.ci(time1, time2, nrep = 10000, conf.level = 0.95)

  myci <- paste0(round(myspear$conf.int[1],2),' - ',round(myspear$conf.int[2],2))

  # scatterplot time1 x time2
  # position for label set to 50%x and 90%y
  xpos <- max(time1,na.rm=T)*.5
  ypos <- max(time2,na.rm=T)*.9
  mycor<-cor.test(time1, time2, method = "spearman", alternative = "greater")$estimate
  #Make a label to show the correlation with 95%CI and N on the plot
  mycorlabel<-paste0('r = ',round(mycor,2),'\n95% CI: ',myci,'\nN = ',length(time1))
  #Find position in top L corner for legend
  legposx <- .25 #not working - needs tweaking!
  legposy <- .85
  #censor extreme values for plotting only
  w<-which(time1<(-remove) )
  time1[w] <- (-remove)
  w<-which(time1>remove) 
  time1[w] <- (remove)
  w<-which(time2<(-remove) )
  time2[w] <- (-remove)
  w<-which(time2>remove) 
  time2[w] <- (remove)
  #now do the plot
  myscatter <- ggplot(dat, aes(x= time1, y= time2)) +
    geom_hline(yintercept=0, color = "black") +
    geom_vline(xintercept=0, color = "black") +
    geom_point(aes(color= handed3)) + 
    geom_rug(aes(color= handed3)) +
    theme_classic(18) +
    ylab(name2) + 
    xlab(name1) +
    annotate("text", x = ax, y = ay, label = mycorlabel) +
    theme(legend.position = c(legposx,legposy), legend.title = element_blank()) + ylim(y1, y2) + xlim(x1, x2) +
    ggtitle(title) + theme(plot.title = element_text(hjust = 0.5))
  # Set relative size of marginal plots (main plot 10x bigger than marginals)
#  bigplot<-ggExtra::ggMarginal(myscatter, groupColour = TRUE, groupFill = TRUE)
  scattername<-paste0('./plots/',scattername,".png")
  ggsave(scattername,plot=myscatter,dpi = 300, width = 6, height = 6,)
  return(list(mycor,myci))
}
```

```{r do.plots,LI,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
#Add a category of handedness that distinguishes extreme left from other left-handers
#In preregistration, extreme left defined as EHI laterality -90 or less.
LIs$handed3 <- as.character(LIs$handedness)
w<-which(LIs$index_EHI<(-89))
LIs$handed3[w]<-'Extreme left-hander'
w<-which(LIs$handed3=='Left')
LIs$handed3[w]<-'Moderate left-hander'
w<-which(LIs$handed3=='Right')
LIs$handed3[w]<-'Right-hander'
# RHYME DECISION
# remove bad cases
RD.dat <- LIs[!is.na(LIs$handed3),]
RD.dat <- RD.dat[RD.dat$exRDT == 0,]
# now plot
dat <- RD.dat
time1<-RD.dat$zlat_RDT_1
time2<-RD.dat$zlat_RDT_2
name1<-"Rhyme decision LI_z (day 1)"
name2<-"Rhyme decision LI_z (day 2)"
scattername<-"RhymeDecisionScatter"
y1= -5
y2= 5
x1= -5
x2= 5
ax= 5*.67
ay= -5*.75
remove= 5
mycors<-densplots(RD.dat, time1,time2,name1,name2,scattername, y1, y2, x1, x2, ax, ay, remove, "(A)") 
combotable[1,2]<-round(mycors[[1]],3)
combotable[1,3]<-mycors[[2]]

# DICHOTIC LISTENING
# remove bad cases
DL.dat <- LIs[!is.na(LIs$handed3),]
DL.dat <- DL.dat[DL.dat$exclude == 0 | DL.dat$exclude == 1,]
# now plot
dat <- DL.dat
time1<-DL.dat$Day1_Dich_acc_Z
time2<-DL.dat$Day2_Dich_acc_Z
name1<-"Dichotic listening LI_z (day 1)"
name2<-"Dichotic listening LI_z (day 2)"
scattername<-"DichoticScatter"
mycors<-densplots(DL.dat, time1,time2,name1,name2,scattername,y1, y2, x1, x2, ax, ay, remove,"(B)")
combotable[2,2]<-round(mycors[[1]],3)
combotable[2,3]<-mycors[[2]]

# CHIMERIC FACES
# remove bad cases
CF.dat <- LIs[!is.na(LIs$handed3),]
CF.dat <- CF.dat[CF.dat$exclude == 0 | CF.dat$exclude == 2,]
# now plot
dat <- CF.dat
time1<-CF.dat$Day1_CF_acc_Z
time2<-CF.dat$Day2_CF_acc_Z
name1<-"Chimeric faces LI_z (day 1)"
name2<-"Chimeric faces LI_z (day 2)"
scattername<-"ChimericScatter"
mycors<-densplots(CF.dat, time1,time2,name1,name2,scattername, y1, y2, x1, x2, ax, ay, remove, "(C)")
combotable[3,2]<-round(mycors[[1]],3)
combotable[3,3]<-mycors[[2]]

# FINGER TAPPING
FT.dat <- LIs[!is.na(LIs$handed3),]
FT.dat <- FT.dat[!is.na(FT.dat$Day1_Finger_acc_Z),]
time1<-FT.dat$Day1_Finger_acc_Z
time2<-FT.dat$Day2_Finger_acc_Z
name1<-"Finger tapping LI_z (day 1)"
name2<-"Finger tapping LI_z (day 2)"
scattername<-"FingerScatter"
mycors<-densplots(FT.dat, time1,time2,name1,name2,scattername, y1, y2, x1, x2, ax, ay, remove, "(D)")
combotable[4,2]<-round(mycors[[1]],3)
combotable[4,3]<-mycors[[2]]

```

Table 4. Test-retest reliability estimates and 95% CIs for each behavioural task.
```{r Tab4,echo=FALSE, comment=FALSE, warning=FALSE}
tab4 <- flextable(combotable)
autofit(tab4, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
```
Note. LIs are scored such that a positive value equates to a left lateralisation at the brain level. A positive value equates to left lateralisation at the brain level. RD= rhyme decision, DL= dichotic listening, CF: chimeric faces, and FT: finger tapping.

![Figure 4. Scatter plots for day 1 and day 2 data for LI_z scores on each laterality task: (A) Rhyme decision, (B) Dichotic listening, (C) Chimeric faces, and (D) Finger tapping. Handedness is colour coded with extreme left-handers identified as those who have a score below -90 on the Edinburgh Handedness Inventory. Note that z-scores are censored at absolute values of 5.](combined_LI_plot.png)

### Optimising the tasks

To determine the smallest number of trials needed to achieve a test-retest reliability of *r*= .65, we recalculated LIs for the dichotic listening and chimeric faces tasks using the range of trials in which different emotions/sounds were presented to each hemiface/ear. The absolute test-retest correlation coefficients are shown in Figure 5. A minimum of 60 trials was required to achieve a test-retest correlation of *r*= .65 using the dichotic listening task. However, as indicated in Figure 5, even with 100 trials the reliability estimate still occasionally approaches this 0.65 threshold. Taking this lower number of trials, it would be able to present only two blocks to achieve this threshold of reliability for the dichotic listening task. For the chimeric faces task, a minimum of 25 trials is necessary to achieve a test-retest reliability of *r*= .65. Recall that all possible combinations of the four emotions would result in 12 stimuli in which a different emotion is shown in each hemiface and 4 where the same emotion appears in each hemiface. Three sets of all possible combinations would then be sufficient to yield a good test-retest reliability here (equating to 1.5 blocks of the chimeric faces in the current study).

![Figure 5. Test-retest correlations for (A) the dichotic listening task and (B) the chimeric faces task when using a variable number of trials to calculate laterality indices. The red dashed line marks the test-retest correlation r= .65. The correlations coefficients are shown as absolute values.](all.cor.plot.png)

### Atypical laterality and handedness

According to Mazoyer et al (2014), extreme left-handers are more likely than other handedness groups to have reversed laterality compared to the rest of the population. Our preregistered analysis accordingly focused on comparing extreme left-handers (those with a score of -90 or below on the Edinburgh Handedness Inventory) with right-handers. We defined 'atypical laterality' as laterality that is in the opposite direction to the majority of participants, with an absolute LI_z value greater than 1.96. Although the focus of our planned analyses included only right handers and extreme left handers, we show data for  the remaining left-handers for completeness in Table 5.

```{r makecontingencies,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
#We start by using z-scores to make a 3-way categorisation into:
#1 sig bias in typical direction; 2: NS bias 3: in atypical direction with sig bias.
#
LIs$rhyme1.cut <- cut(LIs$Day1.RDT_p.corr_LI, 
                       breaks = c(-Inf, -1.96, 1.96, Inf), 
                       labels = c("AtypLat", "Nonlat", "TypLat"), 
                       right = FALSE)
LIs$dichotic1.cut <- cut(LIs$Day1_Dich_acc_Z, 
                       breaks = c(-Inf, -1.96, 1.96, Inf), 
                       labels = c("AtypLat", "Nonlat", "TypLat"), 
                       right = FALSE)
LIs$chimeric1.cut <- cut(LIs$Day1_CF_acc_Z, 
                       breaks =  c(-Inf, -1.96, 1.96, Inf), 
                       labels = c("AtypLat", "Nonlat", "TypLat"), 
                       right = FALSE)
mytabr <- table(LIs$rhyme1.cut)
mytabd <- table(LIs$dichotic1.cut)
mytabc <- table(LIs$chimeric1.cut)
mytabc<-mytabc[c(3,2,1)]
mytab2 <-rbind(mytabr,mytabd,mytabc)
row.names(mytab2)<-c('Rhyme decision','Dichotic','Chimeric Faces')
mytab2<-data.frame(mytab2)

#mytab2%>%
#  kable() %>%
#  kable_styling("condensed")

tabr <- table(LIs$handed3,LIs$rhyme1.cut)
#chisq.test(tabr)

tabd <- table(LIs$handed3,LIs$dichotic1.cut)
#chisq.test(tabd)

tabc <- table(LIs$handed3,LIs$chimeric1.cut)
#chisq.test(tabc)

#Planned test omitted group 'Left' and column Nonlat.
chi.r<-chisq.test(tabr[c(1,3),c(1,3)])

chi.d<-chisq.test(tabd[c(1,3),c(1,3)])
chi.c<-chisq.test(tabc[c(1,3),c(1,3)])

#ignore the headings of these tables!
```

Table 5. Frequencies of laterality typicality for right-, left-, and extreme left-handers across the Rhyme decision, Dichotic listening, and Chimeric faces tasks.
```{r Table2,echo=FALSE, comment=FALSE, warning=FALSE}
tabr<-cbind(tabr,round(100*prop.table(tabr,2),1))
tabd<-cbind(tabd,round(100*prop.table(tabd,2),1) )
tabc<-cbind(tabc,round(100*prop.table(tabc,2),1)) #add percentages
blankrow<-c(NA,NA,NA,NA,NA,NA)
latcats <- as.data.frame(rbind(blankrow,tabr,blankrow,tabd,blankrow,tabc))

row.names(latcats) <- c() #remove rownames as they make problems for printing

#Instead have a new column with these
latcats$nucol <- c('Rhyme decision','Extreme left-hander','Moderate left-hander','Right-hander','Dichotic listening','Extreme left-hander','Moderate left-hander','Right-hander','Chimeric faces','Extreme left-hander','Moderate left-hander','Right-hander')

latcats<-latcats[,c(7,3,6,2,5,1,4)] #now reorder columns to make sense
colnames(latcats)<-c('.','L lateralised','L%','Not lateralised','N%','R lateralised','R%')

myft2 <- flextable(
  latcats)

#NB This table needs formatting - has formatted all numbers to 3 decimal places, which is not what we want. Code below achieves this

num_keys <- c("L%","N%","R%")
int_keys <- c("L lateralised","Not lateralised","R lateralised")

ft <- colformat_num(x = myft2, col_keys = num_keys, digits = 1)
ft <- colformat_int(x = ft, col_keys = int_keys, big.mark = ",")

autofit(ft)
```

```{r pformatfunction, echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
pformat<-function(pvalue){
  pp <- "*p* < .001" #default value
  if (pvalue>.001) {
    pp<-paste0("*p* = ",round(pvalue,3))
  }
  return(pp)
}
```

It is evident that the Rhyme Judgement task has few individuals who are significantly lateralised in either direction. We will therefore not consider it further, as our analyses are only justified if the task is an indicator of cerebral lateralisation. Both Dichotic Listening and Chimeric Faces, on the other hand, have a high proportion of strongly lateralised individuals.  The planned analyses, comparing typically lateralised with atypically lateralised individuals, and extreme left vs right-handed individuals, show significant association between extreme left-handedness and atypical laterality on chi square test. For dichotic listening, chi squared = `r round(chi.d$statistic,2)`, d.f. = 1, `r pformat(p.adjust(chi.d$p.value, method= "bonferroni", n=2))`; for chimeric faces, chi squared = `r round(chi.c$statistic,2)`, d.f. = 1, `r pformat(p.adjust(chi.c$p.value, method= "bonferroni", n=2))`.

With all participants included, on the dichotic listening task, `r round(100*tabd[1,3]/sum(tabd[,3]),1)`% of those with atypical laterality were extreme left-handers, `r round(100*tabd[3,3]/sum(tabd[,3]),1)`% were right-handers, and the remainder less extreme left-handers. For those with typical lateralisation, the corresponding figures were `r round(100*tabd[1,1]/sum(tabd[,1]),1)`% extreme left-handers, `r round(100*tabd[3,1]/sum(tabd[,1]),1)`% right-handers, and the remainder less extreme left-handers. For chimeric faces, the association was even more extreme.  `r round(100*tabc[1,1]/sum(tabc[,1]),1)`% of those with Atypical laterality were extreme left-handers, `r round(100*tabc[3,1]/sum(tabc[,1]),1)`% were right-handers, and the remainder less extreme left-handers. For those with typical lateralisation, `r round(100*tabc[1,3]/sum(tabc[,3]),1)`% were extreme left-handers, `r round(100*tabc[3,3]/sum(tabc[,3]),1)`% were right-handers, and the remainder less extreme left-handers.

```{r contingency2tasks,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
tab2task <- table(LIs$dichotic1.cut,LIs$chimeric1.cut)
chi2task<-chisq.test(tab2task)
w1<-which(LIs$chimeric1.cut=='TypLat') #this is confusingnly mislabelled
w2<-which(LIs$dichotic1.cut=='AtypLat')
w3<-intersect(w1,w2)
atypical2<-LIs$handed3[w3] #this is hardcoded in the text below
```

The planned analysis looking at cases who were atypical on both dichotic listening and chimeric faces was hampered by the fact that the two tasks were not associated (see below), and consequently there were only four cases in this category. Nevertheless, it was noteworthy that two of these were extreme left-handers and the other two were less extreme left-handers.

### Relationship between LIs from different tasks

Our final question was whether the laterality indices from the language and faces laterality tasks correlate? Specifically we had predicted that in right-handers the two language tasks would correlate together, but the chimeric faces would form a separate factor, but in left-handers, the language tasks might fractionate.

Initial scrutiny of Spearman correlations between LI_z scores suggests little evidence for any association between the three laterality indices, as shown in Table 6. Table 6 suggests weak associations between LI_z values for the two language measures (rhyme decision and dichotic listening), but no association between either of these measures and LI_z on chimeric faces.

```{r bigcorrmatrix,echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}

# create subset of data frame for analysis - I used LI_z measures, as they are on similar scale
interRR <- dplyr::select(LIs, 'zlat_RDT_1', 'zlat_RDT_2', 'Day1_Dich_acc_Z',
                         'Day2_Dich_acc_Z', 'Day1_CF_acc_Z', 'Day2_CF_acc_Z','handedness')
# rename
interRR <- interRR %>% 
  rename(
    "Rhyme_1" = zlat_RDT_1,
    "Chimeric_1" = Day1_CF_acc_Z,
    "Dichotic_1" = Day1_Dich_acc_Z,
     "Rhyme_2" = zlat_RDT_2,
    "Chimeric_2" = Day2_CF_acc_Z,
    "Dichotic_2" = Day2_Dich_acc_Z
    )

#remove cases with missing values
interRR<-na.omit(interRR)
# subset into right and left handers
left_dat <- subset(interRR, handedness== "Left")
right_dat <- subset(interRR, handedness== "Right")
left_dat<-left_dat[,1:6]
right_dat<-right_dat[,1:6]
totcorr<-correlate(interRR[,1:6],method='spearman')
Rcorr <- correlate(right_dat,method='spearman')
Lcorr <- correlate(left_dat,method='spearman')

#Change so we have one matrix with R handers above diagonal and L handers below
Lcorr<-as.data.frame(Lcorr[,2:7])
Rcorr <- as.data.frame(Rcorr[,2:7])
allcorr<-Lcorr
for (i in 1:6){
  for (j in i:6){
    allcorr[i,j]<-Rcorr[i,j]

  }
}
allcorr<- round(allcorr,3)
row.names(allcorr)<-colnames(allcorr)

NL<-nrow(left_dat)
NR<-nrow(right_dat)
#for left_dat N = 66
# right_dat
```

Table 6. Spearman correlations between the LI_z scores for the different tasks in two Sessions. The upper diagonal shows correlations for the 143 right-handers, and the lower diagonal shows values for the 66 left-handers.
```{r tab 5, echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
myft5<- flextable(allcorr)
autofit(myft5, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
```
Note. For right-handers, an absolute value of .16 is significant at the .05 level, and a value of .22 is significant at the .01 level. For left-handers, a value of .25 is significant at the .05 level and a value of .32 is significant at the .01 level.

In a more formal test we compared (1) a base model where the LIs are unrelated, (2) a model where the two language tasks are correlated with each other but not with the chimeric faces, (3) a model where all LIs are correlated. This was done separately for right- and left-handers, using a subset of SEM that does not include any latent variables or directional paths. In this approach, we constrain particular covariance patterns and report ‘best’ model fit according AIC weights (Wagenmakers & Farrell, 2004). Akaike weights are transformations of the AIC values, which can be directly interpreted as conditional probabilities for each model. The best fitting model to the data for right- and left-handers is selected as the final model.

```{r prepareKievit, echo=FALSE,include=FALSE, comment=FALSE, warning=FALSE}
library(lavaan)
library(tidyverse)
library(MASS)
library(qpcR)
# created handedness data
# now create subset of data frame for analysis
interRR <- dplyr::select(LIs, zlat_RDT_1, zlat_RDT_2, Day1_Dich_acc_Z,
                         Day2_Dich_acc_Z, Day1_CF_acc_Z, Day2_CF_acc_Z,handedness)
# rename, and reverse sign for CF
interRR <- interRR %>% 
  rename(
    "rhyme1" = zlat_RDT_1,
    "faces1" = Day1_CF_acc_Z,
    "dichotic1" = Day1_Dich_acc_Z,
     "rhyme2" = zlat_RDT_2,
    "faces2" = Day2_CF_acc_Z,
    "dichotic2" = Day2_Dich_acc_Z
    )

#remove cases with missing values
interRR<-na.omit(interRR)
interRR$faces1 <- (-interRR$faces1) #reverse sign for faces
interRR$faces2 <- (-interRR$faces2)
# subset into right and left handers
left_dat <- subset(interRR, handedness== "Left")
right_dat <- subset(interRR, handedness== "Right")

```

As shown in Table 7, the Akaike weights favoured the base model for both left-handers and right-handers. This reflected the lack of cross-task correlation for LI_z measures.

Table 7. Akaike weights for the independent base model, language model, and laterality model.
```{r AICmodels,echo=FALSE, comment=FALSE, warning=FALSE}
#NB in pre-reg we just had time 1 data, but here we use time 1 and 2

 #Model 1 all independent
    model1<-"
    rhyme1~~rhyme2
    rhyme1~~0*dichotic1
    rhyme1~~0*dichotic2
     rhyme1~~0*faces1
    rhyme1~~0*faces2
    rhyme2~~0*dichotic1
    rhyme2~~0*dichotic2
     rhyme2~~0*faces1
    rhyme2~~0*faces2
    dichotic1~~dichotic2
    dichotic1~~0*faces1
    dichotic1~~0*faces2
     dichotic2~~0*faces1
    dichotic2~~0*faces2
    faces1~~faces2
    "
#Model 2: language tasks are correlated
    model2<-"
  rhyme1~~rhyme2
    rhyme1~~dichotic1
    rhyme1~~dichotic2
     rhyme1~~0*faces1
    rhyme1~~0*faces2
    rhyme2~~dichotic1
    rhyme2~~dichotic2
     rhyme2~~0*faces1
    rhyme2~~0*faces2
    dichotic1~~dichotic2
    dichotic1~~0*faces1
    dichotic1~~0*faces2
     dichotic2~~0*faces1
    dichotic2~~0*faces2
    faces1~~faces2
    "
#Model 3:all correlated
    model3<-"
    rhyme1~~rhyme2
    rhyme1~~dichotic1
    rhyme1~~dichotic2
     rhyme1~~faces1
    rhyme1~~faces2
    rhyme2~~dichotic1
    rhyme2~~dichotic2
     rhyme2~~faces1
    rhyme2~~faces2
    dichotic1~~dichotic2
    dichotic1~~faces1
    dichotic1~~faces2
     dichotic2~~faces1
    dichotic2~~faces2
    faces1~~faces2
    "
#Make a little dataframe for AIC weights
AICdf <- data.frame(matrix(NA,nrow=2,ncol=4))
AICdf[,1]<-c('Right-handers','Left-handers')
colnames(AICdf)<-c('Handedness','Independent','Language factor','Laterality factor')
for (myhand in 1:2) {
  mydat<-right_dat
  if(myhand==2) {mydat<-left_dat}

 #Using data from this handedness type, test all 3 models
fit1A <- cfa(model1, data=mydat)
fit2A <- cfa(model2, data=mydat)
fit3A <- cfa(model3, data=mydat)
    aic_vector1 <- c(fitMeasures(fit1A, "aic"),fitMeasures(fit2A, "aic"),fitMeasures(fit3A, "aic"))

    #Confirms that for each model, lowest AIC obtained when the simulated data do match the model
    # 
    # The AIC weights make massive difference for preferred model!
    AICweights<-akaike.weights(aic_vector1)$weights
    # 
 AICdf[myhand,2:4]<-round(AICweights,3)
 }
myft6 <- flextable(AICdf)
autofit(myft6, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
```

### Exploratory analyses

#### Finger tapping and Edinburgh Handedness Inventory

```{r fingerEHI,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
spearCI <- RVAideMemoire::spearman.ci(LIs$Day1_finger_LI, LIs$index_EHI, nrep = 10000, conf.level = 0.95)
spearCI2 <- RVAideMemoire::spearman.ci(LIs$Day2_finger_LI, LIs$index_EHI, nrep = 10000, conf.level = 0.95)
# spearman's test
speartest <- cor.test(LIs$Day1_finger_LI, LIs$index_EHI, method = "spearman", alternative = "greater")
speartest2 <- cor.test(LIs$Day2_finger_LI, LIs$index_EHI, method = "spearman", alternative = "greater")
```

To examine the relationship between the finger tapping task and a conventional hand preference battery, we computed Spearman's rank correlation between LIs produced on the finger tapping task and the Edinburgh Handedness Inventory. For session 1 data, the relationship between the tasks was moderate, *rs*= `r round(speartest$estimate, 2)`, *95% CI* [`r round(spearCI$conf.int[1], 2)`, `r round(spearCI$conf.int[2], 2)`],  *p*< 0.001. For session 2, the correlation was *rs*= `r round(speartest2$estimate, 2)`, *95% CI* [`r round(spearCI2$conf.int[1], 2)`, `r round(spearCI$conf.int[2], 2)`],  *p*< 0.001. This suggests that the influence of hand preference on the tapping task may increase with practice.

#### Handedness measurements and their association with laterality

```{r hand_explore, echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
hand_cors <- data.frame(matrix(ncol = 5, nrow = 1))
  colnames(hand_cors) <- c(" ", "DL 1", "CF 1", "DL 2", "CF 2")
# add rows
# row 1
hand_cors[1, 1] = "EHI"
hand_cors[1, 2] = round(cor(LIs$Day1_DL_acc_LI, LIs$index_EHI, use= "complete.obs"), 2)
hand_cors[1, 3] = round(cor(LIs$Day1_CF_acc_LI, LIs$index_EHI, use= "complete.obs"), 2)
hand_cors[1, 4] = round(cor(LIs$Day2_DL_acc_LI, LIs$index_EHI, use= "complete.obs"), 2)
hand_cors[1, 5] = round(cor(LIs$Day2_CF_acc_LI, LIs$index_EHI, use= "complete.obs"), 2)
# row 2
hand_cors[2, 1] = "FT"
hand_cors[2, 2] = round(cor(LIs$Day1_DL_acc_LI, LIs$Day1_finger_LI, use= "complete.obs"), 2)
hand_cors[2, 3] = round(cor(LIs$Day1_CF_acc_LI, LIs$Day1_finger_LI, use= "complete.obs"), 2)
hand_cors[2, 4] = round(cor(LIs$Day2_DL_acc_LI, LIs$Day2_finger_LI, use= "complete.obs"), 2)
hand_cors[2, 5] = round(cor(LIs$Day2_CF_acc_LI, LIs$Day2_finger_LI, use= "complete.obs"), 2)
# table 
corrTabs <- flextable(
  head(hand_cors))
# same scale
LIs$negEHI <- -LIs$index_EHI
# compare
cocor.result <- cocor(~Day1_DL_acc_LI + negEHI | Day1_DL_acc_LI + Day1_finger_LI, LIs)
cocor.result2 <- cocor(~Day1_CF_acc_LI + negEHI | Day1_CF_acc_LI + Day1_finger_LI, LIs)
cocor.result3 <- cocor(~Day2_DL_acc_LI + negEHI | Day2_DL_acc_LI + Day2_finger_LI, LIs)
cocor.result4 <- cocor(~Day2_CF_acc_LI + negEHI | Day2_CF_acc_LI + Day2_finger_LI, LIs)
# ASSOCIATIONS
# using self report
# dichotic
self_DL <- table(LIs$handedness,LIs$dichotic1.cut)
self_DL.chi <- chisq.test(self_DL)
# chimeric
self_CF <- table(LIs$handedness,LIs$chimeric1.cut)
self_CF.chi <- chisq.test(self_CF)
# combinw
self_hand <- as.data.frame(cbind(self_DL, self_CF))
  self_hand$` ` <- rownames(self_hand)
# table
self_hand_Tabs <- flextable(
  head(self_hand),
  col_keys = c(" ", "DL TypLat", "DL NonLat", "DL AtypLat", "CF TypLat", "CF NonLat", "CF AtypLat"))
# using finger tapping
LIs$finger1.cut <- cut(LIs$Day1_finger_LI, 
                       breaks = c(-Inf, 0, Inf), 
                       labels = c("Right", "Left"), 
                       right = FALSE)
# dichotic
finger_DL <- table(LIs$finger1.cut,LIs$dichotic1.cut)
finger_DL.chi <- chisq.test(finger_DL)
# chimeric
finger_CF <- table(LIs$finger1.cut,LIs$chimeric1.cut)
finger_CF.chi <- chisq.test(finger_CF)
# combinw
finger_hand <- as.data.frame(cbind(finger_DL, finger_CF))
    finger_hand$` ` <- rownames(finger_hand)
# table
finger_hand_Tabs <- flextable(
  head(finger_hand),
  col_keys = c(" ", "DL TypLat", "DL NonLat", "DL AtypLat", "CF TypLat", "CF NonLat", "CF AtypLat"))
# using EHI
LIs$EHI.cut <- cut(LIs$index_EHI, 
                       breaks = c(-Inf, -40, 40, Inf), 
                       labels = c("Left", "Ambidextrous", "Right"), 
                       right = FALSE)
# dichotic
EHI_DL <- table(LIs$EHI.cut,LIs$dichotic1.cut)
EHI_DL.chi <- chisq.test(EHI_DL)
# chimeric
EHI_CF <- table(LIs$EHI.cut,LIs$chimeric1.cut)
EHI_CF.chi <- chisq.test(EHI_CF)
# combinw
EHI_hand <- as.data.frame(cbind(EHI_DL, EHI_CF))
    EHI_hand$` ` <- rownames(EHI_hand)
# table
EHI_hand_Tabs <- flextable(
  head(EHI_hand),
  col_keys = c(" ", "DL TypLat", "DL NonLat", "DL AtypLat", "CF TypLat", "CF NonLat", "CF AtypLat"))
```

<!---Adam I removed the exploratory analyses of different definitions of handedness, as I was worried the paper was getting too long and indigestible. These could be put in a supplementary section, but I think they are too much at this point--->

#### Vocabulary knowledge and laterality

```{r exploreA, echo= FALSE, comment=FALSE, warning=FALSE}
# make sure we have good cases
interR <- LIs[LIs$exclude == 0,]
DL.lex<-RVAideMemoire::spearman.ci(interR$Day1_DL_acc_LI, interR$lexTALE, nrep = 10000, conf.level = 0.95)
DL.lex.est<-cor.test(interR$Day1_DL_acc_LI, interR$lexTALE, method = "spearman")
DL.lex1<-RVAideMemoire::spearman.ci(interR$Day2_DL_acc_LI, interR$lexTALE, nrep = 10000, conf.level = 0.95)
DL.lex.est1<-cor.test(interR$Day2_DL_acc_LI, interR$lexTALE, method = "spearman")
# footedness
foot.dat <- interR[interR$footedness != "Don't know",]
foot.dat$footedness <- as.factor(foot.dat$footedness)
#  levels(droplevels(foot.dat$footedness))
DL.ft.t.1 <- t.test(foot.dat$Day1_DL_acc_LI~ foot.dat$footedness)
DL.ft.t.2 <- t.test(foot.dat$Day2_DL_acc_LI~ foot.dat$footedness)
CF.ft.t.1 <- t.test(foot.dat$Day1_CF_acc_LI~ foot.dat$footedness)
CF.ft.t.2 <- t.test(foot.dat$Day2_CF_acc_LI~ foot.dat$footedness)
# sightedness
foot.dat <- interR[interR$miles != "Don't know",]
foot.dat$miles <- as.factor(foot.dat$miles)
#  levels(droplevels(foot.dat$miles))
DL.st.t <- t.test(foot.dat$Day1_DL_acc_LI~ foot.dat$miles)
DL.st.t <- t.test(foot.dat$Day2_DL_acc_LI~ foot.dat$miles)
CF.st.t <- t.test(foot.dat$Day1_CF_acc_LI~ foot.dat$miles)
CF.st.t <- t.test(foot.dat$Day2_CF_acc_LI~ foot.dat$miles)
```

The LexTALE vocabulary measure had been included largely for the purpose of data exclusion, but we had planned to explore correlations between LexTALE score and the laterality index on the Rhyme Decision task. However, as the rhyme task was less reliable than other measures, and showed only small asymmetries, we chose to examine the relationship between the LexTALE and the other verbal laterality measure, dichotic listening, instead. From session 1 data, after correcting for multiple comparisons, the correlation between the LIs on the dichotic listening and LexTALE score was small and non-significant, *rs*= `r round(DL.lex.est$estimate, 2)`, *95% CI* [`r round(DL.lex$conf.int[1], 2)`, `r round(DL.lex$conf.int[2], 2)`] *p*= `r round(p.adjust(DL.lex.est$p.value[1], n= 2, method= "bonferroni"), 3)`. An identical pattern of results was obtained using session 2 data, *rs*= `r round(DL.lex.est1$estimate, 2)`, *95% CI* [`r round(DL.lex1$conf.int[1], 2)`, `r round(DL.lex1$conf.int[2], 2)`] *p*= `r round(p.adjust(DL.lex.est1$p.value[1], n= 2, method= "bonferroni"), 3)`.

<!--- footedness etc also removed here ; could be some interest in looking at people with unusual laterality patterns on eye/foot/hand, but here it is too much--->

## Discussion

With the advancement of online behaviour research methods, it has become possible to test large samples of participants outside of the lab. In the current study we introduced an online behavioural laterality battery that was administered to a relatively large sample of right- and left-handers. In addition to completing self-report measures of handedness, footedness, and sightedness, participants completed a visual half-field rhyme decision task, a dichotic listening task, a chimeric faces task, and a finger tapping task twice. Thus, we were able to examine the test-retest reliability of each task. We also examined the association between handedness and atypical lateralisation and examined the intercorrelations between LIs on the dichotic listening and chimeric faces tasks. We discuss each of these points in turn.

Each behavioural task yielded a hemispheric advantage in the expected direction, but the size of asymmetries and the test-retest reliability of the different tasks varied. A relatively small proportion of individuals showed significant asymmetry on the rhyme decision task. This task also yielded the poorest test-retest reliability, just falling short of our target minimum cutoff of r = .65.  Compared to early studies (e.g. Krach et al., 2006) the rhyme decision task used here was far more reliable but, compared to other visual half-field tasks, it was far from optimal. Van der Haegen and Brysbaert (2018) have reported that, under lab conditions, visual half-field paradigms can yield test-retest reliabilities of *r*= .77 when using pictures and *r*= .83 when using words. Our study was well-powered, so the relatively poor task validity and reliability are likely to be due to methodological factors. One key difference between van der Haegen and Brysbaert’s and our study was that their language based visual half-field tasks involved overt naming. Speech production is highly lateralised (Kosslyn, 1987) and requiring speech production should yield a hemispheric advantage. Our task was designed to require covert speech, but it is possible that an active speech production component might reveal stronger lateralisation. This would, however, be logistically difficult to implement online. The other main limitation of online testing is that the VHF method depends critically on presenting lateralised visual stimuli that project to one hemisphere. Although we were able to calibrate screen size at the start of the session, and to request participants to sit a certain distance from the screen, it is plausible this was not sufficient to achieve adequately lateralised presentation. While we can format images to be consistent across displays, we cannot control participants’ behaviour. Outside of the lab, participants can freely move their head which will influence the size of images on the retina, and it is difficult to ensure images are shown in foveal or parafoveal vision. One avenue that might be worth exploring is the use of images presented in peripheral vision, as it is unlikely that head movements would result in the image shifting between distinct regions on the retina. This increased difficultly in processing of information in this region may also result in stronger visual field advantages (c.f. Hunter and Brysbaert, 2008).

In contrast to rhyme decision, the other verbal task, dichotic listening, showed excellent validity and reliability. Participants showed a clear right-ear advantage at the population level, and the task exceeded our pre-registered reliability cut-off of *r*= .65. Indeed, the test-retest correlation of the dichotic listening task, *r* = .78, was identical to that reported by Bless et al. (2013) using the iDichotic application, and was higher than that reported for the same dichotic listening task when run under lab conditions (Bless et al., 2013).  We further showed that with as few as 60 trials, the current implementation of the dichotic listening task yielded a test-retest correlation that than *r*= .65, suggesting that those who wish to use this paradigm in the future could use a smaller number of trials and add additional tasks on that are important to their ongoing research questions. It would be of interest to extend this line of work to use the  modified version of dichotic listening described by Westerhausen and Samuelsen (2020), which had even higher reliability. Of course, it is important to screen for adequate and appropriate headphones when devising online measures.

Like the dichotic listening task, the chimeric faces task showed impressive levels of both validity and reliability, this time with a strong right hemisphere bias at the population level. The task has previously be shown to have high split-half reliability (Levy et al., 1983), we show that it can meet the higher standard of achieving a high test-retest correlation. Our subsequent analysis showed that acceptable levels of reliability can be achieved with this task using very few trials: test-retest reliability of *r*= .65 could be found with 25 trials. It seems likely that one reason for the success of the chimeric faces task online is the task’s simplistic nature. From the perspective of the participant, it is a simple task of reporting the emotion expressed in a single  face, and does not require any explicit comparison.

We also introduced a novel finger tapping task, which we designed to quantify handedness behaviourally. This too produced a right hand advantage at the population level, and test-retest reliability of the LIs was good, even though the asymmetries were small in absolute terms. Furthermore, our exploratory analysis indicated that LIs from this task correlated reasonably well with the Edinburgh handedness inventory. Previous studies have reported correlations between hand preference and measures of relative hand skill, but different hand skill measures typically are weakly intercorrelated (Bishop, 1990), and it seems that asymmetries may vary depending on particular aspects of planning and execution of motor coordination that are involved. It would be of interest to see how far LIs resemble that from tasks such as the pegboard task (e.g. Hodgson, Hirst, & Hudson, 2016), which is widely used in neuropsychological assessment. Future work should also examine how other factors such as complexity of key combinations and reach through the midline influence the task.

In addition to assessing lateralisation of the task using traditional LIs, we also examined the potential to use individual *z*-scores to quantify lateralisation on each task. This method of calculating LIs is perhaps more useful for those interested in looking atypical lateralisation, or for cases where there is a need to compare laterality from different tasks on a common scale. Rather than selecting those who fall either side of an arbitrary cut-off, the LI_z score allows researchers to assess whether participants are consistently and statistically lateralised in either a typical or atypical manner. The availability of an online measure means that a researcher who is interested in studying the correlates of atypical language lateralisation could assess participants with our battery to identify those with LI_z scores > 1.96 or <1.96 before inviting them into the lab for more detailed in-person testing. This has the potential to save time and effort by screening participants to identify a sample that is enriched for rare cases of reversed cerebral lateralisation.

While the primary purpose of the study was to validate the battery and establish its test-retest reliability, we addressed two theoretically motivated questions related to handedness and atypicality and intercorrelations between the dichotic listening and chimeric faces task. In our pre-registered analysis of handedness, we categorised participants into right,-, left-, and extreme left-handers. The data are consistent in showing that extreme left-handers showed increased atypicality on both the dichotic listening and chimeric faces task. Mazoyer et al. (2014) reported a relationship between strength of lateralisation and hand preference., which was largely driven by the existence of a group with strong left hand preference and strong right hemisphere dominance. They further claimed that this right hemisphere dominance is almost exclusively present in left-handers. Nevertheless, although strong left-handers were over-represented in this category, we did also observe cases of right-handers with right language dominance on the dichotic listening task. This raises the interesting question of whether right language dominance on dichotic listening is equivalent to right language dominance on more direct measures of brain lateralisation, such as fMRI.

A common mechanism that codetermines lateralisation on a range of skills, extending beyond handedness and language, has been a hotly debated topic in the laterality literature. Several hypotheses postulate complementary lateralisation for left-hemisphere and right-hemisphere brain asymmetries, predicting that LIs on verbal and nonverbal tasks should be related (Behrmann & Plaut, 2015; Hellige, 1983; Kosslyn, 1987). On this view, a person with atypical right-hemisphere language representation should have atypical left-hemisphere representation for nonverbal domains such as face processing.  However, several studies have report no such association, or only a weak relationship (Bradzakova-Trajkov et al., 2010; Bryden et al.,1983; Rosch et al., 2012; Whitehouse & Bishop, 2009; van der Haegen & Brysbaert, 2018). Data from the current work supports a view in which laterality on each task is independent. Not only was our study well powered to examine this relationship, it included a large sample of left-handers who show increased variability in their LIs. Specifically, we found no association between the LI_z indices for rhyme decision, dichotic listening and chimeric faces. The lack of association between the two language tasks, rhyme decision and dichotic listening, must be interpreted cautiously, given that the rhyme decision task was unsatisfactory in several respects: it had relatively low test-retest reliability, and only weak asymmetry in the sample as a whole. It is possible that, as argued by Bishop (2013), there is fractionation between language skills: the dichotic listening task is a purely perceptual task, whereas the rhyme decision task involves implicit speech production. The current result suggests there could be interest in replicating this comparison of the two tasks using a visual-half-field language task that has better reliability and validity.  The lack of association between dichotic listening and chimeric faces is, however, much more convincing, given that both tasks show strong lateralisation at the population level and high test-retest reliablity. This supports the view that lateralisation is determined by a series of independent biases: a left-hemisphere population bias for language and a right-hemisphere bias for facial expressions of emotion.

To conclude, the present study introduced an online behaviour laterality battery that can be used to examine laterality for speech perception, face emotionality, and manual preference with good test-retest reliability. The battery of tasks was administered to a large sample of participants with relative ease, and enabled us to examine several questions related to lateralisation. We find evidence of increased atypical lateralisation in left- relative to right-handers. However, we find no indication that the laterality indices on one task are dependent on another and we conclude that the causes of lateralisation in different domains stem from multiple independent sources.

\newpage
## Data accessibility
All analysis scripts and anonymised data are avaiable on the Open Science Framework (https://osf.io/qsxcv/).

\newpage
## Competing Interests
We declare we have no competing interestes.

\newpage
## Funding 
This study was supported by an Advanced Grant from the European Research Council (694189).

\newpage
## Acknowledgements
We would like to thank Dr Mike Burt (University of Durham, Uk) for providing stimuli for the chimeric faces task. We would also like to thank Professor Kenneth Hugdahl (University of Bergen, Norway) for providing stimuli for the dichotic listening task. Portions of this work were presented at the Online Meeting of the Experimental Psychology Society in July 2020.

\newpage
## References

Badzakova-Trajkov, G., Häberling, I. S., Roberts, R. P., & Corballis, M. C. (2010). Cerebral asymmetries: complementary and independent processes. *PloS one*, *5*, e9682.

Beaumont, J. G. (1982). Studies with verbal stimuli. In J. G. Beaumont (Ed.), *Divided visual field studies of cerebral organisation* (pp. 58–86). London: Academic Press.

Behrmann, M., & Plaut, D. C. (2015). A vision of graded hemispheric specialization. *Annals of the New York Academy of Sciences*, *1359*, 30-46.

Bethmann, A., Tempelmann, C., De Bleser, R., Scheich, H., & Brechmann, A. (2007). Determining language laterality by fMRI and dichotic listening. *Brain research*, *1133*, 145-157.

Bishop DVM. 2013 Cerebral asymmetry and language development: cause, correlate, or consequence? *Science*, *340*, 1230531.

Bishop, D. V. (1990). *Handedness and developmental disorder*. Cambridge University Press.

Bishop, D. V. M. (2017) *Which neuroimaging measures are useful for individual differences research? * [Blog post]. Retrieved from: http://deevybee.blogspot.be/2017/05/whichneuroimaging-measures-are-useful.html?m=1

Bless, J. J., Westerhausen, R., Arciuli, J., Kompus, K., Gudmundsen, M., & Hugdahl, K. (2013). “Right on all occasions?”–On the feasibility of laterality research using a smartphone dichotic listening application. *Frontiers in Psychology*, *4*, 42.

Boles, D. B. (1987). Reaction time asymmetry through bilateral versus unilateral stimulus presentation. *Brain and Cognition*, *6*, 321–333.

Boles, D. B. (1990). What bilateral displays do. *Brain and Cognition*, *12*, 205– 228.

Boles, D. B. (1994). An experimental comparison of stimulus type, display type and input variable contributions to visual field asymmetry. *Brain and Cognition*, *24*, 184–197.

Bourne, V. J. (2006). The divided visual field paradigm: Methodological considerations. *Laterality*, *11*, 373–393

Bryden, M. P. (1970). Laterality effects in dichotic listening: relations with handedness and reading ability in children. *Neuropsychologia*, *8*, 443-450.

Bryden, M. P. (1988). An overview of the dichotic listening procedure and its relation to cerebral organization. In K. Hugdahl (Ed.), *Handbook of dichotic listening: Theory, methods and researc*h (p. 1–43). John Wiley & Sons.

Bryden, M. P., Hécaen, H., & DeAgostini, M. (1983). Patterns of cerebral organization. *Brain and language*, *20*, 249-262.

Burt, D. M., & Perrett, D. I. (1997). Perceptual asymmetries in judgements of facial attractiveness, age, gender, speech and expression. *Neuropsychologia*, *35*, 685-693.

Elias, L. J., Bulman-Fleming, M. B., & McManus, I. C. (1999). Visual temporal asymmetries are related to asymmetries in linguistic perception. *Neuropsychologia*, *37*, 1243-1249.

Geffen, G., Bradshaw, J. L., & Nettleton, N. C. (1972). Hemispheric asymmetry: verbal and spatial encoding of visual stimuli. *Journal of Experimental Psychology*, *95*, 25-31.

Gilbert, C., & Bakan, P. (1973). Visual asymmetry in perception of faces. *Neuropsychologia*, *11*, 355-362.

Hellige, J. B. (1990). Hemispheric asymmetry. *Annual Review of Psychology*, *41*, 55-80.

Hervé, M. (2020). *RVAideMemoire: Testing and Plotting Procedures for Biostatistics*. R package version 0.9-75.https://CRAN.R-project.org/package=RVAideMemoire

Hoaglin, D. C., & Iglewicz, B. (1987). Fine-tuning some resistant rules for outlier labeling. *Journal of the American statistical Association*, *8*, 1147-1149.

Hodgson, J. C., Hirst, R. J., & Hudson, J. M. (2016). Hemispheric speech lateralisation in the developing brain is related to motor praxis ability. *Developmental cognitive neuroscience*, *22*, 9-17.

Hugdahl, K. (2000). Lateralization of cognitive processes in the brain. *Acta psychologica*, *105*, 211-235.

Hugdahl, K. (2003). Dichotic listening in the study of auditory laterality. In K. Hugdahl and R. J. Davidson (Eds.), *The asymmetrical brain* (pp. 441–476). Cambridge, MA: MIT Press.

Hugdahl, K., Westerhausen, R., Alho, K., Medvedev, S., Laine, M., & Hämäläinen, H. (2009). Attention and cognitive control: unfolding the dichotic listening story. *Scandinavian journal of psychology*, *50*, 11-22.

Hunter, Z. R., & Brysbaert, M. (2008). Visual half-field experiments are a good measure of cerebral language dominance if used properly: Evidence from fMRI. *Neuropsychologia*, *46*, 316–325

Iacoboni, M., & Zaidel, E. (1996). Hemispheric independence in word recognition: Evidence from unilateral and bilateral presentations. *Brain and Language*, *53*, 121–140

Karlsson, E. M., Johnstone, L. T., & Carey, D. P. (2019). The depth and breadth of multiple perceptual asymmetries in right handers and non-right handers. *Laterality: Asymmetries of Body, Brain and Cognition*, *24*, 707-739.

Kelley KS, Littenberg B. (2019). Dichotic Listening Test–retest reliability in children. *Journal of Speech, Language, and Hearing Research*, *62*,169-76.

Kimura, D. (1967). Functional asymmetry of the brain in dichotic listening. *Cortex*, *3*, 163-178.

Kosslyn, S. M. (1987). Seeing and imagining in the cerebral hemispheres: a computational approach. *Psychological review*, *94*, 148.

Krach, S., Chen, L. M., & Hartje, W. (2006). Comparison between visual halffield performance and cerebral blood flow changes as indicators of language dominance. *Laterality*, *11*, 122-140

Lemhöfer, K., & Broersma, M. (2012). Introducing LexTALE: A quick and valid lexical test for advanced learners of English. *Behavior research methods*, *44*, 325-343.

Levy, J., Heller, W., Banich, M. T., & Burton, L. A. (1983). Asymmetry of perception in free viewing of chimeric faces. *Brain and cognition*, *2*, 404-419.

Mazoyer, B., Zago, L., Jobard, G., Crivello, F., Joliot, M., Perchey, G., ... & Tzourio-Mazoyer, N. (2014). Gaussian mixture modeling of hemispheric lateralization for language in a large sample of healthy individuals balanced for handedness. *PloS one*, *9*, e101165.

Miles, W. R. (1929). Ocular dominance demonstrated by unconscious sighting. *Journal of experimental psychology*, *12*, 113.

Mishkin, M., & Forgays, D. G. (1952). Word recognition as a function of retinal locus. *Journal of experimental psychology*, *43*, 43.

Ocklenburg, S., & Güntürkün, O. (2018). *The lateralized brain: The neuroscience and evolution of hemispheric asymmetries*. London: Academic Press.

Oldfield, R. C. (1971). The assessment and analysis of handedness: the Edinburgh inventory. *Neuropsychologia*, *9*, 97-113.

Porac, C., & Coren, S. (1976). The dominant eye. *Psychological bulletin*, *83*, 880.

Rosch, R. E., Bishop, D. V., & Badcock, N. A. (2012). Lateralised visual attention is unrelated to language lateralisation, and not influenced by task difficulty–a functional transcranial Doppler study. *Neuropsychologia*, *50*, 810-815.

Roszkowski, M. J., & Snelbecker, G. E. (1982). Temporal stability and predictive validity of self-assessed hand preference with first and second graders. *Brain and Cognition*, *1*, 405-409.

Van der Haegen, L., & Brysbaert, M. (2018). The relationship between behavioral language laterality, face laterality and language performance in left-handers. *PloS one*, *13*, e0208696.

Van Heuven, W. J., Mandera, P., Keuleers, E., & Brysbaert, M. (2014). SUBTLEX-UK: A new and improved word frequency database for British English. *Quarterly journal of experimental psychology*, *67*, 1176-1190.

Voyer, D. (1998). On the reliability and validity of noninvasive laterality measures. *Brain & Cognition*, *36* , 209-36.

Weekes, N. Y., Capetillo-Cunliffe, L., Rayman, J., Iacoboni, M., & Zaidel, E. (1999). Individual differences in the hemispheric specialization of dual route variables. *Brain and language*, *67*, 110-133.

Westerhausen, R. (2019). A primer on dichotic listening as a paradigm for the assessment of hemispheric asymmetry. *Laterality: Asymmetries of Body, Brain and Cognition*, *24*, 740-771.

Westerhausen, R., & Samuelsen, F. (2020, January 31). An optimal dichotic-listening paradigm for the assessment of hemispheric dominance for speech processing. https://doi.org/10.31234/osf.io/tpjes

Whitehouse, A. J., & Bishop, D. V. (2009). Hemispheric division of function is the result of independent probabilistic biases. *Neuropsychologia*, *47*, 1938-1943.

Willemin, J., Hausmann, M., Brysbaert, M., Dael, N., Chmetz, F., Fioravera, A., Gieruc, K., & Mohr, C. (2016). Stability of right visual field advantage in an international lateralized lexical decision task irrespective of participants’ sex, handedness or bilingualism. *Laterality: Asymmetries of Body, Brain and Cognition*, *21*, 502-524.

Woods, K. J., Siegel, M. H., Traer, J., & McDermott, J. H. (2017). Headphone screening to facilitate web-based auditory experiments. *Attention, Perception, & Psychophysics*, *79*, 2064-2072.

Yovel, G., Tambini, A., & Brandman, T. (2008). The asymmetry of the fusiform face area is a stable individual characteristic that underlies the left-visual-field superiority for faces. *Neuropsychologia*, *46*, 3061-3068.